---
title: "aspect_temp_comparison"
output: html_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#initial set up
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
#packages for exporting csvs as tables
library(knitr)
library(kableExtra)
library(pagedown)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

## Loading in Data

```{r}
persb_aspects <- read_csv("nsf/105E_pers_burned/105E_pers_b_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, sensor_id, temp, aspect, tree_name) %>%
  #filtering out gb_w since we are not using this anymore
  filter(tree_name != "gb_w")

persub_aspects <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_aspects.csv") %>%
   mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, sensor_id, temp, aspect, tree_name) 

transb_aspects <- read_csv("nsf/105E_trans_burned/105E_trans_b_aspects.csv") %>%
   mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, sensor_id, temp, aspect, tree_name)

transub_aspects <- read_csv("nsf/105E_trans_unburned/105E_trans_ub_aspects.csv") %>%
   mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, sensor_id, temp, aspect, tree_name) 
```

## Data Wrangling/Filtering

```{r}
#filtered to correct datetime range

filtered_persb_aspects <- persb_aspects %>%
  mutate(zone = "persistent", burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ zone  # Keep existing values for other cases
  )) %>%
  #filtering to datetime range for analysis
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 23:45:00")

filtered_persub_aspects <- persb_aspects %>%
  mutate(zone = "persistent", burn_status = "live") %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 23:45:00")

filtered_transb_aspects <- transb_aspects %>%
  mutate(zone = "transitional", burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ zone  # Keep existing values for other cases
  )) %>%
  #filtering to datetime range for analysis
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 23:45:00")

filtered_transub_aspects <- transb_aspects %>%
  mutate(zone = "transitional", burn_status = "live") %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 23:45:00")
```

```{r}
#making a full aspect dataframe to work from
filtered_full_aspects <- bind_rows(filtered_persb_aspects, filtered_transb_aspects, filtered_persub_aspects, filtered_transub_aspects)
```

Making an hourly dataset to helps smooth out plot

-   averaging all four trees per site values together(by burn status) for this dataset, so there are less lines

```{r}
filtered_hourly_full_aspect <- filtered_full_aspects %>%
  mutate(time = format(datetime, "%H:%M:%S")) %>%  # Create a new time column
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%  # Create an hourly datetime column
  group_by(datetime_hour, zone, burn_status, aspect) %>%  # Group by the hourly datetime
  #mean_hourly_treetemp is the mean of all the db/gb/live trees in the dataframe averaged by burn_status on an hourly timestep
  summarize(mean_hourly_treetemp = mean(temp), .groups = 'drop') %>%
  mutate(month = month(datetime_hour))
```

```{r}
#making a dt_nt dataframe that can be used to do a daily average of the 
```

```{r}
#making a dataframe to compare ONLY greenburned trees north/south/charred/uncharred since the GB trees are the only trees with the charred/uncharred sensors
gb_hourly_full_aspects <- filtered_hourly_full_aspect %>%
  filter(burn_status == "green burn")

#wide dataset, not long, so I can plot individual columns
gb_wide_hourly_aspects <- gb_hourly_full_aspects %>%
  pivot_wider(
    names_from = aspect,  # Column to become new column names
    values_from = mean_hourly_treetemp      # Column to use for values
  )

#now I am making a charred/uncharred dataset and a north/south dataset

gb_char_full_aspects <- gb_hourly_full_aspects %>%
  filter(aspect %in% c("charred", "uncharred"))

gb_northsouth_full_aspects <- gb_hourly_full_aspects %>%
  filter(aspect %in% c("north", "south"))
```

```{r}
feb_filt_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-02-02 00:00:00" & datetime_hour <= "2024-02-23 23:00:00")

march_filt_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-03-01 00:00:00" & datetime_hour <= "2024-03-31 23:00:00")
```

```{r}
#making first of the month hourly daily timeframes to compare
t_feb2_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-02-02 00:00:00" & datetime_hour <= "2024-02-02 23:00:00") %>%
  filter(zone == "transitional")

p_feb2_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-02-02 00:00:00" & datetime_hour <= "2024-02-02 23:00:00") %>%
  filter(zone == "persistent")

t_march1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-03-01 00:00:00" & datetime_hour <= "2024-03-01 23:00:00") %>%
  filter(zone == "transitional")

p_march1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-03-01 00:00:00" & datetime_hour <= "2024-03-01 23:00:00") %>%
  filter(zone == "persistent")

t_april1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-04-01 00:00:00" & datetime_hour <= "2024-04-01 23:00:00") %>%
  filter(zone == "transitional")

p_april1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-04-01 00:00:00" & datetime_hour <= "2024-04-01 23:00:00") %>%
  filter(zone == "persistent")

t_may1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-05-01 00:00:00" & datetime_hour <= "2024-05-01 23:00:00") %>%
  filter(zone == "transitional")

p_may1_hourly_aspect <- filtered_hourly_full_aspect %>%
  filter(datetime_hour >= "2024-05-01 00:00:00" & datetime_hour <= "2024-05-01 23:00:00") %>%
  filter(zone == "persistent")
```

## Stats Tests

To compare the aspects and get values for how they differ

T-tests - DON'T USE THESE (YET)

```{r}
#running an anova test, because I am comparing three or or more aspects to each other
# Perform a t-test comparing 2 aspects at a time
##north/south
northsouth_aspect_ttest <- t.test(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect[filtered_hourly_full_aspect$aspect %in% c("north", "south"), ])

# Summary 
print(northsouth_aspect_ttest)

##north/charred
northcharred_aspect_ttest <- t.test(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect[filtered_hourly_full_aspect$aspect %in% c("north", "charred"), ])

# Summary 
print(northcharred_aspect_ttest)

##north/uncharred
northuncharred_aspect_ttest <- t.test(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect[filtered_hourly_full_aspect$aspect %in% c("north", "uncharred"), ])

# Summary
print(northuncharred_aspect_ttest)

##south/charred
southcharred_aspect_ttest <- t.test(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect[filtered_hourly_full_aspect$aspect %in% c("south", "charred"), ])

# Summary
print(southcharred_aspect_ttest)

##north/uncharred
northuncharred_aspect_ttest <- t.test(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect[filtered_hourly_full_aspect$aspect %in% c("north", "uncharred"), ])

# Summary
print(northuncharred_aspect_ttest)
```

### Levene Test

```{r}
#install.packages("car")
library(car)
```

```{r}
#first have to check to see if variances are equal using the levenes test
# Levene's Test for aspect
leveneTest(mean_hourly_treetemp ~ aspect, data = filtered_hourly_full_aspect)

# Levene's Test for burn status
leveneTest(mean_hourly_treetemp ~ burn_status, data = filtered_hourly_full_aspect)

# Levene's Test for interaction between aspect and burn status
leveneTest(mean_hourly_treetemp~ aspect * burn_status, data = filtered_hourly_full_aspect)
```

These levene tests show that the variances are unequal (all p-values are significant), therefore I am using a Welchs ANOVA test, which does not assume equal variances

### Welch's ANOVA Test

```{r}
# Welch's ANOVA (no assumption of equal variances)
oneway.test(mean_hourly_treetemp ~ aspect * burn_status, data = filtered_hourly_full_aspect)
```

Results:

data: mean_hourly_treetemp and aspect \* burn_status F = 14.262, num df = 9, denom df = 22081, p-value \< 2.2e-16

Interpretation:

1.  **F-value**: The F-value of 14.262 indicates that there is a significant difference between the group means. In the context of ANOVA, a higher F-value suggests a larger difference between the group means relative to the variability within the groups.

2.  **p-value (\< 2.2e-16)**: The p-value is very small (close to 0), which means that the observed differences in mean temperatures between groups (combinations of aspect and burn status) are highly unlikely to have occurred by random chance.

    -   Since the p-value is less than the standard significance level (usually 0.05), you can **reject the null hypothesis** that the group means are equal.

    -   This implies that there is a statistically significant difference in tree temperatures across different aspects and burn statuses.

### POST-HOC Test:

Since I have a significant result, the next step is to identify which specific groups (combinations of aspect and burn status) differ from each other. You can use a post-hoc test such as **Tukey's Honest Significant Difference (HSD)** to determine pairwise comparisons between the groups.

```{r}
# Tukey's HSD post-hoc test
pairwise_results <- pairwise.t.test(filtered_hourly_full_aspect$mean_hourly_treetemp, interaction(filtered_hourly_full_aspect$aspect, filtered_hourly_full_aspect$burn_status), p.adjust.method = "bonferroni")
```

```{r}
#saving results as dataframe
# Convert the matrix of p-values into a data frame
pairwise_results_df <- as.data.frame(as.table(pairwise_results$p.value))

# Save the data frame to a CSV file
write.csv(pairwise_results_df, "nsf/treetemp_data/stats_pairwise_results.csv", row.names = FALSE)
```

Calculating an effect size: the effect size (such as **eta-squared**) can quantify the magnitude of the difference between groups.

```{r}
library(effectsize)

welch_anova_model <- oneway.test(mean_hourly_treetemp ~ aspect * burn_status, data = filtered_hourly_full_aspect)

eta_squared(welch_anova_model)
```

\*\* need to interpret this effect size calculation more... not sure exactly what it is telling me.

## Plotting

```{r}
feb_aspect_BP <- ggplot(feb_filt_hourly_aspect, aes(x = aspect, y = mean_hourly_treetemp)) +
  geom_boxplot() +
  labs(title = "Feb Values by Aspect", x = "aspect", y = "mean_hourly_treetemp") +
  theme_bw() +
  facet_wrap (~ zone + burn_status)

feb_aspect_BP
```

```{r}
#to display the colors
display.brewer.pal(n = 8, name = "Dark2")

#to get the hex number of the colors
brewer.pal(n = 8, name = "Dark2")
```

```{r}
#setting a color-blind friendly custom palette for gb/db/ubn- using RColorBrewer
##blue is unburned, orange is dead burned, green is green burned
burnstatus_colors <- c("#D95F02","#1B7837", adjustcolor("#2166AC", alpha = 0.7))

aspect_colors <- c("#D95F02","#1B7837", "#666666","#2166AC")
```

```{r}
aspect_comparison <- ggplot(data = feb_filt_hourly_aspect) +
  geom_line(aes(x = datetime_hour, y = mean_hourly_treetemp, color = burn_status)) + 
  facet_wrap(~zone + aspect) +
  scale_color_manual(values = burnstatus_colors) +  # Apply custom colors
  labs(color = "Burn Status") +  # Add a label for the color legend
  theme_bw()

aspect_comparison
```

Plotting Individually, then Combining

```{r}
#dataframes for plotting individually
trans_filt_hourly_aspects <- filtered_hourly_full_aspect %>%
  filter( zone == "transitional")

pers_filt_hourly_aspects <- filtered_hourly_full_aspect %>%
  filter( zone == "persistent")
```

```{r}
trans_aspect_comparison <- ggplot(data = trans_filt_hourly_aspects ) +
  geom_line(aes(x = datetime_hour, y = mean_hourly_treetemp, color = burn_status)) + 
  facet_wrap(~zone) + 
  facet_grid(burn_status ~ aspect) +
  scale_color_manual(values = burnstatus_colors) +  # Apply custom colors
  labs(color = "Burn Status") +  # Add a label for the color legend
  theme_bw()

trans_aspect_comparison
```

```{r}
#one-day comparison
t_feb2_aspect_comparison <- ggplot(data = t_feb2_hourly_aspect) +
  geom_line(aes(x = datetime_hour, y = mean_hourly_treetemp, color = burn_status)) + 
  facet_wrap(~zone + aspect) +
  scale_color_manual(values = burnstatus_colors) +  # Apply custom colors
  labs(color = "Burn Status") +  # Add a label for the color legend
  theme_bw()

t_feb2_aspect_comparison 

## should also do this for persistent zone then combine the plots
```

```{r}
t_march1_aspect_comparison <- ggplot(data = t_march1_hourly_aspect) +
  geom_line(aes(x = datetime_hour, y = mean_hourly_treetemp, color = burn_status)) + 
  facet_wrap(~zone + aspect) +
  scale_color_manual(values = burnstatus_colors) +  # Apply custom colors
  labs(color = "Burn Status") +  # Add a label for the color legend
  theme_bw()

t_march1_aspect_comparison 

p_march1_aspect_comparison <- ggplot(data = p_march1_hourly_aspect) +
  geom_line(aes(x = datetime_hour, y = mean_hourly_treetemp, color = burn_status)) + 
  facet_wrap(~zone + aspect) +
  scale_color_manual(values = burnstatus_colors) +  # Apply custom colors
  labs(color = "Burn Status") +  # Add a label for the color legend
  theme_bw()

p_march1_aspect_comparison 
```

Plotting only the GB trees to see if there is a charred/uncharred importance for tree temps

```{r}
gb_aspects_plot <- ggplot(data = gb_wide_hourly_aspects) +
  geom_line(aes(x = datetime_hour, y = north), color = "green") + 
  geom_line(aes(x = datetime_hour, y = south), color = "red") +
  geom_line(aes(x = datetime_hour, y = charred), color = "black", linetype = "dashed") +
  geom_line(aes(x = datetime_hour, y = uncharred), color = "grey", linetype = "dashed") +
  facet_wrap(~zone) +
  labs(color = "Aspect") +  # Add a label for the color legend
  theme_bw()

gb_aspects_plot 

## this plot is too confusing... need to do a daily timestep so there is less variation and the data is easier to process... 
```
