---
title: "daynight_comparison"
author: "Erika Lee"
date: "2024-07-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#initial set up
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(patchwork)
library(RColorBrewer)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

# Loading in Dataframes

```{r}
#reading in dataframes that I need - burned
pers_b_aspects <- read_excel("nsf/105E_pers_burned/105E_pers_b_aspects.xlsx") %>%
  mutate(zone = "persistent")

trans_b_aspects <- read_excel("nsf/105E_trans_burned/105E_trans_b_aspects.xlsx") %>%
  mutate(zone = "transitional")

pers_burned_rad_air <- read_excel("nsf/pers_burned/pers_burned_rad_air.xlsx")

trans_burned_rad_air <- read_excel("nsf/trans_burned/trans_burned_rad_air.xlsx")

#read in dataframes that I need - unburned
pers_ub_aspects <- read_excel("nsf/105E_pers_unburned/105E_pers_ub_aspects.xlsx") %>%
  mutate(zone = "persistent")

trans_ub_aspects <- read_excel("nsf/105E_trans_unburned/105E_trans_ub_aspects.xlsx") %>%
  mutate(zone = "transitional")
```

# Data Wrangling

```{r}
#setting a timeframe
start_datetime <- ymd_hms("2024-02-02 00:00:00")
end_datetime <- ymd_hms("2024-05-25 00:00:00")

#setting a day and night time 
daytime <- hms("08:30:00")
nighttime <- hms("17:30:00")

#creating a dataframe that has all aspect temps, and includes zone and burn status
ns_avg_treetemp_data <- bind_rows(pers_b_aspects, trans_b_aspects, pers_ub_aspects, trans_ub_aspects) %>%
  #creating a timeframe with the same datetimes
  filter(datetime >= start_datetime & datetime <= end_datetime) %>% filter(aspect == c("north", "south")) %>%
  mutate(time = hms(format(datetime, "%H:%M:%S")),  # Extract time from datetime
         phase = case_when(
           time >= daytime & time < nighttime ~ "day",
           time < daytime | time >= nighttime ~ "night",
           TRUE ~ NA_character_  # If none of the conditions match, handle accordingly
         )) %>%
  #creating a burn status row
   mutate(burn_status = case_when(
      startsWith(tree_name, "gb") ~ "green burned",
      startsWith(tree_name, "db") ~ "dead burned",
      startsWith(tree_name, "ub") ~ "unburned",
      TRUE ~ "unknown"  # Default case if none of the above conditions match
    )) %>%
  group_by(datetime, phase, zone, burn_status, aspect) %>%
  summarise(avg_temp = mean(temp, na.rm = TRUE)) %>%
  ungroup()
## this dataframe has all zones and burn statuses, as well as only north and south aspect average temps for every 15 minute increment

#creating an all_averaged temps dataset (not north/south) - this is averaging the n/s temps
all_avg_treetemp_data <- bind_rows(pers_b_aspects, trans_b_aspects, pers_ub_aspects, trans_ub_aspects) %>%
  #creating a timeframe with the same datetimes
  filter(datetime >= start_datetime & datetime <= end_datetime) %>% filter(aspect == c("north", "south")) %>%
  mutate(time = hms(format(datetime, "%H:%M:%S")),  # Extract time from datetime
         phase = case_when(
           time >= daytime & time < nighttime ~ "day",
           time < daytime | time >= nighttime ~ "night",
           TRUE ~ NA_character_  # If none of the conditions match, handle accordingly
         )) %>%
  #creating a burn status row
   mutate(burn_status = case_when(
      startsWith(tree_name, "gb") ~ "green burned",
      startsWith(tree_name, "db") ~ "dead burned",
      startsWith(tree_name, "ub") ~ "unburned",
      TRUE ~ "unknown"  # Default case if none of the above conditions match
    )) %>%
  group_by(datetime, phase, zone, burn_status) %>%
  summarise(avg_temp = mean(temp, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
#writing full dataframe to N-drive
write.xlsx(ns_avg_treetemp_data, "nsf/ns_avg_treetemps.xlsx")

write.xlsx(all_avg_treetemp_data,"nsf/all_avg_treetemps.xlsx")
```

# Plotting

## NS differences for all avg temps

```{r}
color_palette_manual <- c("green burned" = "royalblue2", 
                          "dead burned" = "red3",
                          "unburned" = "#008000")

dtnt_ns_avg_temps_plots <- ggplotly(
  ggplot() +
    geom_line(data = ns_avg_treetemp_data, aes(x = datetime, y = avg_temp, color = burn_status, linetype = aspect), alpha = 0.8) + 
    scale_color_manual(values = color_palette_manual, name = "burn status,") +
    facet_wrap(~zone + phase) +
    labs(x = "Datetime", y = "Avg Temp (C)", title = "Day vs nighttime average temp by sensor aspect") +
    scale_linetype_manual(values = c("dashed", "solid")) # Flipped the order of linetypes
)

dtnt_ns_avg_temps_plots
```

## Total Avg Temps (n/s averaged)

```{r}
#this plot shows the total average temperature of all aspects, not divided by aspects
dtnt_all_avg_temps_plots <- ggplotly(
  ggplot() +
    geom_line(data = all_avg_treetemp_data, aes(x = datetime, y = avg_temp, color = burn_status, linetype = zone), alpha = 0.7) + 
    scale_color_manual(values = color_palette_manual, name = "burn status,") +
    facet_wrap(~phase) +
    labs(x = "Datetime", y = "Avg Temp (C)", title = "Day vs nighttime average temp of all sensors, by zone") +
    scale_linetype_manual(values = c("solid", "dashed")) # Flipped the order of linetypes
)

dtnt_all_avg_temps_plots
```

## Saving plotlys

```{r}
#ns plots
htmlwidgets::saveWidget(dtnt_ns_avg_temps_plots, "daynight_northsouth_allplots.html")

#all avg plots
htmlwidgets::saveWidget(dtnt_all_avg_temps_plots, "daynight_allavg_allplots.html")
```

#next, need to look at including SWin for these plots?

# Creating daily/nightly single value average dataframe

## Data Wrangling

```{r}
#creating dataframes with a single average temp per day - for all tree temps
single_avg_all_treetemps <- all_avg_treetemp_data %>%
  #adding a date column
  mutate(date = as_date(datetime)) %>%
  group_by(date, phase, zone, burn_status) %>%
  summarise(avg_temp = mean(avg_temp, na.rm = TRUE)) %>%
  ungroup()

#creating a dataframe with single average temps per day and per aspect
single_avg_ns_all_treetemps <- ns_avg_treetemp_data %>%
  mutate(date = as_date(datetime)) %>%
  group_by(date, phase, zone, burn_status, aspect) %>%
  summarise(avg_temp = mean(avg_temp, na.rm = TRUE)) %>%
  ungroup()
```

## Plotting

### NS difference for single day/night avg temp

```{r}
single_dtnt_ns_avg_temps_plots <- ggplotly(
  ggplot() +
    geom_line(data = single_avg_ns_all_treetemps, aes(x = date, y = avg_temp, color = burn_status, linetype = aspect), alpha = 0.8) + 
    scale_color_manual(values = color_palette_manual, name = "burn status,") +
    facet_wrap(~zone + phase) +
    labs(x = "Datetime", y = "Avg Temp (C)", title = "Day vs nighttime average temp by sensor aspect") +
    scale_linetype_manual(values = c("dashed", "solid")) # Flipped the order of linetypes
)

single_dtnt_ns_avg_temps_plots
```

### Total Avg Temps (n/s averaged together) for single day/night value

```{r}
single_dtnt_avg_all_temps_plots <- ggplotly(
  ggplot() +
    geom_line(data = single_avg_ns_all_treetemps, aes(x = date, y = avg_temp, color = burn_status, linetype = zone), alpha = 0.7) + 
    scale_color_manual(values = color_palette_manual, name = "burn status,") +
    facet_wrap(~phase) +
    labs(x = "Datetime", y = "Avg Temp (C)", title = "Day vs nighttime average temp of all sensors, by zone") +
    scale_linetype_manual(values = c("solid", "dashed")) # Flipped the order of linetypes
)

single_dtnt_avg_all_temps_plots
```

### Saving Plots

```{r}
#ns single plots
htmlwidgets::saveWidget(single_dtnt_ns_avg_temps_plots, "singlevalue_daynight_northsouth_allplots.html")

#all avg single plots
htmlwidgets::saveWidget(single_dtnt_avg_all_temps_plots, "singlevalue_daynight_allavg_allplots.html")
```
