---
title: "radiation_plot_calcs"
output: html_document
date: "2024-09-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(hms)
library(ggpubr)
#package for scientific notation of numbers
library(scales)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#to display the colors
display.brewer.pal(n = 8, name = "Dark2")

#to get the hex number of the colors
brewer.pal(n = 8, name = "Dark2")
```

```{r}
#setting a color-blind friendly custom palette for gb/db/ubn- using RColorBrewer
##blue is unburned, orange is dead burned, green is green burned
three_custom_colors <- c("#D95F02","#1B7837", adjustcolor("#2166AC", alpha = 0.7))

## orange is burned, blue is unburned
two_custom_colors <- c("#D95F02", adjustcolor("#2166AC", alpha = 0.7))
```

## Calling in Data

```{r}
transb_energytree_table <- read_csv("nsf/drone_imagery/trans_b_energytree_table.csv") %>%
  mutate(zone = "transitional") %>%
  rename(burn_status = burn_status) %>%
  #adding a tree_id column
  mutate(tree_id = row_number()) %>%
  #selecting columns I want
  select(tree_id, zone, burn_status, lat, long)

#this is the csv that has tree diameters for albedo plots on it, use this to choose an average circumference
albedoplots_treechar_alldata <- read_csv("nsf/albedo/albedoplots_treechar_alldata.csv")

#this is the csv that has camera snow depths and if tree wells appeared or not from 105E_transb_snowcamera
transb_snowcamera_data <- read_csv("nsf/camera_snowdepths/2024_105E_transb_snowdepths.csv") %>%
  rename(snowcam_max_depth_cm = "maximum_depth(cm)") %>%
  mutate(date = as.Date(date))

transb_wx_15min <- read_csv("nsf/trans_burned/trans_burned_wx_15min_r.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST'))

#seperating by burned/unburned
##burned
avg_dt_transb_treeLW_burned <- read_csv("nsf/105E_trans_burned/day_avg_filtered_treeLWout_data_bub.csv") %>%
  filter(zone == "transitional", phase == "day", burn_status == "burned") %>%
  rename(b_tree_LWout_MJd = tree_LWout_MJperd, b_treetemp_C = temp, b_tree_LWout_wperm2 = tree_LWout_wperm2)

##unburned
avg_dt_transb_treeLW_unburned <- read_csv("nsf/105E_trans_burned/day_avg_filtered_treeLWout_data_bub.csv") %>%
  filter(zone == "transitional", phase == "day", burn_status == "unburned") %>%
  rename(ub_tree_LWout_MJd = tree_LWout_MJperd, ub_treetemp_C = temp, ub_tree_LWout_wperm2 = tree_LWout_wperm2)


#seperating by tree burn status - gb/db/ub - this dataframe includes all burn status types
avg_dt_transb_treeLW_allstatuses <- read_csv("nsf/105E_trans_burned/day_avg_filtered_treeLWout_data_gbdbub.csv")

##green burned dataframe
avg_dt_transb_treeLW_gb <- avg_dt_transb_treeLW_allstatuses %>%
  #filtering out only the correct burned status, daytime data, & transitional zone
  filter(zone == "transitional", phase == "day", burn_status == "green burn") %>%
  rename(gb_tree_LWout_MJd = tree_LWout_MJperd, gb_treetemp_C = temp_C, gb_tree_LWout_wperm2 = tree_LWout_wperm2, gb_tree_LWout_Jperd = tree_LWout_Jperd) %>%
  #removing double ID first column that appeared for some reason?
  select(-1)

##dead burned dataframe
avg_dt_transb_treeLW_db <- avg_dt_transb_treeLW_allstatuses %>%
  #filtering out only the correct burned status, daytime data, & transitional zone
  filter(zone == "transitional", phase == "day", burn_status == "dead burn") %>%
  rename(db_tree_LWout_MJd = tree_LWout_MJperd, db_treetemp_C = temp_C, db_tree_LWout_wperm2 = tree_LWout_wperm2, db_tree_LWout_Jperd = tree_LWout_Jperd) %>%
  #removing double ID first column that appeared for some reason?
  select(-1)

##unburned dataframe
avg_dt_transb_treeLW_ub <- avg_dt_transb_treeLW_allstatuses %>%
  #filtering out only the correct burned status, daytime data, & transitional zone
  filter(zone == "transitional", phase == "day", burn_status == "unburned") %>%
  rename(ub_tree_LWout_MJd = tree_LWout_MJperd, ub_treetemp_C = temp_C, ub_tree_LWout_wperm2 = tree_LWout_wperm2, ub_tree_LWout_Jperd = tree_LWout_Jperd) %>%
  #removing double ID first column that appeared for some reason?
  select(-1)
```

## Filtering Data

```{r}
filtered_transb_rad_wx <- transb_wx_15min %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00") & datetime <= as.POSIXct("2024-05-25 00:00:00")) %>%
   # Adding in date, time, and phase columns
  mutate(
    date = as.Date(datetime),
    time = format(datetime, "%H:%M:%S"),
    time_hms = hms::as_hms(time),  # Convert to hms object for phase calculation
    phase = case_when(
      time_hms >= hms::as_hms("08:00:00") & time_hms < hms::as_hms("17:00:00") ~ "day",
      time_hms < hms::as_hms("08:00:00") | time_hms >= hms::as_hms("17:00:00") ~ "night",
      TRUE ~ NA_character_)) %>%
      select(datetime, date, time, phase, SnowDepth_m, AirTC_Avg, SWin_Avg, SWout_Avg, SWalbedo_Avg, LWin_Avg, LWout_Avg, SWnet_Avg, LWnet_Avg, NR_Avg)
```

```{r}
#filtering snow depth camera data
filtered_transb_snowcamera_data <- transb_snowcamera_data %>%
  filter(date >= as.POSIXct("2024-02-02") & date <= as.POSIXct("2024-05-24")) %>%
  mutate(snowcam_max_depth_m = snowcam_max_depth_cm/100) %>%
  select(date, snowcam_max_depth_cm, snowcam_max_depth_m, conditions, treewell_developed)
```

## Data Wrangling

Creating a daily average wx station dataframe, then converting to MJ/day

```{r}
#calculations for radiation conversions from W/m2 to MJ/day
## create a daily average dataframe
avg_dt_transb_rad_wx<- filtered_transb_rad_wx %>%
  group_by(date, phase) %>%
  summarize(across(everything(), mean, na.rm = TRUE)) %>%
  select(-c(datetime, time)) %>%
  group_by(date, phase) %>%
  # Convert mean radiation to MJ/day - assuming constent for m2 value
  mutate(SWin_MJd = (SWin_Avg/ 1e6) * 86400, SWout_MJd = (SWout_Avg/ 1e6) * 86400, LWin_MJd = (LWin_Avg/ 1e6) * 86400, LWout_MJd = (LWout_Avg/ 1e6) * 86400, SWnet_MJd = (SWnet_Avg / 1e6) * 86400, LWnet_MJd = (LWnet_Avg/ 1e6) * 86400, NR_MJd = (NR_Avg / 1e6) * 86400) %>%
  rename(SWin_wperm2 = SWin_Avg, SWout_wperm2 = SWout_Avg, SWnet_wperm2 = SWnet_Avg, LWin_wperm2 = LWin_Avg, LWout_wperm2 = LWout_Avg, LWnet_wperm2 = LWnet_Avg, NR_wperm2 = NR_Avg) %>%
  mutate(zone = "transitional", burn_status = "burned") %>%
  #filtering out only daytime values
  filter(phase == "day") %>%
  #reordering so the like values are next to each other
  select(date, phase, zone, burn_status, SnowDepth_m, AirTC_Avg, SWin_wperm2, SWin_MJd, SWout_wperm2, SWout_MJd, SWnet_wperm2, SWnet_MJd, SWalbedo_Avg, LWin_wperm2, LWin_MJd, LWout_wperm2, LWout_MJd, LWnet_wperm2, LWnet_MJd, NR_wperm2, NR_MJd) %>%
  #setting decimal number
   mutate(across(SnowDepth_m:last_col(), ~ round(., 4)))
```

Combining data into one dataframe by burn_status

```{r}
##* I think I want to use this one, not the gb/db/ub, since there is such a small difference between the gb & db LWout values from trees
#by burned/unburned status
transb_comb_radtree <- avg_dt_transb_rad_wx %>%
  mutate(burn_status = "burned") %>%
  full_join(avg_dt_transb_treeLW_burned) %>%
  #recalculating NetRad, as I am not sure the calculations here are correct that are pulled from the weather station
  mutate(NR_MJd_wxs = (SWin_MJd - SWout_MJd) + (LWin_MJd - LWout_MJd)) %>%
  select(-datetime)
#answer is slightly different than weather-station calculated RN, use this value instead!
```

```{r}
#by gb/db/ub status
transb_comb_radtree_gb <- avg_dt_transb_rad_wx %>%
  mutate(burn_status = "green burn") %>%
  full_join(avg_dt_transb_treeLW_gb)

transb_comb_radtree_db <- avg_dt_transb_rad_wx %>%
  mutate(burn_status = "dead burn") %>%
  full_join(avg_dt_transb_treeLW_db)

#won't have data to work with for unburned plots comparison as I don't have drone imagery for this area yet
transb_comb_radtree_ub <- avg_dt_transb_rad_wx %>%
  mutate(burn_status = "unburned") %>%
  full_join(avg_dt_transb_treeLW_ub)
```

```{r}
#getting tree circumfirence max and min for all transitional albedo plots - this will help me ID what average circumference to use
transb_treecircm_data <- albedoplots_treechar_alldata %>%
  #filtering for only transitional burned zone plots
  filter(PlotID %in% c("TLHM", "TLLM")) %>%
  select(PlotID, TreeNum, Species, DBH_cm, PreFire_Status) %>%
  mutate(DBH_m = DBH_cm / 100,
         circ_m = DBH_m * pi) %>%
  summarize(DBH_m_min = min(DBH_m),
    DBH_m_max = max(DBH_m),
    DBH_m_mean = mean(DBH_m),
    circ_m_min = min(circ_m),
    circ_m_max = max(circ_m),
    circ_m_mean = mean(circ_m)
  )
```

```{r}
#setting values based on the min, max and mean circumference of the albedo plots - these will be used in later calculations
min_circ_m <- 0.1979
max_circ_m <- 0.6158
mean_circ_m <- 0.4042
```

Creating a dataframe to add in net tree LWout including all trees in plot area

```{r}
#there are 219 total burned trees in the transitional burned plot area
transb_comb_treeLWout <- avg_dt_transb_treeLW_burned %>%
  select(-c(datetime, zone, burn_status)) %>%
  crossing(transb_energytree_table) %>%
  select(date, phase, zone, burn_status, tree_id, everything()) %>%
  select(-c(lat, long)) %>%
  mutate(min_circ_m = min_circ_m, max_circ_m = max_circ_m, mean_circ_m = mean_circ_m)
```

## Calcs - adding in all trees to LWout data

```{r}
#dataframes seperates by burn_status

gb_transb_energytree_table <- transb_energytree_table %>%
  filter(burn_status == "gb")
#155 gb trees

db_transb_energy_treetabl <- transb_energytree_table %>%
  filter(burn_status == "db")
#64 db trees
```

## Plotting

```{r}
color_values_MJd <- c(
  "SWin_MJd" = "#D95F02",
  "SWout_MJd" = "#E6AB02",
  "LWout_MJd" = "#66A61E",
  "LWin_MJd" = "#2e5e39",
  "gb_tree_LWout_MJd" = "#7570B3",
  "db_tree_LWout_MJd" = "#2166AC")
# will need to include an unburned color later on if I include unburned data in these plot comparisons!

color_values_wperm2 <- c(
  "SWin_wperm2" = "#D95F02",
  "SWout_wperm2" = "#E6AB02",
  "LWout_wperm2" = "#66A61E",
  "LWin_wperm2" = "#2e5e39",
  "gb_tree_LWout_MJd" = "#7570B3",
  "db_tree_LWout_MJd" = "#2166AC")

# Define line types
line_types <- c(
  "Weather Station" = "solid",
  "Tree Temperature Sensors" = "dashed"
)

#changing labels of values in plots for the legend
color_labels <- c(
  "SWin_Avg" = "Shortwave In (Avg)",
  "SWout_Avg" = "Shortwave Out (Avg)",
  "LWin_Avg" = "Longwave In (Avg)",
  "LWout_Avg" = "Longwave Out (Avg)",
  "tree_LWout_avg_wperm2" = "Tree LW Out (Avg)"  # New label for the specific line
)
```

```{r}
#MJ/day plot
gb_combined_day_rad_wx_plot_MJd <- ggplot(data = transb_comb_radtree_gb) + 
  geom_line(aes(x = date, y = SWin_MJd, color = "SWin_MJd", linetype = "Weather Station"), linewidth = 0.6, alpha = 0.9) +
  geom_line(aes(x = date, y = SWout_MJd, color = "SWout_MJd", linetype = "Weather Station"), linewidth = 0.6, alpha = 0.8) +
  geom_line(aes(x = date, y = LWin_MJd, color = "LWin_MJd", linetype = "Weather Station"), linewidth = 0.7, alpha = 0.9) +
  geom_line(aes(x = date, y = LWout_MJd, color = "LWout_MJd", linetype = "Weather Station"), linewidth = 0.7) +
  geom_line(aes(x = date, y = gb_tree_LWout_MJd, color = "gb_tree_LWout_MJd", linetype = "Tree Temperature Sensors")) +
  facet_wrap(~ zone + burn_status, scales = "free_y", dir = "h") +  # Facet by zone and phase
  labs(
    #title = "Transitional Unburned",
    x = "Date",
    y = "Radiation components (MJ/day)",
    color = "Variable", 
    linetype = "Data Source"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values_MJd, labels = color_labels) + 
  scale_linetype_manual(values = line_types)

gb_combined_day_rad_wx_plot_MJd
```

```{r}
#W/m2 plot
gb_combined_day_rad_wx_plot_Wperm2 <- ggplot(data = transb_comb_radtree_gb) + 
  geom_line(aes(x = date, y = SWin_wperm2, color = "SWin_wperm2", linetype = "Weather Station"), linewidth = 0.6, alpha = 0.9) +
  geom_line(aes(x = date, y = SWout_wperm2, color = "SWout_wperm2", linetype = "Weather Station"), linewidth = 0.6, alpha = 0.8) +
  geom_line(aes(x = date, y = LWin_wperm2, color = "LWin_wperm2", linetype = "Weather Station"), linewidth = 0.7, alpha = 0.9) +
  geom_line(aes(x = date, y = LWout_wperm2, color = "LWout_wperm2", linetype = "Weather Station"), linewidth = 0.7) +
  geom_line(aes(x = date, y = gb_tree_LWout_wperm2, color = "gb_tree_LWout_wperm2", linetype = "Tree Temperature Sensors")) +
  facet_wrap(~ zone + burn_status, scales = "free_y", dir = "h") +  # Facet by zone and phase
  labs(
    #title = "Transitional Unburned",
    x = "Date",
    y = "Radiation components (W/m2)",
    color = "Variable", 
    linetype = "Data Source"
    ) +
  theme_bw() +  # Use a minimal theme for a clean look
  scale_color_manual(values = color_values_wperm2, labels = color_labels) +
  scale_linetype_manual(values = line_types)

combined_day_rad_wx_plot_Wperm2
```

\*\* interpretation of difference of MJperday and Wperm2 plots - the Wperm2 plot is showing second-timeframe data, not daily data, which means that the tree temps will be more equal in value to what the weather station is showing. The MJperday plot is showing a daily value for all the variables - once I add all the individual tree values together, perhaps the line for tree_LWout will increase a bit above zero?

Net Radiation Plots - by burn_status

-   the following graph is not really helpful yet, since I have not included all the tree_LWout values in the total net radiation value, that's why the tree_LWout values are plotted as higher than net rad from weather station

-   Need to finish calculations, then this plot can be used!

```{r}
#green burn net radiaiton from WX station and tree RLout
gb_netrad_plot <- ggplot() + 
  geom_col(data = transb_comb_radtree_gb, aes(x = date, y = NR_MJd, fill = "Net Radiation"), position = "stack") +  # Add fill aesthetic to legend
  geom_line(data = transb_comb_radtree_gb, aes(x = date, y = gb_tree_LWout_MJd, color = "Green burned LWout")) +
  geom_line(data = transb_comb_radtree_db, aes(x = date, y = db_tree_LWout_MJd, color = "Dead burned LWout")) +
  facet_wrap(~zone) +
  labs(
    x = "Date",
    y = "Radiation Component (MJ/day)") +
    scale_fill_manual(name = "Weather Station Data", values = "#2166AC") +
  scale_color_manual(
    name = "Tree Temp Sensor Data", 
    values = c("Green burned LWout" = "#1B7837",
               "Dead burned LWout" = "#D95F02")) +
  theme_bw()

gb_netrad_plot
```

```{r}
## only need this plot if I am seperating the net radiation plots by burned status, but the above plot now has the GB and DB data on it
#dead burn net radiaiton from WX station and tree RLout
db_netrad_plot <- ggplot(data = transb_comb_radtree_db) + 
  geom_col(aes(x = date, y = NR_MJd, fill = "Net Radiation (MJ/d)"), position = "stack") +  # Add fill aesthetic to legend
  geom_line(aes(x = date, y = db_tree_LWout_MJd, color = "Dead burned LWout (MJ/d)")) +
  facet_wrap(~zone + burn_status) +
  labs(
    x = "Date",
    y = "Radiation Component (MJ/day)") +
    scale_fill_manual(name = "Weather Station Data", values = "#2166AC") +                         # Custom fill color
  scale_color_manual(name = "Tree Temp Sensor Data", values = "#D95F02") +
  theme_bw()

db_netrad_plot
```

Plotting net SW and LW to see if this matched wyatts data?

```{r}
netrad_plot <- ggplot(data = transb_comb_radtree_data) +
  geom_col(aes(x = date, y = SWnet_wperm2, fill = "SWnet_wperm2"), position = "stack") +
  geom_col(aes(x = date, y = LWnet_wperm2, fill = "LWnet_wperm2"), position = "stack") +
   labs(x = "Date", y = "Net Radiation (W/m²)", fill = "Radiation Type") +
  scale_fill_manual(values = c("SWnet_wperm2" = "orange", "LWnet_wperm2" = "blue")) +  # Custom colors
  theme_bw()

netrad_plot
```

\*creating a bar plot for radiation and a

Looking at Tree LWout Burn Status Differences

```{r}
tree_LWout_data <- ggplot() +
  geom_line(data = avg_dt_transb_treeLW_gb, aes(x = date, y = tree_LWout_MJd), color = "green") +
    geom_line(data = avg_dt_transb_treeLW_db, aes(x = date, y = tree_LWout_MJd), color = "red") +
    geom_line(data = avg_dt_transb_treeLW_ub, aes(x = date, y = tree_LWout_MJd), color = "blue")

tree_LWout_data
```
