---
title: "treetemp_stats"
author: "Erika Lee"
date: "2024-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stats Calculations

```{r}
#initial set up
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
#packages for exporting csvs as tables
library(knitr)
library(kableExtra)
library(pagedown)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

## Loading Data

```{r}
final_15min_ns_b <- read_csv("nsf/treetemp_data/final_15min_ns_b.csv")

final_15min_ns_ub <- read_csv("nsf/treetemp_data/final_15min_ns_ub.csv")

final_hourly_ns_b <- read_csv("nsf/treetemp_data/final_hourly_ns_b.csv")

final_hourly_ns_ub <- read_csv("nsf/treetemp_data/final_hourly_ns_ub.csv")

final_15min_gbdb <- read_csv("nsf/treetemp_data/final_15min_gbdb.csv")

final_15min_gbdb_ub <- read_csv("nsf/treetemp_data/final_15min_gbdb_ub.csv")

final_15min_gbub <- read_csv("nsf/treetemp_data/final_15min_gbub.csv")

final_15min_dbub <- read_csv("nsf/treetemp_data/final_15min_dbub.csv")

final_hourly_gbdb <- read_csv("nsf/treetemp_data/final_hourly_gbdb.csv")

final_hourly_gbdb_ub <- read_csv("nsf/treetemp_data/final_hourly_gbdb_ub.csv")

final_hourly_gbub <- read_csv("nsf/treetemp_data/final_hourly_gbub.csv")

final_hourly_dbub <- read_csv("nsf/treetemp_data/final_hourly_dbub.csv")

final_daynight_singleavg_temps <- read_csv("nsf/treetemp_data/final_daynight_singleavg_temps.csv")
```

```{r}
#new full dataframes to load
full_15min_ns <- read_csv("nsf/treetemp_data/full_15min_ns.csv")

full_hourly_ns <- read_csv("nsf/treetemp_data/full_hourly_ns.csv")

full_15min_gbdbub <- read_csv("nsf/treetemp_data/full_15min_gbdbub.csv")

full_hourly_gbdbub <- read_csv("nsf/treetemp_data/full_hourly_gbdbub.csv")
```

Pulling in Data for gb only comparisons of charred/uncharred data with n/s as well

```{r}
final_15min_gb_temps <- read_csv("nsf/treetemp_data/final_15min_gb_temps.csv") %>%
  mutate(day = as.Date(datetime), month = month(datetime))

final_hourly_gb_temps <- read_csv("nsf/treetemp_data/final_hourly_gb_temps.csv")%>%
  mutate(day = as.Date(datetime_hour), month = month(datetime_hour))
```

## Calculating Stats

### N/S Differenced Data

Based on meeting w/ Stephanie on 7/17/24 - I already calculated avg_temp for 15min timestep in "data_wrangling" rmarkdown, now I will calculate the mean_diff for 15minute timestep, then do the other calculations by month

```{r}
## don't use this one!
#burned n/s stats
stats_monthly_ns_b <- final_15min_ns_b %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

```{r}
## don't use this once! Use the one lower down with combined burned/ub
#unburned n/s stats
stats_monthly_ns_ub <- final_15min_ns_ub %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

#### Creating N/S differencing tables that are organized by burn status (gb, db, ub)

```{r}
#green burned n/s differences
stats_monthly_ns_gb <- stats_monthly_ns_b %>%
  filter(burn_status == "green burn")

#dead burned n/s differences
stats_monthly_ns_db <- stats_monthly_ns_b %>%
  filter(burn_status == "dead burn")

##don't need to do for ub because I already have that
```

### North/South Differenced Data

\*\* differenced in the same way as the charred/uncharred below!

```{r}
stats_monthly_s_n <- final_15min_gb_temps %>%
  mutate(datetime = as.POSIXct(datetime), month = format(datetime, "%m")) %>%
  #calculating temp_diff for 15-minute interval of charred/uncharred
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(south_temp), mean_north_temp = mean(north_temp), mean_temp = mean((south_temp + north_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### Charred/Uncharred Differenced Data

```{r}
#burned n/s stats
stats_monthly_c_uc <- final_15min_gb_temps %>%
  mutate(datetime = as.POSIXct(datetime), month = format(datetime, "%m")) %>%
  #calculating temp_diff for 15-minute interval of charred/uncharred
  mutate(temp_diff = charred_temp - uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_charred_temp = mean(charred_temp), mean_uncharred_temp = mean(uncharred_temp), mean_temp = mean((charred_temp + uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### South / Uncharred Differenced Data

```{r}
# to compare noth/south with charred/uncharred, I need to make new n/s_gb only datasets
stats_monthly_s_uc <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_south_temp, mean_uncharred_temp) %>%
  mutate(temp_diff = mean_south_temp - mean_uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(mean_south_temp), mean_uncharred_temp = mean(mean_uncharred_temp), mean_temp = mean((mean_south_temp + mean_uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### South / Charred Differenced Data

```{r}
stats_monthly_s_c <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_south_temp, mean_charred_temp) %>%
  mutate(temp_diff = mean_south_temp - mean_charred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(mean_south_temp), mean_charred_temp = mean(mean_charred_temp), mean_temp = mean((mean_south_temp + mean_charred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### North / Uncharred Differenced Data

```{r}
stats_monthly_n_uc <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_north_temp, mean_uncharred_temp) %>%
  mutate(temp_diff = mean_north_temp - mean_uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_uncharred_temp = mean(mean_uncharred_temp), mean_temp = mean((mean_north_temp + mean_uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### North / Charred Differenced Data

```{r}
stats_monthly_n_c <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_north_temp, mean_charred_temp) %>%
  mutate(temp_diff = mean_north_temp - mean_charred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_charred_temp = mean(mean_charred_temp), mean_temp = mean((mean_north_temp + mean_charred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/DB Differenced Data

```{r}
#gb/db stats - db minus gb! 
stats_monthly_gbdb <- final_15min_gbdb %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = db_temp - gb_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_gb_temp = mean(gb_temp), mean_db_temp = mean(db_temp), mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/UB Differenced Data

```{r}
#stats calculations
stats_monthly_gbub <- final_15min_gbub %>%
  mutate(temp_diff = gb_temp - ub_temp) %>%
  # Calculating monthly stats
  group_by(month, zone) %>%
  summarize(
    mean_gb_temp = mean(gb_temp, na.rm = TRUE),
    mean_ub_temp = mean(ub_temp, na.rm = TRUE),
    mean_temp = mean(avg_temp, na.rm = TRUE),
    mean_diff = mean(temp_diff, na.rm = TRUE),
    sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))
  ) %>%
  # Rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

```

### DB/UB Differenced Data

```{r}
#stats calculations
stats_monthly_dbub <- final_15min_dbub %>%
  mutate(temp_diff = db_temp - ub_temp) %>%
  # Removing rows with NAs
  drop_na() %>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_db_temp = mean(db_temp, na.rm = TRUE),
    mean_ub_temp = mean(ub_temp, na.rm = TRUE),
    mean_temp = mean(avg_temp, na.rm = TRUE),
    mean_diff = mean(temp_diff, na.rm = TRUE),
    sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))
  ) %>%
  # Rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### Day/Night Differenced Data

```{r}
#monthly stats calculations by day/night zone & burn status
stats_monthly_daynight <- final_daynight_singleavg_temps %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = avg_day_temp - avg_night_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### New Differenced Data from full datasets

#### NS Comparison

```{r}
#creating a monthly dataframe to carry over avg and temp_difference data into full stats dataframe - can't bind this with the stats because we need 15-minute numbers to get the correct stats
monthly_ns_data <- full_15min_ns %>%
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp, na.rm = TRUE), mean_south_temp = mean(south_temp, na.rm = TRUE), ns_avg_temp = mean(ns_avg_temp, na.rm = TRUE), temp_diff = mean(temp_diff, na.rm = TRUE), .groups = 'drop')
```

```{r}
#n/s monthly stats
stats_monthly_ns <- full_15min_ns %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), ns_mean_temp = mean(ns_avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

```{r}
#splitting the above data into gb, db and live
stats_monthly_ns <- stats_monthly_ns %>%
  mutate(burn_status = case_when(
    burn_status == "unburned" ~ "live",
    TRUE ~ burn_status  # Keeps other values as they are
  ))

stats_monthly_ns_gb <- stats_monthly_ns %>%
  filter(burn_status == "green burn")

stats_monthly_ns_db <- stats_monthly_ns %>%
  filter(burn_status == "dead burn")

stats_monthly_ns_live <- stats_monthly_ns %>%
  filter(burn_status == "live")
```

#### Charred/Uncharred/N/S GB Comparisons

```{r}
monthly_gb_data <- final_15min_gb_temps %>%
  mutate(
    datetime = as.POSIXct(datetime),
    month = format(datetime, "%m")
  ) %>%
  select(datetime, day, month, everything()) %>%
  mutate(
    ns_avg_temp = rowMeans(cbind(north_temp, south_temp)),
    c_uc_avg_temp = rowMeans(cbind(charred_temp, uncharred_temp)),
    s_c_avg_temp = rowMeans(cbind(south_temp, charred_temp)),
    s_uc_avg_temp = rowMeans(cbind(south_temp, uncharred_temp)),
    n_c_avg_temp = rowMeans(cbind(north_temp, charred_temp)),
    n_uc_avg_temp = rowMeans(cbind(north_temp, uncharred_temp)),
    ns_temp_diff = north_temp - south_temp,
    c_uc_temp_diff = charred_temp - uncharred_temp,
    s_c_temp_diff = south_temp - charred_temp,
    s_uc_temp_diff = south_temp - uncharred_temp,
    n_c_temp_diff = north_temp - charred_temp,
    n_uc_temp_diff = north_temp - uncharred_temp
  ) %>%
  group_by(month, zone, burn_status) %>%
  summarize(
    mean_north_temp = mean(north_temp, na.rm = TRUE),
    mean_south_temp = mean(south_temp, na.rm = TRUE),
    mean_charred_temp = mean(charred_temp, na.rm = TRUE),
    mean_uncharred_temp = mean(uncharred_temp, na.rm = TRUE),
    ns_avg_temp = mean(ns_avg_temp, na.rm = TRUE),
    c_uc_avg_temp = mean(c_uc_avg_temp, na.rm = TRUE),
    s_c_avg_temp = mean(s_c_avg_temp, na.rm = TRUE),
    s_uc_avg_temp = mean(s_uc_avg_temp, na.rm = TRUE),
    n_c_avg_temp = mean(n_c_avg_temp, na.rm = TRUE),
    n_uc_avg_temp = mean(n_uc_avg_temp, na.rm = TRUE),
    ns_temp_diff = mean(ns_temp_diff, na.rm = TRUE),
    c_uc_temp_diff = mean(c_uc_temp_diff, na.rm = TRUE),
    s_c_temp_diff = mean(s_c_temp_diff, na.rm = TRUE),
    s_uc_temp_diff = mean(s_uc_temp_diff, na.rm = TRUE),
    n_c_temp_diff = mean(n_c_temp_diff, na.rm = TRUE),
    n_uc_temp_diff = mean(n_uc_temp_diff, na.rm = TRUE),
    .groups = 'drop'
  )

```

Don't use the below code

```{r}
## dont use this!
#charred/uncharred monthly stas
stats_monthly_c_uc_V2 <- monthly_gb_data %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(c_uc_mean_temp = mean(c_uc_avg_temp), mean_diff = mean(c_uc_temp_diff), sd = sd(c_uc_temp_diff, na.rm = TRUE),
    coeff_var = (sd(c_uc_temp_diff, na.rm = TRUE) / abs(mean(c_uc_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate(across(where(is.numeric), round, 3))
```

#### GBDBUB Comparison

```{r}
monthly_gbdbub_data <- full_15min_gbdbub %>%
  group_by(month, zone) %>%
  summarize(mean_gb_temp = mean(gb_temp, na.rm = TRUE), mean_db_temp = mean(db_temp, na.rm = TRUE), mean_ub_temp = mean(ub_temp, na.rm = TRUE), gbdb_avg_temp = mean(gbdb_avg_temp, na.rm = TRUE), gbub_avg_temp = mean(gbub_avg_temp, na.rm = TRUE), dbub_avg_temp = mean(dbub_avg_temp, na.rm = TRUE), gbdb_temp_diff = mean(gbdb_temp_diff, na.rm = TRUE), gbub_temp_diff = mean(gbub_temp_diff, na.rm = TRUE), dbub_temp_diff = mean(dbub_temp_diff, na.rm = TRUE), .groups = 'drop')
```

```{r}
#making a stats dataframe with combined data did not work, so I am going to bind the already created gbdbub dataframes so I have a full gbdbub stats dataframe

#gb/db stats - db minus gb! - same as above no V1 version!
stats_monthly_gbdb_V2 <- final_15min_gbdb %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbdb_temp_diff = db_temp - gb_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(gbdb_mean_temp = mean(avg_temp), gbdb_mean_diff = mean(gbdb_temp_diff), gbdb_sd = sd(gbdb_temp_diff, na.rm = TRUE),
    gbdb_coeff_var = (sd(gbdb_temp_diff, na.rm = TRUE) / abs(mean(gbdb_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

#gb/ub stats - DONT USE
stats_monthly_gbub_V1 <- final_15min_gbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbub_temp_diff = gb_temp - ub_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(gbub_mean_temp = mean(avg_temp), gbub_mean_diff = mean(gbub_temp_diff), gbub_sd = sd(gbub_temp_diff, na.rm = TRUE),
    gbub_coeff_var = (sd(gbub_temp_diff, na.rm = TRUE) / abs(mean(gbub_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

#db/ub stats
## this is the same as the V2 version above!
stats_monthly_dbub_V1 <- final_15min_dbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(dbub_temp_diff = db_temp - ub_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  drop_na() %>%
  summarize(dbub_mean_temp = mean(avg_temp), dbub_mean_diff = mean(dbub_temp_diff), dbub_sd = sd(dbub_temp_diff, na.rm = TRUE),
    dbub_coeff_var = (sd(dbub_temp_diff, na.rm = TRUE) / abs(mean(dbub_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

##joining all V2 dataframes
stats_monthly_gbdbub <- stats_monthly_gbdb_V2 %>%
  full_join(stats_monthly_gbub_V2) %>%
  full_join(stats_monthly_dbub_V2)
```

#### Saving New full stats as CSVs

These are full csv tables, not seperated by burn/unburned status

```{r}
write_csv(monthly_ns_data, "nsf/treetemp_data/monthly_ns_data.csv")

write_csv(stats_monthly_ns, "nsf/treetemp_data/stats_monthly_ns.csv")

write_csv(monthly_gbdbub_data, "nsf/treetemp_data/monthly_gbdbub_data.csv")

write_csv(stats_monthly_gbdbub, "nsf/treetemp_data/monthly_gbdbub_data.csv")
```

## Saving Stats CSV

### Exporting CSVs to N-drive

```{r}
write_csv(stats_monthly_ns_b, "nsf/treetemp_data/stats_monthly_ns_b.csv")

write_csv(stats_monthly_ns_ub, "nsf/treetemp_data/stats_monthly_ns_ub.csv")

write_csv(stats_monthly_ns_gb, "nsf/treetemp_data/stats_monthly_ns_gb.csv")

write_csv(stats_monthly_ns_db, "nsf/treetemp_data/stats_monthly_ns_db.csv")

write_csv(stats_monthly_gbdb, "nsf/treetemp_data/stats_monthly_gbdb.csv")

write_csv(stats_monthly_gbub, "nsf/treetemp_data/stats_monthly_gbub.csv")

write_csv(stats_monthly_dbub, "nsf/treetemp_data/stats_monthly_dbub.csv")

write_csv(stats_monthly_daynight, "nsf/treetemp_data/stats_monthly_daynight.csv")

#new charred/uncharred/n/s comparisons
write_csv(stats_monthly_c_uc, "nsf/treetemp_data/stats_montly_c_uc.csv")

write_csv(stats_monthly_s_c, "nsf/treetemp_data/stats_montly_s_c.csv")

write_csv(stats_monthly_s_uc, "nsf/treetemp_data/stats_montly_s_uc.csv")

write_csv(stats_monthly_n_c, "nsf/treetemp_data/stats_montly_n_c.csv")

write_csv(stats_monthly_n_uc, "nsf/treetemp_data/stats_montly_n_uc.csv")

#new gb/db/live n/s comparisons
write_csv(stats_monthly_ns_gb, "nsf/treetemp_data/stats_monthly_ns_gb.csv")

write_csv(stats_monthly_ns_db, "nsf/treetemp_data/stats_monthly_ns_db.csv")

write_csv(stats_monthly_ns_live, "nsf/treetemp_data/stats_monthly_ns_live.csv")
```

```{r}
write_csv(stats_monthly_s_n, "nsf/treetemp_data/stats_monthly_s_n.csv")
```

\*\*not using the below conversion process as of 2024-07-22, just adding csvs to word doc and turning into a table

# Converting CSVs to Tables

```{r}
#creating one NS table with subtitles of burned and unburned (seperate tables)

# Create HTML table with kable and kableExtra
ns_b_statstable <- kable(stats_monthly_ns_b, "html", caption = "Monthly Statistics of North and South Aspect Tree Temp Comparison, for Burned Zone") %>%
  kable_classic(full_width = F, html_font = "Cambria")

ns_ub_statstable <- kable(stats_monthly_ns_ub, "html", caption = "unburned") %>%
  kable_classic(full_width = F, html_font = "Cambria")

# Save the HTML tables to separate files
html_file_ns_b <- "ns_b_stats_table.html"
writeLines(as.character(ns_b_statstable), html_file_ns_b)

html_file_ns_ub <- "ns_ub_stats_table.html"
writeLines(as.character(ns_ub_statstable), html_file_ns_ub)

# Optionally, open the HTML files in the default web browser
browseURL(html_file_ns_b)
browseURL(html_file_ns_ub)
```

#the above process worked, but the html table doesn't look that good. Consider trying to make one large table, with all the sub tables included with sub-headers. That would work better I think?

# Plotting Temp Differences

```{r}

```

# Process to turn csv's into tables after exported

1.  Open CSV in Text Editor

2.  Copy all text

3.  Paste into Word Document

4.  highlight all text in word doc

5.  Click Tables \> Convert \> Convert text to table (make sure comma delimination is highlighted)

6.  Should appear as a table in the word document now

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

# Understanding the Statistics

When you calculate an average difference between two temperatures and analyze the percent difference, average, and coefficient of variation, each metric provides different insights into your data:

1.  **Percent Difference**:

    -   **What it tells you**: The percent difference gives you a relative measure of the difference between the two temperatures compared to their average. It tells you how large the difference is in relation to the overall magnitude of the temperatures.

    -   **Interpretation**: A high percent difference indicates a significant discrepancy between the two temperatures, whereas a low percent difference suggests that the temperatures are relatively similar.

2.  **Average**:

    -   **What it tells you**: The average temperature difference provides a measure of the central tendency of the differences between the two temperatures. It is the mean value of all the calculated differences.

    -   **Interpretation**: This metric gives you an idea of the typical or expected difference between the temperatures. It is useful for understanding the overall trend in the temperature differences.

3.  **Coefficient of Variation (CV)**:

    -   **What it tells you**: The coefficient of variation is a standardized measure of dispersion of the temperature differences. It is calculated as the ratio of the standard deviation to the mean and is usually expressed as a percentage.

    -   **Interpretation**: A high CV indicates that the temperature differences are highly variable, suggesting inconsistency or high variability in the temperature differences. A low CV indicates that the temperature differences are relatively consistent and stable.

### Example Scenario

Suppose you have two sets of temperature readings (in °C) from two different sensors or locations. After calculating the differences between the corresponding readings, you obtain the following statistics:

-   **Average Difference**: 2°C

-   **Percent Difference**: 10%

-   **Coefficient of Variation**: 25%

#### Interpretation:

-   **Average Difference**: On average, there is a 2°C difference between the temperatures recorded by the two sensors or at the two locations.

-   **Percent Difference**: The 10% difference indicates that the temperature difference is relatively small compared to the average temperature values of the readings.

-   **Coefficient of Variation**: A CV of 25% suggests moderate variability in the temperature differences. While there is some fluctuation in the differences, it is not extremely high, indicating that the differences are somewhat consistent but not entirely uniform.

In summary, these metrics together help you understand the magnitude, relative size, and variability of the temperature differences in your data.

## My Understanding of the Data

-   very high coefficient of variation and percent differences mean that there is a lot of variability of temperatures, even on a daily and monthly scale.

\*\*Made edits to test git connection

# T-Tests

Loading Data for tree temps

```{r}
pers_b_aspects <- read_csv("nsf/105E_pers_burned/105E_pers_b_aspects.csv") %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

trans_b_aspects <- read_csv("nsf/105E_trans_burned/105E_trans_b_aspects.csv") %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

pers_ub_aspects <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_aspects.csv") %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

trans_ub_aspects <- read_csv("nsf/105E_trans_unburned/105E_trans_ub_aspects.csv") %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))
```

Filtering dataframes for the purpose of temp comparisons

```{r}
filtered_pers_b_aspects <- pers_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_pers_ub_aspects <- pers_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_trans_b_aspects <- trans_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_trans_ub_aspects <- trans_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")
```

## T-Test Dataframes

### Green burned trees

```{r}
#using earlier dataframes to compare on a hourly timestep 
##testing with the hourly gb dataset - final_hourly_gb_temps

t_test_hourly_gb_s_n_result <- t.test(final_hourly_gb_temps$mean_south_temp, final_hourly_gb_temps$mean_north_temp)

print(t_test_hourly_gb_s_n_result)
```

-   above is statistically significant!

```{r}
t_test_hourly_gb_c_uc_result <- t.test(final_hourly_gb_temps$mean_charred_temp, final_hourly_gb_temps$mean_uncharred_temp)

print(t_test_hourly_gb_c_uc_result )
```

-   above is statistically significant

```{r}
t_test_hourly_gb_s_c_result <- t.test(final_hourly_gb_temps$mean_south_temp, final_hourly_gb_temps$mean_charred_temp)

print(t_test_hourly_gb_s_c_result)
```

-   above is NOT statistically significant

```{r}
t_test_hourly_gb_s_uc_result <- t.test(final_hourly_gb_temps$mean_south_temp, final_hourly_gb_temps$mean_uncharred_temp)

print(t_test_hourly_gb_s_uc_result)
```

-   Above is NOT statistically significant!

```{r}
t_test_hourly_gb_n_c_result <- t.test(final_hourly_gb_temps$mean_north_temp, final_hourly_gb_temps$mean_charred_temp)

print(t_test_hourly_gb_n_c_result)
```

-   above is statistically significant!

```{r}
t_test_hourly_gb_n_uc_result <- t.test(final_hourly_gb_temps$mean_north_temp, final_hourly_gb_temps$mean_uncharred_temp)

print(t_test_hourly_gb_n_uc_result)
```

-   above is NOT statistically significant!

Monthly GB T-Tests - to look at monthly trends

```{r}
#creating monthly dataframes
final_hourly_gb_temps <- final_hourly_gb_temps %>%
  mutate(datetime_hour = as.POSIXct(datetime_hour, format=("%Y-%m-%d %H:%M:%S")),
                               datetime_hour = force_tz(datetime_hour, tz = 'MST'))

feb_gb_data <- final_hourly_gb_temps %>%
  filter(month == "2")

march_gb_data <- final_hourly_gb_temps %>%
  filter(month == "3")

april_gb_data <- final_hourly_gb_temps %>%
  filter(month == "4")

may_gb_data <- final_hourly_gb_temps %>%
  filter(month == "5")
```

Monthly T-Tests to compare seasonal differences

Monthly GB N-S comparison

```{r}
t_test_feb_gb_s_n_result <- t.test(feb_gb_data$mean_south_temp, feb_gb_data$mean_north_temp)

print(t_test_feb_gb_s_n_result)
## above is every so slightly significant

t_test_march_gb_s_n_result <- t.test(march_gb_data$mean_south_temp, march_gb_data$mean_north_temp)

print(t_test_march_gb_s_n_result)
##above is significant

t_test_april_gb_s_n_result <- t.test(april_gb_data$mean_south_temp, april_gb_data$mean_north_temp)

print(t_test_april_gb_s_n_result)
## above is significant

t_test_may_gb_s_n_result <- t.test(may_gb_data$mean_south_temp, may_gb_data$mean_north_temp)

print(t_test_may_gb_s_n_result)
#above is not significant!
```

Monthly GB C-UC Comparison

```{r}
t_test_feb_gb_c_uc_result <- t.test(feb_gb_data$mean_charred_temp, feb_gb_data$mean_uncharred_temp)

print(t_test_feb_gb_c_uc_result)
## above is not significant

t_test_march_gb_c_uc_result <- t.test(march_gb_data$mean_charred_temp, march_gb_data$mean_uncharred_temp)

print(t_test_march_gb_c_uc_result)
## above is not significant

t_test_april_gb_c_uc_result <- t.test(april_gb_data$mean_charred_temp, april_gb_data$mean_uncharred_temp)

print(t_test_april_gb_c_uc_result)
## above is not significant? - p value = 0.03027

t_test_may_gb_c_uc_result <- t.test(may_gb_data$mean_charred_temp, may_gb_data$mean_uncharred_temp)

print(t_test_may_gb_c_uc_result)
## above is significant!
```

Monthly GB S-C comparison

```{r}
t_test_feb_gb_s_c_result <- t.test(feb_gb_data$mean_south_temp, feb_gb_data$mean_charred_temp)

print(t_test_feb_gb_s_c_result)
## above is not significant - 0.195

t_test_march_gb_s_c_result <- t.test(march_gb_data$mean_south_temp, march_gb_data$mean_charred_temp)

print(t_test_march_gb_s_c_result)
## above is not significant - 0.2871

t_test_april_gb_s_c_result <- t.test(april_gb_data$mean_south_temp, april_gb_data$mean_charred_temp)

print(t_test_april_gb_s_c_result)
## above is not significant - 0.4538

t_test_may_gb_s_c_result <- t.test(may_gb_data$mean_south_temp, may_gb_data$mean_charred_temp)

print(t_test_may_gb_s_c_result)
## above IS Significant! p = 0.002422
```

Monthly GB S-UC Comparison

```{r}
## none of these comparisons are significant

t_test_feb_gb_s_uc_result <- t.test(feb_gb_data$mean_south_temp, feb_gb_data$mean_uncharred_temp)

print(t_test_feb_gb_s_uc_result)
## above is not significant - 0.4534

t_test_march_gb_s_uc_result <- t.test(march_gb_data$mean_south_temp, march_gb_data$mean_uncharred_temp)

print(t_test_march_gb_s_uc_result)
## above is not significant - 0.2455

t_test_april_gb_s_uc_result <- t.test(april_gb_data$mean_south_temp, april_gb_data$mean_uncharred_temp)

print(t_test_april_gb_s_uc_result)
## above is not significant - 0.1629

t_test_may_gb_s_uc_result <- t.test(may_gb_data$mean_south_temp, may_gb_data$mean_uncharred_temp)

print(t_test_may_gb_s_uc_result)
## above is not significant - 0.4025
```

Monthly GB N-C comparison

```{r}
## all of these combinations are significant differences!

t_test_feb_gb_n_c_result <- t.test(feb_gb_data$mean_north_temp, feb_gb_data$mean_charred_temp)

print(t_test_feb_gb_n_c_result)
## above is not significant? - 0.086

t_test_march_gb_n_c_result <- t.test(march_gb_data$mean_north_temp, march_gb_data$mean_charred_temp)

print(t_test_march_gb_n_c_result)
## above is significant - 0.007399

t_test_april_gb_n_c_result <- t.test(april_gb_data$mean_north_temp, april_gb_data$mean_charred_temp)

print(t_test_april_gb_n_c_result)
## above is significant - 0.001116

t_test_may_gb_n_c_result <- t.test(may_gb_data$mean_north_temp, may_gb_data$mean_charred_temp)

print(t_test_may_gb_n_c_result)
## above IS Significant! p = 0.00265
```

Monthly GB N-UC Comparisons

```{r}
## none of these comparisons are significant

t_test_feb_gb_n_uc_result <- t.test(feb_gb_data$mean_north_temp, feb_gb_data$mean_uncharred_temp)

print(t_test_feb_gb_n_uc_result)
## above is significant? - 0.02531

t_test_march_gb_n_uc_result <- t.test(march_gb_data$mean_north_temp, march_gb_data$mean_uncharred_temp)

print(t_test_march_gb_n_uc_result)
## above is significant? - 0.01287

t_test_april_gb_n_uc_result <- t.test(april_gb_data$mean_north_temp, april_gb_data$mean_uncharred_temp)

print(t_test_april_gb_n_uc_result)
## above is not significant - 0.3065

t_test_may_gb_n_uc_result <- t.test(may_gb_data$mean_north_temp, may_gb_data$mean_uncharred_temp)

print(t_test_may_gb_n_uc_result)
## above is not significant - 0.3222
```

### Dead Burned Trees

```{r}
#have to set up a new dataframe with pivoted wider values
final_hourly_db_temps <- full_hourly_gbdbub %>%
  # Pivot wider, creating separate columns for db_south_temp and db_north_temp
  select(datetime_hour, day, month, zone, aspect, db_temp) %>%
  pivot_wider(
    names_from = aspect,      # The column to pivot by (north/south aspects)
    values_from = db_temp,    # The values to populate the new columns
    names_prefix = "db_"      # Prefix to add to the new column names
  ) %>%
  rename(
    south_temp = db_south,
    north_temp = db_north
  )
```

```{r}
t_test_hourly_db_s_n_result <- t.test(final_hourly_db_temps$south_temp, final_hourly_db_temps$north_temp)

print(t_test_hourly_db_s_n_result)
```

Experimenting with facet wrapping and plotting data - DONT LIKE THIS! Not sure how to interpret it

```{r}
# Group by your variable (e.g., month) and run t-tests for each group
t_test_results <- final_hourly_db_temps %>%
  group_by(month, zone) %>%
  nest() %>%
  mutate(t_test = map(data, ~ tidy(t.test(.x$south_temp, .x$north_temp)))) %>%
  unnest(t_test)

# Plot the p-values or estimates and use facet_wrap for each group
ggplot(t_test_results, aes(x = month, y = estimate)) +
  geom_point() +
  facet_wrap(~ month) +  # Facet by the grouping variable
  labs(title = "T-test Results by Month", x = "Month", y = "Mean Difference in Temperature") +
  theme_minimal()
```

Monthly DB S-N comparisons

```{r}
#making monthly dataframes
feb_db_data <- final_hourly_db_temps %>%
  filter(month == "2")

march_db_data <- final_hourly_db_temps %>%
  filter(month == "3")

april_db_data <- final_hourly_db_temps %>%
  filter(month == "4")

may_db_data <- final_hourly_db_temps %>%
  filter(month == "5")
```

```{r}
t_test_feb_db_s_n_result <- t.test(feb_db_data$south_temp, feb_db_data$north_temp)

print(t_test_feb_db_s_n_result)
## above is significant

t_test_march_db_s_n_result <- t.test(march_db_data$south_temp, march_db_data$north_temp)

print(t_test_march_db_s_n_result)
##above is significant

t_test_april_db_s_n_result <- t.test(april_db_data$south_temp, april_db_data$north_temp)

print(t_test_april_db_s_n_result)
## above is significant

t_test_may_db_s_n_result <- t.test(may_db_data$south_temp, may_db_data$north_temp)

print(t_test_may_db_s_n_result)
#above is not significant
```

### Live Trees

```{r}
#have to set up a new dataframe with pivoted wider values
final_hourly_ub_temps <- full_hourly_gbdbub %>%
  # Pivot wider, creating separate columns for db_south_temp and db_north_temp
  select(datetime_hour, day, month, zone, aspect, ub_temp) %>%
  pivot_wider(
    names_from = aspect,      # The column to pivot by (north/south aspects)
    values_from = ub_temp,    # The values to populate the new columns
    names_prefix = "ub_"      # Prefix to add to the new column names
  ) %>%
  rename(
    south_temp = ub_south,
    north_temp = ub_north
  )
```

```{r}
t_test_hourly_ub_s_n_result <- t.test(final_hourly_ub_temps$south_temp, final_hourly_ub_temps$north_temp)

print(t_test_hourly_ub_s_n_result)
```

-   the above is significant!

Monthly live S-N comparisons

```{r}
#making monthly dataframes
feb_ub_data <- final_hourly_ub_temps %>%
  filter(month == "2")

march_ub_data <- final_hourly_ub_temps %>%
  filter(month == "3")

april_ub_data <- final_hourly_ub_temps %>%
  filter(month == "4")

may_ub_data <- final_hourly_ub_temps %>%
  filter(month == "5")
```

```{r}
t_test_feb_ub_s_n_result <- t.test(feb_ub_data$south_temp, feb_ub_data$north_temp)

print(t_test_feb_ub_s_n_result)
## above is significant

t_test_march_ub_s_n_result <- t.test(march_ub_data$south_temp, march_ub_data$north_temp)

print(t_test_march_ub_s_n_result)
##above is significant

t_test_april_ub_s_n_result <- t.test(april_ub_data$south_temp, april_ub_data$north_temp)

print(t_test_april_ub_s_n_result)
## above is significant

t_test_may_ub_s_n_result <- t.test(may_ub_data$south_temp, may_ub_data$north_temp)

print(t_test_may_ub_s_n_result)
#above is not significant
```

## Zone-specific T-Tests

### Green Burned

```{r}
#setting up dataframes
p_final_hourly_gb_temps <- final_hourly_gb_temps %>%
  filter (zone == "persistent")

t_final_hourly_gb_temps <- final_hourly_gb_temps %>%
  filter (zone == "transitional")

p_t_test_hourly_gb_s_n_result <- t.test(p_final_hourly_gb_temps$mean_south_temp, p_final_hourly_gb_temps$mean_north_temp)

print(p_t_test_hourly_gb_s_n_result)

t_t_test_hourly_gb_s_n_result <- t.test(t_final_hourly_gb_temps$mean_south_temp, t_final_hourly_gb_temps$mean_north_temp)

print(t_t_test_hourly_gb_s_n_result)
```

Monthly GB S-N Comparison - by zone

```{r}
p_feb_gb_data <- feb_gb_data %>%
  filter(zone == "persistent")

t_feb_gb_data <- feb_gb_data %>%
  filter(zone == "transitional")

p_march_gb_data <- feb_gb_data %>%
  filter(zone == "persistent")

t_march_gb_data <- feb_gb_data %>%
  filter(zone == "transitional")

p_april_gb_data <- feb_gb_data %>%
  filter(zone == "persistent")

t_april_gb_data <- feb_gb_data %>%
  filter(zone == "transitional")

p_may_gb_data <- feb_gb_data %>%
  filter(zone == "persistent")

t_may_gb_data <- feb_gb_data %>%
  filter(zone == "transitional")
```

### Dead Burned

```{r}
p_final_hourly_db_temps <- final_hourly_db_temps %>%
  filter (zone == "persistent")

t_final_hourly_db_temps <- final_hourly_db_temps %>%
  filter (zone == "transitional")

p_t_test_hourly_db_s_n_result <- t.test(p_final_hourly_db_temps$south_temp, p_final_hourly_db_temps$north_temp)

print(p_t_test_hourly_db_s_n_result)

t_t_test_hourly_db_s_n_result <- t.test(t_final_hourly_db_temps$south_temp, t_final_hourly_db_temps$north_temp)

print(t_t_test_hourly_db_s_n_result)
```

Monthly DB S-N Comparisons by zone

```{r}
#dataframes
p_feb_db_data <- feb_db_data %>%
  filter(zone == "persistent")

t_feb_db_data <- feb_db_data %>%
  filter(zone == "transitional")

p_march_db_data <- march_db_data %>%
  filter(zone == "persistent")

t_march_db_data <- march_db_data %>%
  filter(zone == "transitional")

p_april_db_data <- april_db_data %>%
  filter(zone == "persistent")

t_april_db_data <- april_db_data %>%
  filter(zone == "transitional")

p_may_db_data <- may_db_data %>%
  filter(zone == "persistent")

t_may_db_data <- may_db_data %>%
  filter(zone == "transitional")
```

Persistent

```{r}
p_t_test_feb_db_s_n_result <- t.test(p_feb_db_data$south_temp, p_feb_db_data$north_temp)

print(p_t_test_feb_db_s_n_result)
## above is significant

p_t_test_march_db_s_n_result <- t.test(p_march_db_data$south_temp, p_march_db_data$north_temp)

print(p_t_test_march_db_s_n_result)
##above is significant

p_t_test_april_db_s_n_result <- t.test(p_april_db_data$south_temp, p_april_db_data$north_temp)

print(p_t_test_april_db_s_n_result)
## above is significant

p_t_test_may_db_s_n_result <- t.test(p_may_db_data$south_temp, p_may_db_data$north_temp)

print(p_t_test_may_db_s_n_result)
#above is not significant
```

Transitional

```{r}
t_t_test_feb_db_s_n_result <- t.test(t_feb_db_data$south_temp, t_feb_db_data$north_temp)

print(t_t_test_feb_db_s_n_result)
## above is significant

t_t_test_march_db_s_n_result <- t.test(t_march_db_data$south_temp, t_march_db_data$north_temp)

print(t_t_test_march_db_s_n_result)
##above is significant

t_t_test_april_db_s_n_result <- t.test(t_april_db_data$south_temp, t_april_db_data$north_temp)

print(t_t_test_april_db_s_n_result)
## above is significant

t_t_test_may_db_s_n_result <- t.test(t_may_db_data$south_temp, t_may_db_data$north_temp)

print(t_t_test_may_db_s_n_result)
#above is not significant
```

### Live

```{r}
p_final_hourly_ub_temps <- final_hourly_ub_temps %>%
  filter (zone == "persistent")

t_final_hourly_ub_temps <- final_hourly_ub_temps %>%
  filter (zone == "transitional")

p_t_test_hourly_ub_s_n_result <- t.test(p_final_hourly_ub_temps$south_temp, p_final_hourly_ub_temps$north_temp)

print(p_t_test_hourly_ub_s_n_result)

t_t_test_hourly_ub_s_n_result <- t.test(t_final_hourly_ub_temps$south_temp, t_final_hourly_ub_temps$north_temp)

print(t_t_test_hourly_ub_s_n_result)
```

Monthly Live S-N Comparisons by zone

```{r}
#dataframes
p_feb_ub_data <- feb_ub_data %>%
  filter(zone == "persistent")

t_feb_ub_data <- feb_ub_data %>%
  filter(zone == "transitional")

p_march_ub_data <- march_ub_data %>%
  filter(zone == "persistent")

t_march_ub_data <- march_ub_data %>%
  filter(zone == "transitional")

p_april_ub_data <- april_ub_data %>%
  filter(zone == "persistent")

t_april_ub_data <- april_ub_data %>%
  filter(zone == "transitional")

p_may_ub_data <- may_ub_data %>%
  filter(zone == "persistent")

t_may_ub_data <- may_ub_data %>%
  filter(zone == "transitional")
```

Persistent

```{r}
p_t_test_feb_ub_s_n_result <- t.test(p_feb_ub_data$south_temp, p_feb_ub_data$north_temp)

print(p_t_test_feb_ub_s_n_result)
## above is not significant

p_t_test_march_ub_s_n_result <- t.test(p_march_ub_data$south_temp, p_march_ub_data$north_temp)

print(p_t_test_march_ub_s_n_result)
##above is not significant

p_t_test_april_ub_s_n_result <- t.test(p_april_ub_data$south_temp, p_april_ub_data$north_temp)

print(p_t_test_april_ub_s_n_result)
## above is not significant

p_t_test_may_ub_s_n_result <- t.test(p_may_ub_data$south_temp, p_may_ub_data$north_temp)

print(p_t_test_may_ub_s_n_result)
#above is not significant
```

transitional

```{r}
t_t_test_feb_ub_s_n_result <- t.test(t_feb_ub_data$south_temp, t_feb_ub_data$north_temp)

print(t_t_test_feb_ub_s_n_result)
## above is significant

t_t_test_march_ub_s_n_result <- t.test(t_march_ub_data$south_temp, t_march_ub_data$north_temp)

print(t_t_test_march_ub_s_n_result)
##above is significant

t_t_test_april_ub_s_n_result <- t.test(t_april_ub_data$south_temp, t_april_ub_data$north_temp)

print(p_t_test_april_ub_s_n_result)
## above is significant

t_t_test_may_ub_s_n_result <- t.test(t_may_ub_data$south_temp, t_may_ub_data$north_temp)

print(t_t_test_may_ub_s_n_result)
#above is not significant
```
