---
title: "treetemp_datawrangling"
author: "Erika Lee"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#to clear prior environment
rm(list = ls())
```

## Calling in data

#### Tree temp data

```{r}
#calling in tree temp data from n-drive
## burned
pers_b_aspects <- read_csv("nsf/105E_pers_burned/105E_pers_b_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_b_aspects <- read_csv("nsf/105E_trans_burned/105E_trans_b_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

pers_ub_aspects <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_ub_aspects <- read_csv("nsf/105E_trans_unburned/105E_trans_ub_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = with_tz(datetime, tz = 'MST'))
```

#### Weather station data

NOTE - I need to rerun the wx_datawrangling Rmarkdown before calling in these dataframes if I change the range of dates I am looking for!

```{r}
transb_wx_15min_r <- read_csv('nsf/trans_burned/trans_burned_wx_15min_r.csv') %>%
  #have to force the MST otherwise it recognizes it as UTC
  mutate(datetime = force_tz(datetime, tz = 'MST'))

transub_wx_15min_r <- read_csv('nsf/trans_unburned/trans_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persb_wx_hourly_r <- read_csv('nsf/pers_burned/pers_burned_wx_hourly_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persub_wx_15min_r <- read_csv('nsf/pers_unburned/pers_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))
```

## Setting consistent timeframe

```{r}
#setting a consistent timeframe for all tree temp data for all burned/unburned zones
##has to start at 2024-02-02 due to errors in data at transitional burned site

filtered_pers_b_aspects <- pers_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_pers_ub_aspects <- pers_ub_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_b_aspects <- trans_b_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_ub_aspects <- trans_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

```{r}
#setting consistent timeframe for weather data
#filtered weather dataframes
filtered_persb_wx_hourly <- persb_wx_hourly_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_persub_wx_15min <- persub_wx_15min_r %>%
  select(datetime, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
  
filtered_transb_wx_15min <- transb_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_transub_wx_15min <- transub_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

## Solar Exposure Times

```{r}
# Specify PSZ and TSZ locations
##pers_b - from 105E location!
p_b_lat <- 40.56379284
p_b_lon <- -105.867215
##pers_ub - from 105E locations
p_ub_lat <- 40.5638075
p_ub_lon <- -105.8705715

##trans_b - from 105E locations!
t_b_lat <- 40.60600629
t_b_lon <- -105.8333335

##trans_ub - from 105E locations!
t_ub_lat <- 40.568396
t_ub_lon <- -105.587383

## use the p_b_lat/long since all of these dates/times are almost identical for sunrise/sunset

# Create a date sequence from February 2nd to May 24th, 2024 for solar exposure
solex_dates <- seq(as.Date("2024-02-02"), as.Date("2024-07-01"), by = "day")

# Get solar times for each day
p_b_solar_times <- getSunlightTimes(date = solex_dates, lat = p_b_lat, lon = p_b_lon, tz = "MST") %>%
  mutate(
    day = as.Date(date),  # Ensure 'day' is of Date type for joining
    sunrise = with_tz(as.POSIXct(sunrise), tzone = "MST"),  # Set MST timezone
    sunset = with_tz(as.POSIXct(sunset), tzone = "MST")     # Set MST timezone
  )
```

```{r}
write_csv(p_b_solar_times, "nsf/treetemp_data/p_b_solar_times.csv")
```

# Data Wrangling for Comparison

```{r}
#adding in a burn status column to the aspect dataframes
filtered_pers_b_aspects <- filtered_pers_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_b_aspects <- filtered_trans_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

## NS Pivoted Dataframes

PSZ

```{r}
#creating a dataframe for each tree
persb_gb_e_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_e') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_e')

persb_gb_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

persb_db_n_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

persb_db_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
persb_gbdb_combined_data <- bind_rows(persb_gb_e_diff_data, persb_gb_s_diff_data, persb_db_n_diff_data, persb_db_s_diff_data)
```

TSZ

```{r}
#creating a dataframe for each tree
transb_gb_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_n')

transb_gb_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

transb_db_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

transb_db_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
transb_gbdb_combined_data <- bind_rows(transb_gb_n_diff_data, transb_gb_s_diff_data, transb_db_n_diff_data, transb_db_s_diff_data)
```

Combined NS Dataframe

```{r}
#creating a full dataframe for both pers and transitional data for later plotting of the south-north comparison

##individually editing pers and trans combined data before combining both dataframes
persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "persistent")

transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "transitional")

#creating a full combined dataset
full_ns_comparison_data <- bind_rows(persb_gbdb_combined_data, transb_gbdb_combined_data)
```

```{r}
#exporting full ns comparison dataframe
write_csv(full_ns_comparison_data, "nsf/treetemp_data/full_ns_comparison_data.csv")
```

```{r}
# important step, I use these later to create the full_ns_dataframe
#creating a combined new dataframe from original dataframes with individual tree temps - May not want a full combined dataset!
full_burned_combined_data <- filtered_pers_b_aspects %>%
  full_join(filtered_trans_b_aspects)

full_unburned_combined_data <- filtered_pers_ub_aspects %>%
  full_join(filtered_trans_ub_aspects)
```

Combined Full, Filtered Aspect/temp dataframes

```{r}
#exporting important dataframes to N-Drive. These are what are used for stats calculations
write_csv(full_burned_combined_data, "nsf/treetemp_data/full_burned_combined_data.csv")

write_csv(full_unburned_combined_data, "nsf/treetemp_data/full_unburned_combined_data.csv")
```

## N/S Dataframes

GB Dataframe wrangling to remove charred/uncharred temps

-   I don't think I use these, but keeping just in case

```{r}
#green burned dataframe
full_gb_data <- full_burned_combined_data %>%
  filter(burn_status == "green burn") %>%
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_gb_temps <- full_gb_data%>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south, charred_temp = charred, uncharred_temp = uncharred) %>%
  summarize(north_temp, south_temp, charred_temp, uncharred_temp)

final_hourly_gb_temps <-final_gb_temps %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%  # Create an hourly datetime column
  group_by(datetime_hour, day, month, zone, snow_phase, burn_status) %>%  # Group by the hourly datetime
  #mean_hourly_treetemp is the mean of all the db/gb/live trees in the dataframe averaged by burn_status on an hourly timestep
  summarize(north_temp = mean(north_temp), south_temp = mean(south_temp), charred_temp = mean(charred_temp), uncharred_temp = mean(uncharred_temp), ns_avg_temp = ((north_temp + south_temp)/2), cuc_avg_temp = ((charred_temp + uncharred_temp)/2), sc_avg_temp = ((south_temp + charred_temp)/2), suc_avg_temp = ((south_temp + uncharred_temp)/2), nc_avg_temp = ((north_temp + charred_temp)/2), nuc_avg_temp = ((north_temp + uncharred_temp)/2), ns_temp_diff = (south_temp - north_temp), cuc_temp_diff = (charred_temp - uncharred_temp), sc_temp_diff = (south_temp - charred_temp), suc_temp_diff = (south_temp - uncharred_temp), nc_temp_diff = (north_temp - charred_temp), nuc_temp_diff = (north_temp - uncharred_temp), .groups = 'drop')
```

```{r}
# I don't think I use these, but keeping them anyways
##hourly
write_csv(final_hourly_gb_temps, "nsf/treetemp_data/final_hourly_gb_temps.csv")

##15-minute
write_csv(final_gb_temps, "nsf/treetemp_data/final_15min_gb_temps.csv")
```

### PSZ 

Burned

Creating a clean dataframe with columns for north and south temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree, so there is two gb and two db
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_persb_ns_temps <- full_pers_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

Unburned

```{r}
#persistent unburned final dataset
full_pers_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_pers_ub_ns_temps <- full_pers_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_pers_ns_data <- final_persb_ns_temps %>%
  full_join(final_pers_ub_ns_temps)
```

### TSZ

Burned

```{r}
#transitional burned 
full_trans_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_transb_ns_temps <- full_trans_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

Unburned

```{r}
#tranitional unburned final dataset
full_trans_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_trans_ub_ns_temps <- full_trans_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_trans_ns_data <- final_transb_ns_temps %>%
  full_join(final_trans_ub_ns_temps)
```

Final Full N/S dataframe

```{r}
full_15min_ns <- full_pers_ns_data %>%
  full_join(full_trans_ns_data) %>%
#adding in mean temp of n/s temps
  mutate(ns_avg_temp = ((north_temp + south_temp)/2), ns_temp_diff = (south_temp - north_temp))
```

## DB/GB/UB Dataframes

### PSZ

Burned

Creating a clean dataframe with columns for gb, db, and ub temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_gbdb_temps <- full_pers_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

Unburned

```{r}
#persistent unburned final dataset
full_pers_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_ub_temps <- full_pers_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(datetime, day, month, aspect, zone, snow_phase, ub_temp) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

### TSZ 

Burned

```{r}
#transitional burned 
full_trans_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_gbdb_temps <- full_trans_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp)
```

Unburned

```{r}
#transitional unburned final dataset
full_trans_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_ub_temps <- full_trans_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(-burn_status) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

Combining Burned/Unburned dataframes

```{r}
final_pers_gbdb_temps <- final_pers_gbdb_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))

final_pers_ub_temps <- final_pers_ub_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))
```

```{r}
full_pers_gbdbub_data <- final_pers_gbdb_temps %>%
  left_join(final_pers_ub_temps)
```

```{r}
full_trans_gbdbub_data <- final_trans_gbdb_temps %>%
  left_join(final_trans_ub_temps)
```

```{r}
full_15min_gbdbub <- rbind(full_pers_gbdbub_data, full_trans_gbdbub_data)
```

```{r}
write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

# Final 15-minute Dataframes

Combining burned and unburned dataframes by zone and 15-minute timestep

## GB/DB/UB Dataframes

```{r}
# GB/DB comparison
final_15min_gbdb <- final_pers_gbdb_temps %>%
  full_join(final_trans_gbdb_temps) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((gb_temp + db_temp) / 2)) %>%
  select(datetime, day, month, zone, snow_phase, everything())

#adding in an hourly version of this final_15min_gbdb dataframe
final_hourly_gbdb <- final_15min_gbdb %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, aspect) %>%
  summarize(gb_temp = mean(gb_temp, na.rm = TRUE), db_temp = mean(db_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#creating dataframe for combined ub temps
final_15min_gbdb_ub <- final_pers_ub_temps %>%
  full_join(final_trans_ub_temps) %>%
  #adding in day & month
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  select(datetime, day, month, zone, snow_phase, everything()) 

#adding in an hourly version of this final_15min_ub dataframe
final_hourly_gbdb_ub <- final_15min_gbdb_ub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, aspect) %>%
  summarize(ub_temp = mean(ub_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#creating a combined gbdbub full dataframe
## this is an important step in data wrangling - need this for final csvs
full_15min_gbdbub <- final_15min_gbdb %>%
  select(-avg_temp) %>%
  full_join(final_15min_gbdb_ub) %>%
  #remove weird NAs
  drop_na()
```

```{r}
# Adding in temp comparison column by burn condition!
full_15min_gbdbub <- full_15min_gbdbub %>%
  #adding in mean temp of gbdb/gbub/dbub temps
  mutate(gbdb_avg_temp = ((gb_temp + db_temp) / 2)) %>%
  mutate(gbub_avg_temp = ((gb_temp + ub_temp) / 2)) %>%
  mutate(dbub_avg_temp = ((db_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything()) %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbdb_temp_diff = gb_temp - db_temp) %>%
  mutate(gbub_temp_diff = gb_temp - ub_temp) %>%
  mutate(dbub_temp_diff = db_temp - ub_temp)
```

```{r}
write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

```{r}
write_csv(full_15min_ns, "nsf/treetemp_data/full_15min_ns.csv")

write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

# Final Data frames

## Adding in Day/Night Phase

```{r}
filtered_pers_b_aspects_r <- filtered_pers_b_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_pers_ub_aspects_r <- filtered_pers_ub_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_trans_b_aspects_r <- filtered_trans_b_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_trans_ub_aspects_r <- filtered_trans_ub_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
full_15min_ns <- full_15min_ns %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = with_tz(datetime, tz = 'MST')) %>%
  mutate(day = as.Date(day, format = ("%Y-%m-%d"))) %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, north_temp, south_temp, solarNoon, sunrise, sunset, dawn, dusk)

full_15min_gbdbub <- full_15min_gbdbub %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = with_tz(datetime, tz = 'MST'), ,
    day = as.Date(datetime, tz= "MST")) %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, aspect, gb_temp, db_temp, ub_temp, solarNoon, sunrise, sunset, dawn, dusk)
```

##  Exporting

```{r}
#exporting above dataframes
write.csv(filtered_pers_b_aspects_r, "nsf/treetemp_data/filtered_pers_b_aspects_r.csv")

write.csv(filtered_pers_ub_aspects_r, "nsf/treetemp_data/filtered_pers_ub_aspects_r.csv")

write.csv(filtered_trans_b_aspects_r, "nsf/treetemp_data/filtered_trans_b_aspects_r.csv")

write.csv(filtered_trans_ub_aspects_r, "nsf/treetemp_data/filtered_trans_ub_aspects_r.csv")
```

```{r}
write_csv(full_15min_ns, "nsf/treetemp_data/full_15min_ns.csv")

write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```
