---
title: "treetemp_datawrangling"
author: "Erika Lee"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

### Calling in data

Tree temp data

```{r}
#calling in tree temp data from n-drive
## burned
pers_b_aspects <- read_excel("nsf/105E_pers_burned/105E_pers_b_aspects.xlsx") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_b_aspects <- read_excel("nsf/105E_trans_burned/105E_trans_b_aspects.xlsx") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

##unburned
pers_ub_aspects <- read_excel("nsf/105E_pers_unburned/105E_pers_ub_aspects.xlsx") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_ub_aspects <- read_excel("nsf/105E_trans_unburned/105E_trans_ub_aspects.xlsx") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))
```

Weather station data:

NOTE - I need to rerun the wx_datawrangling Rmarkdown before calling in these dataframes if I change the range of dates I am looking for!

```{r}
transb_wx_15min_r <- read_excel('nsf/trans_burned/trans_burned_wx_15min_r.xlsx') %>%
  #have to force the MST otherwise it recognizes it as UTC
  mutate(datetime = force_tz(datetime, tz = 'MST'))

transub_wx_15min_r <- read_excel('nsf/trans_unburned/trans_unburned_wx_15min_r.xlsx') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persb_wx_15min_r <- read_excel('nsf/pers_burned/pers_burned_wx_15min_r.xlsx') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persub_wx_15min_r <- read_excel('nsf/pers_unburned/pers_unburned_wx_15min_r.xlsx') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))
```

### Setting consistent timeframe

```{r}
#setting a consistent timeframe for all tree temp data for all burned/unburned zones
##has to start at 2024-02-02 due to errors in data at transitional burned site

#tree temp dataframes
start_time <- ymd_hms("2024-02-02 00:00:00")
end_time <- ymd_hms("2024-05-24 00:00:00")

filtered_pers_b_aspects <- pers_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_pers_ub_aspects <- pers_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_trans_b_aspects <- trans_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")

filtered_trans_ub_aspects <- trans_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S"))) %>%
  filter(datetime >= "2024-02-02 00:00:00" & datetime <= "2024-05-24 00:00:00")
```

```{r}
#setting consistent timeframe for weather data
#filtered weather dataframes
filtered_persb_wx_15min <- persb_wx_15min %>%
  filter(datetime >= start_time & datetime <= end_time) %>%
  #removed UTM column
  select(-TIMESTAMP)

filtered_persub_wx_15min <- persub_wx_15min %>%
  select(datetime, everything()) %>%
  filter(datetime >= start_time & datetime <= end_time)
  
filtered_transb_wx_15min <- transb_wx_15min %>%
  filter(datetime >= start_time & datetime <= end_time)

filtered_transun_wx_15min <- transub_wx_15min %>%
  filter(datetime >= start_time & datetime <= end_time)
```

# Tree Temp Comparison

```{r}
#adding in a burn status column to the aspect dataframes
filtered_pers_b_aspects <- filtered_pers_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "persistent") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m")))

filtered_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "persistent") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m")))

filtered_trans_b_aspects <- filtered_trans_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "transitional") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m")))

filtered_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  mutate(zone = "transitional") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m")))
```

## Differencing Dead burn and Green burn trees - For Plotting later!

### Persistent Zone

```{r}
#creating a dataframe with green burn temp minus dead burn temp for the same aspects
##north aspect
persb_north_diff_data <- filtered_pers_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'north') %>%
  # Filter out gb_w tree
  filter(tree_name != 'gb_w') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_e_minus_db_s = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_e_minus_db_n = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
persb_north_diff_long <- persb_north_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "persistent") %>%
  mutate(aspect = "north")
```

North Aspect Boxplots

```{r}
# Plotting the boxplots
persb_north_diff_plot <- ggplot(persb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(persb_north_diff_plot)
```

```{r}
##south aspect
persb_south_diff_data <- filtered_pers_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'south') %>%
  # Filter out gb_w tree
  filter(tree_name != 'gb_w') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_e_minus_db_s = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_e_minus_db_n = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
persb_south_diff_long <- persb_south_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "persistent") %>%
  mutate(aspect = "south")
```

South Aspect Boxplots

```{r}
# Plotting the boxplots
persb_south_diff_plot <- ggplot(persb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(persb_south_diff_plot)
```

Combining Plots

```{r}
#combinging north and south plots
persb_combined_aspectdiff_treetemp_plots <- persb_north_diff_plot + persb_south_diff_plot + plot_layout(nrow = 2) + plot_annotation(title = "Persistent Zone, Green & Dead Burn Tree Temp Difference")

persb_combined_aspectdiff_treetemp_plots 

#saving combined plot
ggsave("persb_combined_aspectdiff_treetemp_plots.png", plot = persb_combined_aspectdiff_treetemp_plots , width = 8, height = 10, dpi = 300, bg = "white")
```

### Transitional Zone

```{r}
#creating a dataframe with green burn temp minus dead burn temp for the same aspects
##north aspect
transb_north_diff_data <- filtered_trans_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'north') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_n_minus_db_s = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_n_minus_db_n = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
transb_north_diff_long <- transb_north_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "transitional") %>%
  mutate(aspect = "north")
```

North Aspect Boxplots

```{r}
# Plotting the boxplots
transb_north_diff_plot <- ggplot(transb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(transb_north_diff_plot)
```

```{r}
##south aspect
transb_south_diff_data <- filtered_trans_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'south') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_n_minus_db_s = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_n_minus_db_n = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
transb_south_diff_long <- transb_south_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "transitional") %>%
  mutate(aspect = "south")
```

South Aspect Boxplots

```{r}
# Plotting the boxplots
transb_south_diff_plot <- ggplot(transb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(transb_south_diff_plot)
```

Combining Plots

```{r}
#combinging north and south plots
transb_combined_aspectdiff_treetemp_plots <- transb_north_diff_plot + transb_south_diff_plot + plot_layout(nrow = 2) + plot_annotation(title = "Transitional Zone, Green & Dead Burn Tree Temp Difference")

transb_combined_aspectdiff_treetemp_plots 

#saving combined plot
ggsave("transb_combined_aspectdiff_treetemp_plots.png", plot = transb_combined_aspectdiff_treetemp_plots , width = 8, height = 10, dpi = 300, bg = "white")
```

Combining Dataframes

```{r}
#creating a combined dataframe for the db-gb comparisons
full_dbgb_comparisons_data <- bind_rows(persb_north_diff_long, persb_south_diff_long, transb_north_diff_long, transb_south_diff_long)
```

Saving Combined dataframe

```{r}
write_csv(full_dbgb_comparisons_data, "nsf/treetemp_data/full_dbgb_comparisons_data.csv")
```

## Differencing by Individual Tree Aspect

### Persistent Zone

```{r}
#creating a dataframe for each tree
persb_gb_e_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_e') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_e')

persb_gb_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

persb_db_n_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

persb_db_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
persb_gbdb_combined_data <- bind_rows(persb_gb_e_diff_data, persb_gb_s_diff_data, persb_db_n_diff_data, persb_db_s_diff_data)
```

Plotting

```{r}
#plotting persistent north-south tree boxplots
persb_gbdb_diff_plot <- ggplot(persb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Persistent Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(persb_gbdb_diff_plot)
```

### Transitional Zone

```{r}
#creating a dataframe for each tree
transb_gb_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_n')

transb_gb_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

transb_db_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

transb_db_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
transb_gbdb_combined_data <- bind_rows(transb_gb_n_diff_data, transb_gb_s_diff_data, transb_db_n_diff_data, transb_db_s_diff_data)
```

Plotting

```{r}
#plotting transitional north-south tree boxplots
transb_gbdb_diff_plot <- ggplot(transb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Transitional Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(transb_gbdb_diff_plot)
```

Combined Plots

```{r}
#combinging north and south plots
combined_bytree_aspect_treetemp_plots <- persb_gbdb_diff_plot + transb_gbdb_diff_plot + plot_layout(nrow = 1) + plot_annotation(title = "Temperature Differences by Tree - south minus north aspect")

combined_bytree_aspect_treetemp_plots

#saving combined plot
ggsave("combined_bytree_aspect_treetemp_plots.png", plot = combined_bytree_aspect_treetemp_plots , width = 10, height = 6, dpi = 300, bg = "white")
```

Combined NS Dataframe

```{r}
#creating a full dataframe for both pers and transitional data for later plotting of the south-north comparison

##individually editing pers and trans combined data before combining both dataframes
# Add new column 'burn_status' based on conditions, and new column 'zone' based on location
persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "persistent")

transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "transitional")

#creating a full combined dataset
full_ns_comparison_data <- bind_rows(persb_gbdb_combined_data, transb_gbdb_combined_data)
```

```{r}
#exporting full ns comparison dataframe
write_csv(full_ns_comparison_data, "nsf/treetemp_data/full_ns_comparison_data.csv")
```

**Interpretation of box plots**

-   Consistent Boxes: Similar boxes indicate that the general trend and variability within the middle 50% of the data are consistent across different aspects.

-   Many Outliers: The numerous outliers highlight that there are frequent instances of temperature differences that are much higher or lower than the typical range. This could point to specific factors or conditions that cause these deviations.

# Pivoting to long dataframes for temp comparisons

```{r}
#creating a combined new dataframe from original dataframes with individual tree temps - May not want a full combined dataset!
full_burned_combined_data <- bind_rows(filtered_pers_b_aspects, filtered_trans_b_aspects)

full_unburned_combined_data <- bind_rows(filtered_pers_ub_aspects, filtered_trans_ub_aspects)
```

Combined Full, Filtered Aspect/temp dataframes

```{r}
#exporting important dataframes to N-Drive. These are what are used for stats calculations
write_csv(full_burned_combined_data, "nsf/treetemp_data/full_burned_combined_data.csv")

write_csv(full_unburned_combined_data, "nsf/treetemp_data/full_unburned_combined_data.csv")
```

## N/S Aspect Dataframes

\*\* need to create a dataframe for only the gb trees and do the following comparisons, because I have to subtract it form the north/south aspects dataframes because deadburned trees do not have charred/uncharred aspect temps

\*\* Use the following final_hourly_GB temps dataset to do plotting of charred/uncharred comparisons, as well as for stats calculations!

```{r}
#green burned dataframe
full_gb_data <- full_burned_combined_data %>%
  filter(burn_status == "green burn") %>%
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_gb_temps <- full_gb_data%>%
  group_by(datetime, day, month, burn_status, zone) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south, charred_temp = charred, uncharred_temp = uncharred) %>%
  summarize(north_temp, south_temp, charred_temp, uncharred_temp)

final_hourly_gb_temps <-final_gb_temps %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%  # Create an hourly datetime column
  group_by(datetime_hour, day, month, zone, burn_status) %>%  # Group by the hourly datetime
  #mean_hourly_treetemp is the mean of all the db/gb/live trees in the dataframe averaged by burn_status on an hourly timestep
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), mean_charred_temp = mean(charred_temp), mean_uncharred_temp = mean(uncharred_temp), .groups = 'drop')
```

```{r}
#writing this gb csv to use in stats calculations!

##hourly
write_csv(final_hourly_gb_temps, "nsf/treetemp_data/final_hourly_gb_temps.csv")
##15-minute
write_csv(final_gb_temps, "nsf/treetemp_data/final_15min_gb_temps.csv")
```

### Pers_b

Creating a clean dataframe with columns for north and south temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "persistent") %>%
  #remove gb_w tree, so there is two gb and two db
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_persb_ns_temps <- full_pers_ns_data %>%
  group_by(datetime, day, month, burn_status, zone) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

### Pers_ub

```{r}
#persistent unburned final dataset
full_pers_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "unburned") %>%
  filter(zone == "persistent") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_pers_ub_ns_temps <- full_pers_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
full_pers_ns_data <- final_persb_ns_temps %>%
  full_join(final_pers_ub_ns_temps)
```

### Trans_b

```{r}
#transitional burned 
full_trans_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "transitional") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_transb_ns_temps <- full_trans_ns_data %>%
  group_by(datetime, day, month, burn_status, zone) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

### Trans_ub

```{r}
#tranitional unburned final dataset
full_trans_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "unburned") %>%
  filter(zone == "transitional") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_trans_ub_ns_temps <- full_trans_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
full_trans_ns_data <- final_transb_ns_temps %>%
  full_join(final_trans_ub_ns_temps)
```

## DB/GB/UB Dataframes \*\*\*

### Pers_b

Creating a clean dataframe with columns for gb, db, and ub temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "persistent") %>%
  #remove gb_w tree, so there is two gb and two db
  filter(!(tree_name == "gb_w"))

clean_persb_gbdb_temps <- full_pers_gbdb_data %>%
  group_by(datetime, burn_status, zone, aspect) %>%
  pivot_wider(names_from = burn_status, values_from = temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  # Selecting only rows where tree_name is "gb_s" or "db_s"
  filter(tree_name %in% c("gb_s", "db_s")) %>%
  select(c(-tree_name, -sensor_id))


##creatinging individual dataframes for north and south temps, then going to rebind them to get rid of the NA issues
gb_persb_temps <- clean_persb_gbdb_temps %>%
  select(-db_temp) %>%
  # Filter out rows with NA
  filter(!is.na(gb_temp))

db_persb_temps <- clean_persb_gbdb_temps %>%
  select(-gb_temp) %>%
  filter(!is.na(db_temp))

##joining north/south dataframes
final_persb_gbdb_temps <- gb_persb_temps %>%
  full_join(db_persb_temps, by = c("datetime", "aspect", "zone"))
```

### Trans_b

```{r}
#transitional burned 

full_trans_gbdb_data <- full_burned_combined_data %>%
  #there is a missing datetime for db_north, so I have to filter this dataset specifically to exclude the first datetime reading
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "transitional")

clean_transb_gbdb_temps <- full_trans_gbdb_data %>%
  group_by(datetime, burn_status, zone, aspect) %>%
  pivot_wider(names_from = burn_status, values_from = temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  # Selecting only rows where tree_name is "gb_s" or "db_s"
  filter(tree_name %in% c("gb_s", "db_s")) %>%
  select(c(-tree_name, -sensor_id))


##creatinging individual dataframes for north and south temps, then going to rebind them to get rid of the NA issues
gb_transb_temps <- clean_transb_gbdb_temps %>%
  select(-db_temp) %>%
  # Filter out rows with NA
  filter(!is.na(gb_temp))

db_transb_temps <- clean_transb_gbdb_temps %>%
  select(-gb_temp) %>%
  filter(!is.na(db_temp))

##joining north/south dataframes
final_transb_gbdb_temps <- gb_transb_temps %>%
  full_join(db_transb_temps, by = c("datetime", "aspect", "zone"))
```

### Pers_ub

```{r}
full_pers_ub_data <- full_unburned_combined_data %>%
  filter(zone == "persistent")

clean_pers_ub_temps <- full_pers_ub_data %>%
  group_by(datetime, burn_status, zone, aspect) %>%
  pivot_wider(names_from = burn_status, values_from = temp) %>%
  rename(ub_temp = "alive") %>%
  #selecting one tree, "ub_s"
  filter(tree_name == "ub_s") %>%
  select(c(-tree_name, -sensor_id))

# final pers_ub temps dataframe
final_pers_ub_temps <- clean_pers_ub_temps
```

### Trans_ub

```{r}
full_trans_ub_data <- full_unburned_combined_data %>%
  filter(zone == "transitional")

clean_trans_ub_temps <- full_trans_ub_data %>%
  group_by(datetime, burn_status, zone, aspect) %>%
  pivot_wider(names_from = burn_status, values_from = temp) %>%
  rename(ub_temp = "alive") %>%
  #selecting one tree, "ub_e"
  filter(tree_name == "ub_e") %>%
  select(c(-tree_name, -sensor_id))

# final pers_ub temps dataframe
final_trans_ub_temps <- clean_trans_ub_temps
```

# Final 15-minute Dataframes

Combining burned and unburned dataframes by zone and 15-minute timestep - this is what is used for stats calculations

-   important step
-   Also adding avg_temp for 15-minute timestep

## North/South Aspects

```{r}
#combining persistent and transitional unburned to create a 15min_unburned datatable

final_15min_ns_ub <- bind_rows(final_pers_ub_ns_temps, final_trans_ub_ns_temps) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((south_temp + north_temp) / 2)) %>%
  mutate(burn_status = "unburned") %>%
  select(datetime, day, month, zone, burn_status, everything())
```

```{r}
#combining persistent and transitional burnedto create a 15min_burned datatable
final_15min_ns_b <- bind_rows(final_persb_ns_temps, final_transb_ns_temps) %>%
  group_by(datetime, day, month, zone, burn_status) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((south_temp + north_temp) / 2)) %>%
  select(datetime, day, month, zone, burn_status, everything())
```

```{r}
#adding in an hourly version of this final_ns dataframe
##unburned
final_hourly_ns_ub <- final_15min_ns_ub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, burn_status) %>%
  summarize(north_temp = mean(north_temp, na.rm = TRUE), south_temp = mean(south_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
final_hourly_ns_b <- final_15min_ns_b %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, burn_status) %>%
  summarize(north_temp = mean(north_temp, na.rm = TRUE), south_temp = mean(south_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#writing final n/s combined dataframes to n-drive
write_csv(final_15min_ns_b, "nsf/treetemp_data/final_15min_ns_b.csv")

write_csv(final_15min_ns_ub, "nsf/treetemp_data/final_15min_ns_ub.csv")

write_csv(final_hourly_ns_b, "nsf/treetemp_data/final_hourly_ns_b.csv")

write_csv(final_hourly_ns_ub, "nsf/treetemp_data/final_hourly_ns_ub.csv")
```

## GB/DB/UB Aspects

```{r}
# GB/DB comparison
final_15min_gbdb <- bind_rows(final_persb_gbdb_temps, final_transb_gbdb_temps) %>%
  #adding in day & month
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((gb_temp + db_temp) / 2)) %>%
  select(datetime, day, month, zone, everything())

#adding in an hourly version of this final_15min_gbdb dataframe
final_hourly_gbdb <- final_15min_gbdb %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d"), tz = "MST"),
         month = as.numeric(format(datetime, "%m")),
         hour = as.numeric(format(datetime, "%H"))) %>%
  group_by(datetime_hour, day, hour, month, zone, aspect) %>%
  summarize(gb_temp = mean(gb_temp, na.rm = TRUE), db_temp = mean(db_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#creating dataframe for combined ub temps

final_15min_gbdb_ub <- bind_rows(final_pers_ub_temps, final_trans_ub_temps) %>%
  #adding in day & month
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  select(datetime, day, month, zone, everything())

#adding in an hourly version of this final_15min_ub dataframe
final_hourly_gbdb_ub <- final_15min_gbdb_ub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d"), tz = "MST"),
         month = as.numeric(format(datetime, "%m")),
         hour = as.numeric(format(datetime, "%H"))) %>%
  group_by(datetime_hour, day, hour, month, zone, aspect) %>%
  summarize(ub_avg_temp = mean(ub_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#GB/UB comparison

##have to create a new dataframe for this step that gets rid of db temps
final_gb_temps <- final_15min_gbdb %>%
  select(c(-db_temp, -avg_temp))

final_15min_gbub <- final_gb_temps %>%
  full_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "aspect", "zone")) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((gb_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything())

#adding in an hourly version of this final_15min_gbdb dataframe
final_hourly_gbub <- final_15min_gbub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d"), tz = "MST"),
         month = as.numeric(format(datetime, "%m")),
         hour = as.numeric(format(datetime, "%H"))) %>%
  group_by(datetime_hour, day, hour, month, zone, aspect) %>%
  summarize(gb_temp = mean(gb_temp, na.rm = TRUE), 
            # Replace ub_temp with the correct column name if needed
            ub_temp = mean(ub_temp, na.rm = TRUE), 
            avg_temp = mean(avg_temp, na.rm = TRUE), 
            .groups = 'drop')
```

```{r}
#DB/UB comparison
final_db_temps <- final_15min_gbdb %>%
  select(c(-gb_temp, -avg_temp))

final_15min_dbub <- final_db_temps %>%
  full_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "aspect", "zone")) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((db_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything())

final_hourly_dbub <- final_15min_dbub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d"), tz = "MST"),
         month = as.numeric(format(datetime, "%m")),
         hour = as.numeric(format(datetime, "%H"))) %>%
  group_by(datetime_hour, day, hour, month, zone, aspect) %>%
  summarize(db_temp = mean(db_temp, na.rm = TRUE), 
            # Replace ub_temp with the correct column name if needed
            ub_temp = mean(ub_temp, na.rm = TRUE), 
            avg_temp = mean(avg_temp, na.rm = TRUE), 
            .groups = 'drop')
```

### Writing CSV

```{r}
#writing final gb/db/ub 15-minute comparison dataframes to n-drive
write_csv(final_15min_gbdb, "nsf/treetemp_data/final_15min_gbdb.csv")

write_csv(final_hourly_gbdb, "nsf/treetemp_data/final_hourly_gbdb.csv")

#this one is only the unburned temps! Not a comparison
write_csv(final_15min_gbdb_ub, "nsf/treetemp_data/final_15min_gbdb_ub.csv")

write_csv(final_hourly_gbdb_ub, "nsf/treetemp_data/final_hourly_gbdb_ub.csv")

write_csv(final_15min_gbub, "nsf/treetemp_data/final_15min_gbub.csv")

write_csv(final_hourly_gbub, "nsf/treetemp_data/final_hourly_gbub.csv")

write_csv(final_15min_dbub, "nsf/treetemp_data/final_15min_dbub.csv")

write_csv(final_hourly_dbub, "nsf/treetemp_data/final_hourly_dbub.csv")
```

# Day/Night timeframes

## Creating day-only dataframes

```{r}
#persistent daytime dataframes
dt_pers_b_aspects <- filtered_pers_b_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

#transitional daytime dataframes
dt_trans_b_aspects <- filtered_trans_b_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

## Plotting - DB/GB differencing

### Persistent Zone

```{r}
#manipulating the dataset to only be hours between 6 am and 9 pm MST
dt_persb_north_diff_long <- persb_north_diff_long %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_persb_south_diff_long <- persb_south_diff_long %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

Plotting

```{r}
#plotting
dt_persb_north_diff_plot <- ggplot(dt_persb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(dt_persb_north_diff_plot)

dt_persb_south_diff_plot <- ggplot(dt_persb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(dt_persb_south_diff_plot)
```

## Plotting - aspect differencing

### Persistent Zone

```{r}
#creating new dataframe with only daytime hours
dt_persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

```{r}
#plotting
dt_persb_gbdb_diff_plot <- ggplot(dt_persb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime Persistent Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(dt_persb_gbdb_diff_plot)
```

### Transitional Zone

```{r}
#creating new dataframe with only daytime hours
dt_transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

Plotting

```{r}
#plotting
dt_transb_gbdb_diff_plot <- ggplot(dt_transb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime Transitional Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(dt_transb_gbdb_diff_plot)
```

# Day vs Night Stats Dataframes

### Calling in Data

The single_avg temps are where a single tree was chosen, then the north/south aspects were averaged

```{r}
#dataframe with single mean daily values calculated for daytime and nighttime, averaging not between N/S but for all trees.
single_avg_all_treetemps <- read_excel("nsf/single_avg_all_treetemps.xlsx")
```

Data Wrangling

```{r}
#manipulating dataframe to run stats on
stats_single_avg_all_treetemps <- single_avg_all_treetemps %>%
  group_by(date, phase, burn_status, zone) %>%
  pivot_wider(names_from = phase, values_from = avg_temp) %>%
  rename(avg_day_temp = day, avg_night_temp = night) %>%
  reframe(avg_day_temp, avg_night_temp)
  
##creatinging individual dataframes for day and night temps, then going to rebind them to get rid of the NA issues
day_all_temps <- stats_single_avg_all_treetemps %>%
  select(-avg_night_temp) %>%
  filter(!is.na(avg_day_temp))# Filter out rows with NA

night_all_temps <- stats_single_avg_all_treetemps %>%
  select(-avg_day_temp) %>%
  filter(!is.na(avg_night_temp))

##joining north/south dataframes
final_daynight_singleavg_temps <- day_all_temps %>%
  inner_join(night_all_temps) %>%
  #adding in week and month columns
  mutate(
    month = format(date, "%Y-%m"),
    week = week(date)  # Extract week number from datetime
  ) %>%
  mutate(avg_temp = (avg_day_temp + avg_night_temp) / 2)
```

Exporting final daynight single_avg dataset

```{r}
write_csv(final_daynight_singleavg_temps, "nsf/treetemp_data/final_daynight_singleavg_temps.csv")
```

# Creating a single comparison dataframe for N/S and GB/DB/UB comparisons and stats

### 15-minute timestep

```{r}
full_15min_ns <- bind_rows(full_pers_ns_data, full_trans_ns_data) %>%
  #adding in day, month and time columns
  mutate(datetime = as.POSIXct(datetime, tz = "MST"), 
         day = format(datetime, "%Y-%m-%d"),  # Extract the day in 'YYYY-MM-DD' format as a string
         month = format(datetime, "%m"),      # Extract the month as a numeric string
         time = format(datetime, "%H:%M:%S")) %>%
#adding in mean temp of gbdb/gbub/dbub temps
  mutate(ns_avg_temp = mean((north_temp + south_temp) / 2)) %>%
  select(datetime, time, day, month, zone, everything()) %>%
  #calculating temp difference for 15-minute intervals
  mutate(temp_diff = south_temp - north_temp)
```

```{r}
# GB/DB/UB comparison
full_15min_gbdbub <- bind_rows(final_persb_gbdb_temps, final_transb_gbdb_temps) %>%
  full_join(final_15min_gbdb_ub) %>%
  #adding in day, month and time columns
 mutate(datetime = as.POSIXct(datetime, tz = "MST"), 
         day = format(datetime, "%Y-%m-%d"),  # Extract the day in 'YYYY-MM-DD' format as a string
         month = format(datetime, "%m"),      # Extract the month as a numeric string
         time = format(datetime, "%H:%M:%S")) %>%
  #adding in mean temp of gbdb/gbub/dbub temps
  mutate(gbdb_avg_temp = mean((gb_temp + db_temp) / 2)) %>%
  mutate(gbub_avg_temp = mean((gb_temp + ub_temp) / 2)) %>%
  mutate(dbub_avg_temp = mean((db_temp + ub_temp) / 2)) %>%
  select(datetime, time, day, month, zone, everything()) %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbdb_temp_diff = gb_temp - db_temp) %>%
  mutate(gbub_temp_diff = gb_temp - ub_temp) %>%
  mutate(dbub_temp_diff = db_temp - ub_temp)
```

Exporting full 15-minute gbdbub and n/s comparison dataframes

```{r}
write_csv(full_15min_ns, "nsf/treetemp_data/full_15min_ns.csv")

write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

### hourly timestep

```{r}
full_hourly_ns_b <- bind_rows(final_persb_ns_temps, final_transb_ns_temps) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = format(datetime, "%Y-%m-%d"),
         month = format(datetime, "%m"),
         time = format(datetime, "%H:%M:%S")) %>%
  mutate(ns_avg_temp = mean((north_temp + south_temp) / 2)) %>%
  select(datetime, time, day, month, zone, everything()) %>%
  #calculating temp difference for 15-minute intervals
  mutate(temp_diff = south_temp - north_temp) %>%
  select(datetime, day, time, month, zone, burn_status, north_temp, south_temp, ns_avg_temp, temp_diff)

full_hourly_ns_ub <- bind_rows(final_pers_ub_ns_temps, final_trans_ub_ns_temps) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  mutate(day = format(datetime, "%Y-%m-%d"),
         month = format(datetime, "%m"),
         time = format(datetime, "%H:%M:%S")) %>%
  mutate(ns_avg_temp = mean((north_temp + south_temp) / 2)) %>%
  select(datetime, time, day, month, zone, everything()) %>%
  #calculating temp difference for 15-minute intervals
  mutate(temp_diff = south_temp - north_temp) %>%
  select(datetime, day, time, month, zone, burn_status, north_temp, south_temp, ns_avg_temp, temp_diff)

#making a final dataframe that has burned and unburned 
full_hourly_ns <- bind_rows(full_hourly_ns_b, full_hourly_ns_ub)
```

```{r}
#gb/db/ub comparisons
# Combine datasets and calculate hourly averages
full_hourly_gbdbub <- final_persb_gbdb_temps %>%
  full_join(final_transb_gbdb_temps) %>%
  full_join(final_15min_gbdb_ub) %>% # Ensure you're joining on the correct column
  # Add date and hour columns
  mutate(datetime_hour = floor_date(datetime, unit = "hour"),
         day = format(datetime_hour, "%Y-%m-%d"),
         month = format(datetime_hour, "%m"),
         time = format(datetime_hour, "%H:%M:%S")) %>%
  # Calculate temperature differences
  mutate(gbdb_temp_diff = gb_temp - db_temp, 
         gbub_temp_diff = gb_temp - ub_temp, 
         dbub_temp_diff = db_temp - ub_temp) %>%
  # Group by the new hourly datetime column
  group_by(datetime_hour, day, month, zone, aspect) %>%
  summarize(
    db_temp = mean(db_temp, na.rm = TRUE), 
    ub_temp = mean(ub_temp, na.rm = TRUE), 
    gb_temp = mean(gb_temp, na.rm = TRUE), 
    gbdb_avg_temp = mean(c(gb_temp, db_temp), na.rm = TRUE), 
    gbub_avg_temp = mean(c(gb_temp, ub_temp), na.rm = TRUE), 
    dbub_avg_temp = mean(c(db_temp, ub_temp), na.rm = TRUE), 
    gbdb_temp_diff = mean(gbdb_temp_diff, na.rm = TRUE), 
    gbub_temp_diff = mean(gbub_temp_diff, na.rm = TRUE), 
    dbub_temp_diff = mean(dbub_temp_diff, na.rm = TRUE), 
    .groups = 'drop'
  )
```

```{r}
write_csv(full_hourly_ns, "nsf/treetemp_data/full_hourly_ns.csv")

write_csv(full_hourly_gbdbub, "nsf/treetemp_data/full_hourly_gbdbub.csv")
```

## Plotting charred/uncharred comparisons

```{r}
#making a daily dt/nt dataset
daily_final_gb_temps <- final_hourly_gb_temps %>%
  mutate(
    date = date(datetime_hour),  # Create the date column
    phase = case_when(
      hour(datetime_hour) >= 8 & hour(datetime_hour) < 17 ~ "day",
      TRUE ~ "night"
    )) %>%
  group_by(date, month, phase, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_south_temp = mean(mean_south_temp), mean_charred_temp = mean(mean_charred_temp), mean_uncharred_temp = mean(mean_uncharred_temp))

#adding day/night columns to hourly gb_temps
final_hourly_gb_temps <- final_hourly_gb_temps %>%
  mutate(
    date = date(datetime_hour),  # Create the date column
    phase = case_when(
      hour(datetime_hour) >= 8 & hour(datetime_hour) < 17 ~ "day",
      TRUE ~ "night"
    )) %>%
  select(datetime_hour, date, month, phase, everything())
```

```{r}
#day/night mean temps plots
dtnt_gb_aspect_comparison <- ggplot(data = daily_final_gb_temps) +
  geom_line(aes(x = date, y = mean_north_temp, color = "North Temp")) +
  geom_line(aes(x = date, y = mean_south_temp, color = "South Temp")) +
  geom_line(aes(x = date, y = mean_charred_temp, color = "Charred Temp")) +
  geom_line(aes(x = date, y = mean_uncharred_temp, color = "Uncharred Temp")) +
  facet_wrap(~ zone + phase, dir = "v") +
  scale_color_manual(
    values = c("North Temp" = "green", "South Temp" = "red", "Charred Temp" = "black", "Uncharred Temp" = "grey"),
    name = "Temperature Type"  # Legend title
  ) + 
  facet_wrap(~phase + zone)


dtnt_gb_aspect_comparison 
```

The above plot makes it look like the charred temps are the highest, not sure why?

```{r}
#day/night hourly temps plots
hourly_gb_aspect_comparison <- ggplot(data = final_hourly_gb_temps) +
  geom_line(aes(x = date, y = mean_north_temp, color = "North Temp")) +
  geom_line(aes(x = date, y = mean_south_temp, color = "South Temp")) +
  geom_line(aes(x = date, y = mean_charred_temp, color = "Charred Temp")) +
  geom_line(aes(x = date, y = mean_uncharred_temp, color = "Uncharred Temp")) +
  facet_wrap(~ zone + phase, dir = "v") +
  scale_color_manual(
    values = c("North Temp" = "green", "South Temp" = "red", "Charred Temp" = "black", "Uncharred Temp" = "grey"),
    name = "Temperature Type"  # Legend title
  ) + 
  facet_wrap(~phase + zone)


hourly_gb_aspect_comparison 
```

### Box Plot of hourly data

```{r}
## have to make a new, long, dataframe first
final_hourly_gb_temps_long <- final_hourly_gb_temps %>%
  #renaming aspect variables so it only says "south" not "mean_south_temp"
  rename_with(~ gsub("_temp", "", gsub("mean_", "", .)), starts_with("mean_")) %>%
  pivot_longer(cols = c(north, south, charred, uncharred),
               names_to = "temp_type", values_to = "temperature")
```

```{r}
#creating a daytime dataframe and a night time dataframe to make the facet_wrapped plots more easy to see
##persistent
dt_p_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "day", zone == "persistent")

nt_p_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "night", zone == "persistent")

##transitional
dt_t_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "day", zone == "transitional")

nt_t_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "night", zone == "transitional")
```

```{r}
#hourly boxplot
hourly_gb_aspect_comparison_boxplot <- ggplot(data = final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(hourly_gb_aspect_comparison_boxplot)
```

```{r}
#dt hourly boxplot - persistent
dt_p_hourly_gb_aspect_comparison_boxplot <- ggplot(data = dt_p_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_p_hourly_gb_aspect_comparison_boxplot)
```

```{r}
dt_t_hourly_gb_aspect_comparison_boxplot <- ggplot(data = dt_t_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_t_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt plots next to each other
dt_hourly_gb_comb_plots <- ggarrange(
  dt_p_hourly_gb_aspect_comparison_boxplot,
  dt_t_hourly_gb_aspect_comparison_boxplot,
  nrow = 1,
  common.legend = TRUE,
  legend = "right")

# adding a title to combined plot
dt_hourly_gb_comb_plots <- annotate_figure(
  dt_hourly_gb_comb_plots,
  top = text_grob("Daytime Hourly Mean Temps", size = 14)
)

# Display the combined plot
dt_hourly_gb_comb_plots
```

Night Time Plots by zone

```{r}
#nt hourly boxplot - persistent
nt_p_hourly_gb_aspect_comparison_boxplot <- ggplot(data = nt_p_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(nt_p_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#nighttime hourly boxplot - transitional
nt_t_hourly_gb_aspect_comparison_boxplot <- ggplot(data = nt_t_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(nt_t_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt plots next to each other
nt_hourly_gb_comb_plots <- ggarrange(
  nt_p_hourly_gb_aspect_comparison_boxplot, 
  nt_t_hourly_gb_aspect_comparison_boxplot,
  nrow = 1,
  common.legend = TRUE,
  legend = "right")

nt_hourly_gb_comb_plots
```

Boxplots by Mean Daily Data

```{r}
## have to make a new, long, dataframe first
final_daily_gb_temps_long <- daily_final_gb_temps %>%
  #renaming aspect variables so it only says "south" not "mean_south_temp"
  rename_with(~ gsub("_temp", "", gsub("mean_", "", .)), starts_with("mean_")) %>%
  pivot_longer(cols = c(north, south, charred, uncharred),
               names_to = "temp_type", values_to = "temperature")
```

Day/Night Long Dataframes - for boxplot plotting

```{r}
dt_p_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "day", zone == "persistent")

nt_p_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "night", zone == "persistent")

##transitional
dt_t_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "day", zone == "transitional")

nt_t_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "night", zone == "transitional")
```

Plotting Daily Means by Phase + Zone

```{r}
#dt hourly boxplot - persistent
dt_p_daily_gb_aspect_comparison_boxplot <- ggplot(data = dt_p_final_daily_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_p_daily_gb_aspect_comparison_boxplot)
```

```{r}
#dt hourly boxplot - persistent
dt_t_daily_gb_aspect_comparison_boxplot <- ggplot(data = dt_t_final_daily_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_t_daily_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt plots next to each other
dt_daily_gb_comb_plots <- ggarrange(
  dt_p_daily_gb_aspect_comparison_boxplot,
  dt_t_daily_gb_aspect_comparison_boxplot,
  nrow = 1,
  common.legend = TRUE,
  legend = "right")

# adding a title to combined plot
dt_daily_gb_comb_plots <- annotate_figure(
  dt_daily_gb_comb_plots,
  top = text_grob("Daytime Daily Mean Temps", size = 14)
)

# Display the combined plot
dt_daily_gb_comb_plots
```

\
\
\
\

# Display the combined plot

dt_hourly_gb_comb_plots

```{r}
#new dataframe with charred/uncharred/north/south aspects for feb

feb_t_full_aspects <- filtered_trans_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(month == "2")

feb_p_full_aspects <- filtered_pers_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(tree_name !="gb_w") %>%
  filter(month == "2")

may_t_full_aspects <- filtered_trans_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(month == "5")

may_p_full_aspects <- filtered_pers_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(tree_name !="gb_w") %>%
  filter(month == "5")

```

```{r}

```

```{r}
#plotting feb dates to see if there is a substantial difference between char patterns
t_feb_aspect_plot <- ggplot(data = feb_t_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

t_feb_aspect_plot

p_feb_aspect_plot <- ggplot(data = feb_p_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

p_feb_aspect_plot
```

```{r}
t_may_aspect_plot <- ggplot(data = may_t_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

t_may_aspect_plot

p_may_aspect_plot <- ggplot(data = feb_p_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

p_may_aspect_plot
```

\*\* at the pers may plot -\> looks like charred trees are actually getting much warmer than in feb.
