---
title: "treetemp_stats"
author: "Erika Lee"
date: "2024-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stats Calculations

```{r}
#initial set up
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
#packages for exporting csvs as tables
library(knitr)
library(kableExtra)
library(pagedown)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

## Loading Data

```{r}
final_15min_ns_b <- read_csv("nsf/treetemp_data/final_15min_ns_b.csv")

final_15min_ns_ub <- read_csv("nsf/treetemp_data/final_15min_ns_ub.csv")

final_hourly_ns_b <- read_csv("nsf/treetemp_data/final_hourly_ns_b.csv")

final_hourly_ns_ub <- read_csv("nsf/treetemp_data/final_hourly_ns_ub.csv")

final_15min_gbdb <- read_csv("nsf/treetemp_data/final_15min_gbdb.csv")

final_15min_gbdb_ub <- read_csv("nsf/treetemp_data/final_15min_gbdb_ub.csv")

final_15min_gbub <- read_csv("nsf/treetemp_data/final_15min_gbub.csv")

final_15min_dbub <- read_csv("nsf/treetemp_data/final_15min_dbub.csv")

final_hourly_gbdb <- read_csv("nsf/treetemp_data/final_hourly_gbdb.csv")

final_hourly_gbdb_ub <- read_csv("nsf/treetemp_data/final_hourly_gbdb_ub.csv")

final_hourly_gbub <- read_csv("nsf/treetemp_data/final_hourly_gbub.csv")

final_hourly_dbub <- read_csv("nsf/treetemp_data/final_hourly_dbub.csv")

final_daynight_singleavg_temps <- read_csv("nsf/treetemp_data/final_daynight_singleavg_temps.csv")
```

```{r}
#new full dataframes to load
full_15min_ns <- read_csv("nsf/treetemp_data/full_15min_ns.csv")

full_hourly_ns <- read_csv("nsf/treetemp_data/full_hourly_ns.csv")

full_15min_gbdbub <- read_csv("nsf/treetemp_data/full_15min_gbdbub.csv")

full_hourly_gbdbub <- read_csv("nsf/treetemp_data/full_hourly_gbdbub.csv")
```

Pulling in Data for gb only comparisons of charred/uncharred data with n/s as well

```{r}
final_15min_gb_temps <- read_csv("nsf/treetemp_data/final_15min_gb_temps.csv") %>%
  mutate(day = as.Date(datetime), month = month(datetime))

final_hourly_gb_temps <- read_csv("nsf/treetemp_data/final_hourly_gb_temps.csv")%>%
  mutate(day = as.Date(datetime_hour), month = month(datetime_hour))
```

## Calculating Stats

### N/S Differenced Data

Based on meeting w/ Stephanie on 7/17/24 - I already calculated avg_temp for 15min timestep in "data_wrangling" rmarkdown, now I will calculate the mean_diff for 15minute timestep, then do the other calculations by month

```{r}
## don't use this one!
#burned n/s stats
stats_monthly_ns_b <- final_15min_ns_b %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

```{r}
## don't use this once! Use the one lower down with combined burned/ub
#unburned n/s stats
stats_monthly_ns_ub <- final_15min_ns_ub %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp), mean_south_temp = mean(south_temp), mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

#### Creating N/S differencing tables that are organized by burn status (gb, db, ub)

```{r}
#green burned n/s differences
stats_monthly_ns_gb <- stats_monthly_ns_b %>%
  filter(burn_status == "green burn")

#dead burned n/s differences
stats_monthly_ns_db <- stats_monthly_ns_b %>%
  filter(burn_status == "dead burn")

##don't need to do for ub because I already have that
```

### North/South Differenced Data

\*\* differenced in the same way as the charred/uncharred below!

```{r}
stats_monthly_s_n <- final_15min_gb_temps %>%
  mutate(datetime = as.POSIXct(datetime), month = format(datetime, "%m")) %>%
  #calculating temp_diff for 15-minute interval of charred/uncharred
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(south_temp), mean_north_temp = mean(north_temp), mean_temp = mean((south_temp + north_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### Charred/Uncharred Differenced Data

```{r}
#burned n/s stats
stats_monthly_c_uc <- final_15min_gb_temps %>%
  mutate(datetime = as.POSIXct(datetime), month = format(datetime, "%m")) %>%
  #calculating temp_diff for 15-minute interval of charred/uncharred
  mutate(temp_diff = charred_temp - uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_charred_temp = mean(charred_temp), mean_uncharred_temp = mean(uncharred_temp), mean_temp = mean((charred_temp + uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### South / Uncharred Differenced Data

```{r}
# to compare noth/south with charred/uncharred, I need to make new n/s_gb only datasets
stats_monthly_s_uc <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_south_temp, mean_uncharred_temp) %>%
  mutate(temp_diff = mean_south_temp - mean_uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(mean_south_temp), mean_uncharred_temp = mean(mean_uncharred_temp), mean_temp = mean((mean_south_temp + mean_uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### South / Charred Differenced Data

```{r}
stats_monthly_s_c <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_south_temp, mean_charred_temp) %>%
  mutate(temp_diff = mean_south_temp - mean_charred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_south_temp = mean(mean_south_temp), mean_charred_temp = mean(mean_charred_temp), mean_temp = mean((mean_south_temp + mean_charred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### North / Uncharred Differenced Data

```{r}
stats_monthly_n_uc <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_north_temp, mean_uncharred_temp) %>%
  mutate(temp_diff = mean_north_temp - mean_uncharred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_uncharred_temp = mean(mean_uncharred_temp), mean_temp = mean((mean_north_temp + mean_uncharred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### North / Charred Differenced Data

```{r}
stats_monthly_n_c <- final_hourly_gb_temps %>%
  select(datetime_hour, month, zone, burn_status, mean_north_temp, mean_charred_temp) %>%
  mutate(temp_diff = mean_north_temp - mean_charred_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_charred_temp = mean(mean_charred_temp), mean_temp = mean((mean_north_temp + mean_charred_temp)/2), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/DB Differenced Data

```{r}
#gb/db stats - db minus gb! 
stats_monthly_gbdb <- final_15min_gbdb %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = db_temp - gb_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/UB Differenced Data

```{r}
#stats calculations
stats_monthly_gbub <- final_15min_gbub %>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### DB/UB Differenced Data

```{r}
#stats calculations
stats_monthly_dbub <- final_15min_dbub %>%
  # Removing rows with NAs
  drop_na() %>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### Day/Night Differenced Data

```{r}
#monthly stats calculations by day/night zone & burn status
stats_monthly_daynight <- final_daynight_singleavg_temps %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = avg_day_temp - avg_night_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### New Differenced Data from full datasets

#### NS Comparison

```{r}
#creating a monthly dataframe to carry over avg and temp_difference data into full stats dataframe - can't bind this with the stats because we need 15-minute numbers to get the correct stats
monthly_ns_data <- full_15min_ns %>%
  group_by(month, zone, burn_status) %>%
  summarize(mean_north_temp = mean(north_temp, na.rm = TRUE), mean_south_temp = mean(south_temp, na.rm = TRUE), ns_avg_temp = mean(ns_avg_temp, na.rm = TRUE), temp_diff = mean(temp_diff, na.rm = TRUE), .groups = 'drop')
```

```{r}
#n/s monthly stats
stats_monthly_ns <- full_15min_ns %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(ns_mean_temp = mean(ns_avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

#### Charred/Uncharred/N/S GB Comparisons

```{r}
monthly_gb_data <- final_15min_gb_temps %>%
  mutate(
    datetime = as.POSIXct(datetime),
    month = format(datetime, "%m")
  ) %>%
  select(datetime, day, month, everything()) %>%
  mutate(
    ns_avg_temp = rowMeans(cbind(north_temp, south_temp)),
    c_uc_avg_temp = rowMeans(cbind(charred_temp, uncharred_temp)),
    s_c_avg_temp = rowMeans(cbind(south_temp, charred_temp)),
    s_uc_avg_temp = rowMeans(cbind(south_temp, uncharred_temp)),
    n_c_avg_temp = rowMeans(cbind(north_temp, charred_temp)),
    n_uc_avg_temp = rowMeans(cbind(north_temp, uncharred_temp)),
    ns_temp_diff = north_temp - south_temp,
    c_uc_temp_diff = charred_temp - uncharred_temp,
    s_c_temp_diff = south_temp - charred_temp,
    s_uc_temp_diff = south_temp - uncharred_temp,
    n_c_temp_diff = north_temp - charred_temp,
    n_uc_temp_diff = north_temp - uncharred_temp
  ) %>%
  group_by(month, zone, burn_status) %>%
  summarize(
    mean_north_temp = mean(north_temp, na.rm = TRUE),
    mean_south_temp = mean(south_temp, na.rm = TRUE),
    mean_charred_temp = mean(charred_temp, na.rm = TRUE),
    mean_uncharred_temp = mean(uncharred_temp, na.rm = TRUE),
    ns_avg_temp = mean(ns_avg_temp, na.rm = TRUE),
    c_uc_avg_temp = mean(c_uc_avg_temp, na.rm = TRUE),
    s_c_avg_temp = mean(s_c_avg_temp, na.rm = TRUE),
    s_uc_avg_temp = mean(s_uc_avg_temp, na.rm = TRUE),
    n_c_avg_temp = mean(n_c_avg_temp, na.rm = TRUE),
    n_uc_avg_temp = mean(n_uc_avg_temp, na.rm = TRUE),
    ns_temp_diff = mean(ns_temp_diff, na.rm = TRUE),
    c_uc_temp_diff = mean(c_uc_temp_diff, na.rm = TRUE),
    s_c_temp_diff = mean(s_c_temp_diff, na.rm = TRUE),
    s_uc_temp_diff = mean(s_uc_temp_diff, na.rm = TRUE),
    n_c_temp_diff = mean(n_c_temp_diff, na.rm = TRUE),
    n_uc_temp_diff = mean(n_uc_temp_diff, na.rm = TRUE),
    .groups = 'drop'
  )

```

Don't use the below code

```{r}
## dont use this!
#charred/uncharred monthly stas
stats_monthly_c_uc_V2 <- monthly_gb_data %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(c_uc_mean_temp = mean(c_uc_avg_temp), mean_diff = mean(c_uc_temp_diff), sd = sd(c_uc_temp_diff, na.rm = TRUE),
    coeff_var = (sd(c_uc_temp_diff, na.rm = TRUE) / abs(mean(c_uc_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate(across(where(is.numeric), round, 3))
```

#### GBDBUB Comparison

```{r}
monthly_gbdbub_data <- full_15min_gbdbub %>%
  group_by(month, zone) %>%
  summarize(mean_gb_temp = mean(gb_temp, na.rm = TRUE), mean_db_temp = mean(db_temp, na.rm = TRUE), mean_ub_temp = mean(ub_temp, na.rm = TRUE), gbdb_avg_temp = mean(gbdb_avg_temp, na.rm = TRUE), gbub_avg_temp = mean(gbub_avg_temp, na.rm = TRUE), dbub_avg_temp = mean(dbub_avg_temp, na.rm = TRUE), gbdb_temp_diff = mean(gbdb_temp_diff, na.rm = TRUE), gbub_temp_diff = mean(gbub_temp_diff, na.rm = TRUE), dbub_temp_diff = mean(dbub_temp_diff, na.rm = TRUE), .groups = 'drop')
```

```{r}
#making a stats dataframe with combined data did not work, so I am going to bind the already created gbdbub dataframes so I have a full gbdbub stats dataframe

#gb/db stats - db minus gb! 
stats_monthly_gbdb_V2 <- final_15min_gbdb %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbdb_temp_diff = db_temp - gb_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(gbdb_mean_temp = mean(avg_temp), gbdb_mean_diff = mean(gbdb_temp_diff), gbdb_sd = sd(gbdb_temp_diff, na.rm = TRUE),
    gbdb_coeff_var = (sd(gbdb_temp_diff, na.rm = TRUE) / abs(mean(gbdb_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

#gb/ub stats
stats_monthly_gbub_V2 <- final_15min_gbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbub_temp_diff = gb_temp - ub_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(gbub_mean_temp = mean(avg_temp), gbub_mean_diff = mean(gbub_temp_diff), gbub_sd = sd(gbub_temp_diff, na.rm = TRUE),
    gbub_coeff_var = (sd(gbub_temp_diff, na.rm = TRUE) / abs(mean(gbub_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

#db/ub stats
stats_monthly_dbub_V2 <- final_15min_dbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(dbub_temp_diff = db_temp - ub_temp)%>%
  #calculating monthly stats
  group_by(month, zone) %>%
  drop_na() %>%
  summarize(dbub_mean_temp = mean(avg_temp), dbub_mean_diff = mean(dbub_temp_diff), dbub_sd = sd(dbub_temp_diff, na.rm = TRUE),
    dbub_coeff_var = (sd(dbub_temp_diff, na.rm = TRUE) / abs(mean(dbub_temp_diff, na.rm = TRUE)))) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)

##joining all V2 dataframes
stats_monthly_gbdbub <- stats_monthly_gbdb_V2 %>%
  full_join(stats_monthly_gbub_V2) %>%
  full_join(stats_monthly_dbub_V2)
```

#### Saving New full stats as CSVs

These are full csv tables, not seperated by burn/unburned status

```{r}
write_csv(monthly_ns_data, "nsf/treetemp_data/monthly_ns_data.csv")

write_csv(stats_monthly_ns, "nsf/treetemp_data/stats_monthly_ns.csv")

write_csv(monthly_gbdbub_data, "nsf/treetemp_data/monthly_gbdbub_data.csv")

write_csv(stats_monthly_gbdbub, "nsf/treetemp_data/monthly_gbdbub_data.csv")
```

## Saving Stats CSV

### Exporting CSVs to N-drive

```{r}
write_csv(stats_monthly_ns_b, "nsf/treetemp_data/stats_monthly_ns_b.csv")

write_csv(stats_monthly_ns_ub, "nsf/treetemp_data/stats_monthly_ns_ub.csv")

write_csv(stats_monthly_ns_gb, "nsf/treetemp_data/stats_monthly_ns_gb.csv")

write_csv(stats_monthly_ns_db, "nsf/treetemp_data/stats_monthly_ns_db.csv")

write_csv(stats_monthly_gbdb, "nsf/treetemp_data/stats_monthly_gbdb.csv")

write_csv(stats_monthly_gbub, "nsf/treetemp_data/stats_monthly_gbub.csv")

write_csv(stats_monthly_dbub, "nsf/treetemp_data/stats_monthly_dbub.csv")

write_csv(stats_monthly_daynight, "nsf/treetemp_data/stats_monthly_daynight.csv")

#new charred/uncharred/n/s comparisons
write_csv(stats_monthly_c_uc, "nsf/treetemp_data/stats_montly_c_uc.csv")

write_csv(stats_monthly_s_c, "nsf/treetemp_data/stats_montly_s_c.csv")

write_csv(stats_monthly_s_uc, "nsf/treetemp_data/stats_montly_s_uc.csv")

write_csv(stats_monthly_n_c, "nsf/treetemp_data/stats_montly_n_c.csv")

write_csv(stats_monthly_n_uc, "nsf/treetemp_data/stats_montly_n_uc.csv")
```

```{r}
write_csv(stats_monthly_s_n, "nsf/treetemp_data/stats_monthly_s_n.csv")
```

\*\*not using the below conversion process as of 2024-07-22, just adding csvs to word doc and turning into a table

# Converting CSVs to Tables

```{r}
#creating one NS table with subtitles of burned and unburned (seperate tables)

# Create HTML table with kable and kableExtra
ns_b_statstable <- kable(stats_monthly_ns_b, "html", caption = "Monthly Statistics of North and South Aspect Tree Temp Comparison, for Burned Zone") %>%
  kable_classic(full_width = F, html_font = "Cambria")

ns_ub_statstable <- kable(stats_monthly_ns_ub, "html", caption = "unburned") %>%
  kable_classic(full_width = F, html_font = "Cambria")

# Save the HTML tables to separate files
html_file_ns_b <- "ns_b_stats_table.html"
writeLines(as.character(ns_b_statstable), html_file_ns_b)

html_file_ns_ub <- "ns_ub_stats_table.html"
writeLines(as.character(ns_ub_statstable), html_file_ns_ub)

# Optionally, open the HTML files in the default web browser
browseURL(html_file_ns_b)
browseURL(html_file_ns_ub)
```

#the above process worked, but the html table doesn't look that good. Consider trying to make one large table, with all the sub tables included with sub-headers. That would work better I think?

# Plotting Temp Differences

```{r}

```

# Process to turn csv's into tables after exported

1.  Open CSV in Text Editor

2.  Copy all text

3.  Paste into Word Document

4.  highlight all text in word doc

5.  Click Tables \> Convert \> Convert text to table (make sure comma delimination is highlighted)

6.  Should appear as a table in the word document now

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

# Understanding the Statistics

When you calculate an average difference between two temperatures and analyze the percent difference, average, and coefficient of variation, each metric provides different insights into your data:

1.  **Percent Difference**:

    -   **What it tells you**: The percent difference gives you a relative measure of the difference between the two temperatures compared to their average. It tells you how large the difference is in relation to the overall magnitude of the temperatures.

    -   **Interpretation**: A high percent difference indicates a significant discrepancy between the two temperatures, whereas a low percent difference suggests that the temperatures are relatively similar.

2.  **Average**:

    -   **What it tells you**: The average temperature difference provides a measure of the central tendency of the differences between the two temperatures. It is the mean value of all the calculated differences.

    -   **Interpretation**: This metric gives you an idea of the typical or expected difference between the temperatures. It is useful for understanding the overall trend in the temperature differences.

3.  **Coefficient of Variation (CV)**:

    -   **What it tells you**: The coefficient of variation is a standardized measure of dispersion of the temperature differences. It is calculated as the ratio of the standard deviation to the mean and is usually expressed as a percentage.

    -   **Interpretation**: A high CV indicates that the temperature differences are highly variable, suggesting inconsistency or high variability in the temperature differences. A low CV indicates that the temperature differences are relatively consistent and stable.

### Example Scenario

Suppose you have two sets of temperature readings (in °C) from two different sensors or locations. After calculating the differences between the corresponding readings, you obtain the following statistics:

-   **Average Difference**: 2°C

-   **Percent Difference**: 10%

-   **Coefficient of Variation**: 25%

#### Interpretation:

-   **Average Difference**: On average, there is a 2°C difference between the temperatures recorded by the two sensors or at the two locations.

-   **Percent Difference**: The 10% difference indicates that the temperature difference is relatively small compared to the average temperature values of the readings.

-   **Coefficient of Variation**: A CV of 25% suggests moderate variability in the temperature differences. While there is some fluctuation in the differences, it is not extremely high, indicating that the differences are somewhat consistent but not entirely uniform.

In summary, these metrics together help you understand the magnitude, relative size, and variability of the temperature differences in your data.

## My Understanding of the Data

-   very high coefficient of variation and percent differences mean that there is a lot of variability of temperatures, even on a daily and monthly scale.

\*\*Made edits to test git connection
