---
title: "treetemp_stats"
author: "Erika Lee"
date: "2024-07-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Stats Calculations

```{r}
#initial set up
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(xlsx)
#packages for exporting csvs as tables
library(knitr)
library(kableExtra)
library(pagedown)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

## Loading Data

```{r}
final_15min_ns_b <- read_csv("nsf/treetemp_data/final_15min_ns_b.csv")

final_15min_ns_ub <- read_csv("nsf/treetemp_data/final_15min_ns_ub.csv")

final_15min_gbdb_b <- read_csv("nsf/treetemp_data/final_15min_gbdb_b.csv")

final_15min_gbdb_ub <- read_csv("nsf/treetemp_data/final_15min_gbdb_ub.csv")

final_single_avg_all_treetemps <- read_csv("nsf/treetemp_data/final_daynight_singleavg_temps.csv")
```

## Calculating Stats

### N/S Differenced Data

Based on meeting w/ Stephanie on 7/17/24 - I already calculated avg_temp for 15min timestep in "data_wrangling" rmarkdown, now I will calculate the mean_diff for 15minute timestep, then do the other calculations by month

```{r}
#burned n/s stats
stats_monthly_ns_b <- final_15min_ns_b %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100
  ) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

```{r}
#unburned n/s stats
stats_monthly_ns_ub <- final_15min_ns_ub %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = south_temp - north_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100
  ) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/DB Differenced Data

```{r}
#gb/db stats - db minus gb! 
stats_monthly_gbdb <- final_15min_gbdb_b %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = db_temp - gb_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100
  ) %>%
  select(-burn_status) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### GB/UB Differenced Data

```{r}
#have to create new dataframes to difference gb from unburned

##manipulating loaded gbdb_ub dataframe first
final_15min_gbdb_ub <- final_15min_gbdb_ub %>%
  #removing avg_temp column from ub dataframe
  select(-c(avg_temp, burn_status))

final_15min_gbub <- final_15min_gbdb_b %>%
  #removing db data from dataframe
  select(-c(db_temp, avg_temp, burn_status)) %>%
  inner_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "zone", "aspect")) %>%
  #calculating avg_temp for 15-minute interval
  mutate(avg_temp = (gb_temp+ub_temp)/2) 
```

```{r}
#exporting new gb/ub dataframe
write_csv(final_15min_gbub, "nsf/treetemp_data/final_15min_gbub.csv")
```

```{r}
#stats calculations
stats_monthly_gbub <- final_15min_gbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = gb_temp - ub_temp) %>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

### DB/UB Differenced Data

```{r}
#have to create new dataframes to difference db from unburned
final_15min_dbub <- final_15min_gbdb_b %>%
  #removing db data from dataframe
  select(-c(gb_temp, avg_temp, burn_status)) %>%
  inner_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "zone", "aspect")) %>%
  #calculating avg_temp for 15-minute interval
  mutate(avg_temp = (db_temp+ub_temp)/2)
```

```{r}
#exporting new gb/ub dataframe
write_csv(final_15min_dbub, "nsf/treetemp_data/final_15min_dbub.csv")
```

```{r}
#stats calculations
stats_monthly_dbub <- final_15min_dbub %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = db_temp - ub_temp) %>%
  #calculating monthly stats
  group_by(month, zone) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

Day/Night Differenced Data

```{r}
#monthly stats calculations by day/night zone & burn status
stats_monthly_daynight <- final_daynight_singleavg_temps %>%
  #calculating temp_diff for 15-minute interval
  mutate(temp_diff = avg_day_temp - avg_night_temp) %>%
  #calculating monthly stats
  group_by(month, zone, burn_status) %>%
  summarize(mean_temp = mean(avg_temp), mean_diff = mean(temp_diff), sd = sd(temp_diff, na.rm = TRUE),
    coeff_var = (sd(temp_diff, na.rm = TRUE) / abs(mean(temp_diff, na.rm = TRUE))) * 100) %>%
  #rounding numeric columns to 3 decimals
  mutate_if(is.numeric, round, 3)
```

## Saving Stats CSV

### Exporting CSVs to N-drive

```{r}
write_csv(stats_monthly_ns_b, "nsf/treetemp_data/stats_monthly_ns_b.csv")

write_csv(stats_monthly_ns_ub, "nsf/treetemp_data/stats_monthly_ns_ub.csv")

write_csv(stats_monthly_gbdb, "nsf/treetemp_data/stats_monthly_gbdb.csv")

write_csv(stats_monthly_gbub, "nsf/treetemp_data/stats_monthly_gbub.csv")

write_csv(stats_monthly_dbub, "nsf/treetemp_data/stats_monthly_dbub.csv")

write_csv(stats_monthly_daynight, "nsf/treetemp_data/stats_monthly_daynight.csv")
```

# Converting CSVs to Tables

```{r}
#creating one NS table with subtitles of burned and unburned (seperate tables)

# Create HTML table with kable and kableExtra
ns_b_statstable <- kable(stats_monthly_ns_b, "html", caption = "Monthly Statistics of North and South Aspect Tree Temp Comparison, for Burned Zone") %>%
  kable_classic(full_width = F, html_font = "Cambria")

ns_ub_statstable <- kable(stats_monthly_ns_ub, "html", caption = "unburned") %>%
  kable_classic(full_width = F, html_font = "Cambria")

# Save the HTML tables to separate files
html_file_ns_b <- "ns_b_stats_table.html"
writeLines(as.character(ns_b_statstable), html_file_ns_b)

html_file_ns_ub <- "ns_ub_stats_table.html"
writeLines(as.character(ns_ub_statstable), html_file_ns_ub)

# Optionally, open the HTML files in the default web browser
browseURL(html_file_ns_b)
browseURL(html_file_ns_ub)
```

```{r}

```

#the above process worked, but the html table doesn't look that good. Consider trying to make one large table, with all the sub tables included with sub-headers. That would work better I think?

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

# Understanding the Statistics

When you calculate an average difference between two temperatures and analyze the percent difference, average, and coefficient of variation, each metric provides different insights into your data:

1.  **Percent Difference**:

    -   **What it tells you**: The percent difference gives you a relative measure of the difference between the two temperatures compared to their average. It tells you how large the difference is in relation to the overall magnitude of the temperatures.

    -   **Interpretation**: A high percent difference indicates a significant discrepancy between the two temperatures, whereas a low percent difference suggests that the temperatures are relatively similar.

2.  **Average**:

    -   **What it tells you**: The average temperature difference provides a measure of the central tendency of the differences between the two temperatures. It is the mean value of all the calculated differences.

    -   **Interpretation**: This metric gives you an idea of the typical or expected difference between the temperatures. It is useful for understanding the overall trend in the temperature differences.

3.  **Coefficient of Variation (CV)**:

    -   **What it tells you**: The coefficient of variation is a standardized measure of dispersion of the temperature differences. It is calculated as the ratio of the standard deviation to the mean and is usually expressed as a percentage.

    -   **Interpretation**: A high CV indicates that the temperature differences are highly variable, suggesting inconsistency or high variability in the temperature differences. A low CV indicates that the temperature differences are relatively consistent and stable.

### Example Scenario

Suppose you have two sets of temperature readings (in °C) from two different sensors or locations. After calculating the differences between the corresponding readings, you obtain the following statistics:

-   **Average Difference**: 2°C

-   **Percent Difference**: 10%

-   **Coefficient of Variation**: 25%

#### Interpretation:

-   **Average Difference**: On average, there is a 2°C difference between the temperatures recorded by the two sensors or at the two locations.

-   **Percent Difference**: The 10% difference indicates that the temperature difference is relatively small compared to the average temperature values of the readings.

-   **Coefficient of Variation**: A CV of 25% suggests moderate variability in the temperature differences. While there is some fluctuation in the differences, it is not extremely high, indicating that the differences are somewhat consistent but not entirely uniform.

In summary, these metrics together help you understand the magnitude, relative size, and variability of the temperature differences in your data.

## My Understanding of the Data

-   very high coefficient of variation and percent differences mean that there is a lot of variability of temperatures, even on a daily and monthly scale.
