---
title: "treetemp_fullcode_clean"
author: "Erika Lee"
date: "2025-05-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#installing packages / set working directory
packages <- c('tidyverse',
              'lterdatasampler', 'dplyr', 'ggplot2', 'readr', 'readxl', 'lubridate', 'plotly', 'openxlsx', 'plotly', 'rstatix')

install.packages(packages)

library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

#adjust working directly as appropriate for where data is stored
setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

# Initial Data Processing

Processing composite datasets - includes 2-m sensors (2023-2025) and 1/0.5-m sensors (2024-2025)

-   2-m sensor data for burned sites did not change from 23/24 winter to 24/25 winter

-   2-m sensor data for unburned sites DID change between 23/24 and 24/25 -\> moved one tree's sensors to other tree, therefore the new composite unburned data (24/25 winter) includes both 2-m and 0.5-m sensor data

-   If processing -\> change location of where excel sheets are being read from to whatever your working directory is

## 2-m sensor Data

### PSZ Burned

```{r}
#read in temperature data
pers_b_tree_composite=read_excel("nsf/105E_pers_burned/105E_pers_b_composite.xlsx")

#filter temp data to full data frame
pers_b_tree_fulldata <- pers_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_b_aspects <- pers_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(5)",
      "Temp_C_Avg(9)",
      "Temp_C_Avg(13)", "Temp_C_Avg(15)"
    ),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(14)", "Temp_C_Avg(16)"),
    "south",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)", "Temp_C_Avg(11)"),
    "charred",
    aspect
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)", "Temp_C_Avg(12)"),
    "uncharred",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)",
      "Temp_C_Avg(3)",
      "Temp_C_Avg(4)"
    ),
    "gb_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(5)",
      "Temp_C_Avg(6)",
      "Temp_C_Avg(7)",
      "Temp_C_Avg(8)"
    ),
    "gb_e",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(9)",
      "Temp_C_Avg(10)",
      "Temp_C_Avg(11)",
      "Temp_C_Avg(12)"
    ),
    "gb_s",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(13)", "Temp_C_Avg(14)"),
    "db_n",
    tree_name
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(15)", "Temp_C_Avg(16)"),
    "db_s",
    tree_name
  ))
```

### PSZ Unburned

\*\* only works for data from 2023-2024, as the sensors were moved for 2024-25 winter

```{r}
#read in temperature data
pers_ub_tree_composite=read_excel("nsf/105E_pers_unburned/105E_pers_ub_composite.xlsx")

#filter temp data to full data frame
pers_ub_tree_fulldata <- pers_ub_tree_composite %>%
  mutate(datetime = as.POSIXct(TIMESTAMP, format=("%Y-%m-%d %H:%M:%S")), datetime = force_tz(datetime, tz = 'MST')) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
pers_ub_aspects <- pers_ub_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(3)"),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(4)"),
    "south",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)"
    ),
    "ub_n",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(3)", 
      "Temp_C_Avg(4)"
    ),
    "ub_s",
    tree_name
  ))
```

### TSZ Burned

```{r}
#read in temperature data
trans_b_tree_composite=read_excel("nsf/105E_trans_burned/105E_trans_b_composite.xlsx")

#filter temp data to full data frame
trans_b_tree_fulldata <- trans_b_tree_composite %>%
  mutate(datetime = ymd_hms(TIMESTAMP)) %>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP, Record)) %>%
  na.omit()

#adding in an aspect column 
trans_b_aspects <- trans_b_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(5)", "Temp_C_Avg(9)", "Temp_C_Avg(11)"),
    "north",
    "other"
  )) %>%
mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(2)", "Temp_C_Avg(6)", "Temp_C_Avg(10)", "Temp_C_Avg(12)"),
    "south",
    aspect)) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(3)", "Temp_C_Avg(7)"),
    "charred",
    aspect)) %>%
    mutate(aspect = ifelse(
    sensor_id %in% c("Temp_C_Avg(4)", "Temp_C_Avg(8)"),
    "uncharred",
    aspect)) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(1)", "Temp_C_Avg(2)", "Temp_C_Avg(3)", "Temp_C_Avg(4)"),
    "gb_n",
    "other"
  )) %>%
mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(5)", "Temp_C_Avg(6)", "Temp_C_Avg(7)", "Temp_C_Avg(8)"),
    "gb_s",
    tree_name)) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(9)", "Temp_C_Avg(10)"),
    "db_n",
    tree_name)) %>%
    mutate(tree_name = ifelse(
    sensor_id %in% c("Temp_C_Avg(11)", "Temp_C_Avg(12)"),
    "db_s",
    tree_name))
```

### TSZ Unburned

\*\* only works for data from 2023-2024, as the sensors were moved for 2024-25 winter

```{r}
trans_ub_tree_composite=read_excel("nsf/105E_trans_unburned/105E_trans_ub_composite.xlsx")

#filter temp data to full data frame
trans_ub_tree_fulldata <- trans_ub_tree_composite %>%
  mutate(datetime = as.POSIXct(TIMESTAMP, format = ("%Y-%m-%d %H:%M:%S")),
         datetime = force_tz(datetime, tz = 'MST'))%>%
  select(datetime, everything()) %>%
  select(-c(TIMESTAMP,RECORD))

#adding in an aspect column - WITH ONLY 14 SENSORS!
trans_ub_aspects <- trans_ub_tree_fulldata %>%
  pivot_longer(cols = starts_with("Temp"),
               names_to = "sensor_id",
               values_to = "temp") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(3)"),
    "north",
    "other"
  )) %>%
  mutate(aspect = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(2)",
      "Temp_C_Avg(4)"),
    "south",
    aspect
  )) %>%
  #adding in sensor name column
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(1)",
      "Temp_C_Avg(2)"
    ),
    "ub_w",
    "other"
  )) %>%
  mutate(tree_name = ifelse(
    sensor_id %in% c(
      "Temp_C_Avg(3)", 
      "Temp_C_Avg(4)"
    ),
    "ub_e",
    tree_name
  )) 
```

## 1 and 0.5-m sensor Data

PSZ Burned

```{r}
# bs stands for "below snow" - this is the 1 and 0.5 m data
bs_pers_b_alldata <- read_csv("nsf/105E_pers_burned/105E_pers_b_bs_composite.csv")
```

```{r}
bs_pers_b_aspects <- bs_pers_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "GB_E_N_1_Avg",
      "GB_E_N_05_Avg",
      "DB_N_N_1_Avg",
      "DB_N_N_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "GB_E") ~ "gb_e",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB_N") ~ "db_n",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "GB_E") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "DB") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "PSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

### PSZ Unburned

```{r}
bs_pers_ub_alldata <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_bs_composite.csv")
```

```{r}
bs_pers_ub_aspects <- bs_pers_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "PSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

TSZ Burned

```{r}
bs_trans_b_alldata <- read_csv("nsf/105E_trans_burned/105E_trans_b_bs_composite.csv")
```

```{r}
bs_trans_b_aspects <- bs_trans_b_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^GB|^DB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "gb_s_n_1_Avg",
      "gb_s_n_05_Avg",
      "db_s_n_1_Avg",
      "db_s_n_05_Avg"),
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = case_when(
    str_starts(tree_id, "gb_s") ~ "gb_s",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db_s") ~ "db_s",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  #add burn status column
   # Add tree_name column based on the start of tree_id
  mutate(burn_status = case_when(
    str_starts(tree_id, "gb") ~ "GB",  # If tree_id starts with "GB_E"
    str_starts(tree_id, "db") ~ "DB",  # If tree_id starts with "DB_N"
    TRUE ~ NA_character_  # If neither condition is met, set to NA
  )) %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "1_Avg") ~ 1,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(zone = "TSZ", day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```

Trans Unburned

```{r}
bs_trans_ub_alldata <- read_csv("nsf/105E_trans_unburned/105E_trans_ub_bs_composite.csv")
```

```{r}
bs_trans_ub_aspects <- bs_trans_ub_alldata %>%
  mutate(TIMESTAMP = paste0(TIMESTAMP, ":00")) %>%  # Add missing seconds
  mutate(datetime = mdy_hms(TIMESTAMP, tz = "MST")) %>%
  select(-TIMESTAMP) %>%
  select(datetime, everything()) %>%
  pivot_longer(cols = matches("^UB"),  # Use matches() for multiple prefixes
  names_to = "tree_id",
  values_to = "temp_C") %>%
  #add in aspect column
  mutate(aspect = ifelse(
    tree_id %in% c(
      "UB_N_N_2_Avg",
      "UB_N_N_05_Avg"), 
    "north",
    "south"
  )) %>%
  # Add tree_name column based on the start of tree_id
  mutate(tree_name = "ub_n", burn_status = "live", zone = "TSZ") %>%
  mutate(sensor_height_m = case_when(
    str_ends(tree_id, "2_Avg") ~ 2,  # If tree_id ends with "1_Avg"
    str_ends(tree_id, "05_Avg") ~ 0.5,  # If tree_id ends with "05_Avg"
    TRUE ~ NA_real_  # If neither condition is met, set to NA
  )) %>%
  mutate(day = as.Date(datetime, tz = "MST")) %>%
  select(datetime, day, zone, burn_status, tree_name, aspect, sensor_height_m, temp_C)
```
