---
title: "treetemp_datawrangling"
author: "Erika Lee"
date: "2024-06-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Wrangling

```{r}
library(tidyverse)
library(lterdatasampler)
library(dplyr)
library(ggplot2)
library(readr)
library(readxl)
library(lubridate)
library(plotly)
library(openxlsx)
library(plotly)
library(rstatix)
library(htmlwidgets)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(suncalc)

setwd("/Volumes/wcnr-network/Research/Kampf/Private/field_data")
```

```{r}
#to clear prior environment
rm(list = ls())
```

### Calling in data

Tree temp data

```{r}
#calling in tree temp data from n-drive
## burned
pers_b_aspects <- read_csv("nsf/105E_pers_burned/105E_pers_b_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_b_aspects <- read_csv("nsf/105E_trans_burned/105E_trans_b_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

pers_ub_aspects <- read_csv("nsf/105E_pers_unburned/105E_pers_ub_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = force_tz(datetime, tz = 'MST'))

trans_ub_aspects <- read_csv("nsf/105E_trans_unburned/105E_trans_ub_aspects.csv") %>%
  mutate(datetime = as.POSIXct(datetime, format=("%Y-%m-%d %H:%M:%S")),
                               datetime = with_tz(datetime, tz = 'MST'))
```

Weather station data:

NOTE - I need to rerun the wx_datawrangling Rmarkdown before calling in these dataframes if I change the range of dates I am looking for!

```{r}
transb_wx_15min_r <- read_csv('nsf/trans_burned/trans_burned_wx_15min_r.csv') %>%
  #have to force the MST otherwise it recognizes it as UTC
  mutate(datetime = force_tz(datetime, tz = 'MST'))

transub_wx_15min_r <- read_csv('nsf/trans_unburned/trans_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persb_wx_hourly_r <- read_csv('nsf/pers_burned/pers_burned_wx_hourly_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))

persub_wx_15min_r <- read_csv('nsf/pers_unburned/pers_unburned_wx_15min_r.csv') %>%
  mutate(datetime = force_tz(datetime, tz = 'MST'))
```

### Setting consistent timeframe

```{r}
#setting a consistent timeframe for all tree temp data for all burned/unburned zones
##has to start at 2024-02-02 due to errors in data at transitional burned site

filtered_pers_b_aspects <- pers_b_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_pers_ub_aspects <- pers_ub_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_b_aspects <- trans_b_aspects %>%
 mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_trans_ub_aspects <- trans_ub_aspects %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

```{r}
#setting consistent timeframe for weather data
#filtered weather dataframes
filtered_persb_wx_hourly <- persb_wx_hourly_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_persub_wx_15min <- persub_wx_15min_r %>%
  select(datetime, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
  
filtered_transb_wx_15min <- transb_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))

filtered_transub_wx_15min <- transub_wx_15min_r %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST")) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST"))
```

# Tree Temp Comparison

```{r}
#adding in a burn status column to the aspect dataframes
filtered_pers_b_aspects <- filtered_pers_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "PSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-05-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_b_aspects <- filtered_trans_b_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "gb") ~ "green burn",
    startsWith(tree_name, "db") ~ "dead burn",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  #adding a zone column
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

filtered_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  mutate(burn_status = case_when(
    startsWith(tree_name, "ub") ~ "live",
    TRUE ~ NA_character_  # Optional: handle other cases if needed
  )) %>%
  mutate(zone = "TSZ") %>%
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  mutate(snow_phase = case_when(
      day >= as.Date("2024-04-06") ~ "ablation",
      TRUE ~ "accumulation"
    )) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, sensor_id, tree_name, everything()) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

## Differencing Dead burn and Green burn trees - For Plotting later!

### Persistent Zone

```{r}
#creating a dataframe with green burn temp minus dead burn temp for the same aspects
##north aspect
persb_north_diff_data <- filtered_pers_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'north') %>%
  # Filter out gb_w tree
  filter(tree_name != 'gb_w') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_e_minus_db_s = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_e_minus_db_n = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
persb_north_diff_long <- persb_north_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "persistent") %>%
  mutate(aspect = "north")
```

North Aspect Boxplots

```{r}
# Plotting the boxplots
persb_north_diff_plot <- ggplot(persb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(persb_north_diff_plot)
```

```{r}
##south aspect
persb_south_diff_data <- filtered_pers_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'south') %>%
  # Filter out gb_w tree
  filter(tree_name != 'gb_w') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_e_minus_db_s = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_e_minus_db_n = mean(temp[tree_name == 'gb_e']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
persb_south_diff_long <- persb_south_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "persistent") %>%
  mutate(aspect = "south")
```

South Aspect Boxplots

```{r}
# Plotting the boxplots
persb_south_diff_plot <- ggplot(persb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(persb_south_diff_plot)
```

Combining Plots

```{r}
#combinging north and south plots
persb_combined_aspectdiff_treetemp_plots <- persb_north_diff_plot + persb_south_diff_plot + plot_layout(nrow = 2) + plot_annotation(title = "Persistent Zone, Green & Dead Burn Tree Temp Difference")

persb_combined_aspectdiff_treetemp_plots 

#saving combined plot
ggsave("persb_combined_aspectdiff_treetemp_plots.png", plot = persb_combined_aspectdiff_treetemp_plots , width = 8, height = 10, dpi = 300, bg = "white")
```

### Transitional Zone

```{r}
#creating a dataframe with green burn temp minus dead burn temp for the same aspects
##north aspect
transb_north_diff_data <- filtered_trans_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'north') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_n_minus_db_s = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_n_minus_db_n = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
transb_north_diff_long <- transb_north_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "transitional") %>%
  mutate(aspect = "north")
```

North Aspect Boxplots

```{r}
# Plotting the boxplots
transb_north_diff_plot <- ggplot(transb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(transb_north_diff_plot)
```

```{r}
##south aspect
transb_south_diff_data <- filtered_trans_b_aspects %>%
  # Filter for north aspect
  filter(aspect == 'south') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    gb_n_minus_db_s = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_s']),
    gb_s_minus_db_s = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_s']),
    gb_n_minus_db_n = mean(temp[tree_name == 'gb_n']) - mean(temp[tree_name == 'db_n']),
    gb_s_minus_db_n = mean(temp[tree_name == 'gb_s']) - mean(temp[tree_name == 'db_n'])
  ) %>%
  ungroup()

# Reshape data to long format
transb_south_diff_long <- transb_south_diff_data %>%
  pivot_longer(cols = starts_with("gb"), names_to = "comparison", values_to = "difference") %>%
  #adding in a zone and a burn status for later dataframe combinations
  mutate(burn_status = "burned") %>%
  mutate(zone = "transitional") %>%
  mutate(aspect = "south")
```

South Aspect Boxplots

```{r}
# Plotting the boxplots
transb_south_diff_plot <- ggplot(transb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(transb_south_diff_plot)
```

Combining Plots

```{r}
#combinging north and south plots
transb_combined_aspectdiff_treetemp_plots <- transb_north_diff_plot + transb_south_diff_plot + plot_layout(nrow = 2) + plot_annotation(title = "Transitional Zone, Green & Dead Burn Tree Temp Difference")

transb_combined_aspectdiff_treetemp_plots 

#saving combined plot
ggsave("transb_combined_aspectdiff_treetemp_plots.png", plot = transb_combined_aspectdiff_treetemp_plots , width = 8, height = 10, dpi = 300, bg = "white")
```

Combining Dataframes

```{r}
#creating a combined dataframe for the db-gb comparisons
full_dbgb_comparisons_data <- bind_rows(persb_north_diff_long, persb_south_diff_long, transb_north_diff_long, transb_south_diff_long)
```

Saving Combined dataframe

```{r}
write_csv(full_dbgb_comparisons_data, "nsf/treetemp_data/full_dbgb_comparisons_data.csv")
```

## Differencing by Individual Tree Aspect

### Persistent Zone

```{r}
#creating a dataframe for each tree
persb_gb_e_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_e') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_e')

persb_gb_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

persb_db_n_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

persb_db_s_diff_data <- filtered_pers_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
persb_gbdb_combined_data <- bind_rows(persb_gb_e_diff_data, persb_gb_s_diff_data, persb_db_n_diff_data, persb_db_s_diff_data)
```

Plotting

```{r}
#plotting persistent north-south tree boxplots
persb_gbdb_diff_plot <- ggplot(persb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Persistent Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(persb_gbdb_diff_plot)
```

### Transitional Zone

```{r}
#creating a dataframe for each tree
transb_gb_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_n')

transb_gb_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'gb_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'gb_s')

transb_db_n_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_n') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_n')

transb_db_s_diff_data <- filtered_trans_b_aspects %>%
  # Filter for gb_e tree
  filter(tree_name == 'db_s') %>%
  # Calculate differences by datetime
  group_by(datetime) %>%
  summarise(
    south_minus_north = mean(temp[aspect == 'south']) - mean(temp[aspect == 'north'])) %>%
  ungroup() %>%
  pivot_longer(cols = starts_with("south"), names_to = "comparison", values_to = "difference") %>%
  mutate(treename = 'db_s')

#creating a combined new dataframe
transb_gbdb_combined_data <- bind_rows(transb_gb_n_diff_data, transb_gb_s_diff_data, transb_db_n_diff_data, transb_db_s_diff_data)
```

Plotting

```{r}
#plotting transitional north-south tree boxplots
transb_gbdb_diff_plot <- ggplot(transb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Transitional Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(transb_gbdb_diff_plot)
```

Combined Plots

```{r}
#combinging north and south plots
combined_bytree_aspect_treetemp_plots <- persb_gbdb_diff_plot + transb_gbdb_diff_plot + plot_layout(nrow = 1) + plot_annotation(title = "Temperature Differences by Tree - south minus north aspect")

combined_bytree_aspect_treetemp_plots

#saving combined plot
ggsave("combined_bytree_aspect_treetemp_plots.png", plot = combined_bytree_aspect_treetemp_plots , width = 10, height = 6, dpi = 300, bg = "white")
```

Combined NS Dataframe

```{r}
#creating a full dataframe for both pers and transitional data for later plotting of the south-north comparison

##individually editing pers and trans combined data before combining both dataframes
# Add new column 'burn_status' based on conditions, and new column 'zone' based on location
persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "persistent")

transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  mutate(burn_status = case_when(
    startsWith(treename, "gb") ~ "green_burned",
    startsWith(treename, "db") ~ "dead_burned",
    TRUE ~ NA_character_  # Handle other cases if needed
  )) %>%
  mutate(zone = "transitional")

#creating a full combined dataset
full_ns_comparison_data <- bind_rows(persb_gbdb_combined_data, transb_gbdb_combined_data)
```

```{r}
#exporting full ns comparison dataframe
write_csv(full_ns_comparison_data, "nsf/treetemp_data/full_ns_comparison_data.csv")
```

**Interpretation of box plots**

-   Consistent Boxes: Similar boxes indicate that the general trend and variability within the middle 50% of the data are consistent across different aspects.

-   Many Outliers: The numerous outliers highlight that there are frequent instances of temperature differences that are much higher or lower than the typical range. This could point to specific factors or conditions that cause these deviations.

# Pivoting to long dataframes for temp comparisons

```{r}
#creating a combined new dataframe from original dataframes with individual tree temps - May not want a full combined dataset!
full_burned_combined_data <- filtered_pers_b_aspects %>%
  full_join(filtered_trans_b_aspects)

full_unburned_combined_data <- filtered_pers_ub_aspects %>%
  full_join(filtered_trans_ub_aspects)
```

Combined Full, Filtered Aspect/temp dataframes

```{r}
#exporting important dataframes to N-Drive. These are what are used for stats calculations
write_csv(full_burned_combined_data, "nsf/treetemp_data/full_burned_combined_data.csv")

write_csv(full_unburned_combined_data, "nsf/treetemp_data/full_unburned_combined_data.csv")
```

## N/S Aspect Dataframes

\*\* need to create a dataframe for only the gb trees and do the following comparisons, because I have to subtract it form the north/south aspects dataframes because deadburned trees do not have charred/uncharred aspect temps

\*\* Use the following final_hourly_GB temps dataset to do plotting of charred/uncharred comparisons, as well as for stats calculations!

```{r}
#green burned dataframe
full_gb_data <- full_burned_combined_data %>%
  filter(burn_status == "green burn") %>%
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_gb_temps <- full_gb_data%>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south, charred_temp = charred, uncharred_temp = uncharred) %>%
  summarize(north_temp, south_temp, charred_temp, uncharred_temp)

final_hourly_gb_temps <-final_gb_temps %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%  # Create an hourly datetime column
  group_by(datetime_hour, day, month, zone, snow_phase, burn_status) %>%  # Group by the hourly datetime
  #mean_hourly_treetemp is the mean of all the db/gb/live trees in the dataframe averaged by burn_status on an hourly timestep
  summarize(north_temp = mean(north_temp), south_temp = mean(south_temp), charred_temp = mean(charred_temp), uncharred_temp = mean(uncharred_temp), ns_avg_temp = ((north_temp + south_temp)/2), cuc_avg_temp = ((charred_temp + uncharred_temp)/2), sc_avg_temp = ((south_temp + charred_temp)/2), suc_avg_temp = ((south_temp + uncharred_temp)/2), nc_avg_temp = ((north_temp + charred_temp)/2), nuc_avg_temp = ((north_temp + uncharred_temp)/2), ns_temp_diff = (south_temp - north_temp), cuc_temp_diff = (charred_temp - uncharred_temp), sc_temp_diff = (south_temp - charred_temp), suc_temp_diff = (south_temp - uncharred_temp), nc_temp_diff = (north_temp - charred_temp), nuc_temp_diff = (north_temp - uncharred_temp), .groups = 'drop')
```

```{r}
#writing this gb csv to use in stats calculations!

##hourly
write_csv(final_hourly_gb_temps, "nsf/treetemp_data/final_hourly_gb_temps.csv")

##15-minute
write_csv(final_gb_temps, "nsf/treetemp_data/final_15min_gb_temps.csv")
```

### Pers_b

Creating a clean dataframe with columns for north and south temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree, so there is two gb and two db
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_persb_ns_temps <- full_pers_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

### Pers_ub

```{r}
#persistent unburned final dataset
full_pers_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_pers_ub_ns_temps <- full_pers_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_pers_ns_data <- final_persb_ns_temps %>%
  full_join(final_pers_ub_ns_temps)
```

### Trans_b

```{r}
#transitional burned 
full_trans_ns_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_transb_ns_temps <- full_trans_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

### Trans_ub

```{r}
#tranitional unburned final dataset
full_trans_ub_ns_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop')

final_trans_ub_ns_temps <- full_trans_ub_ns_data %>%
  group_by(datetime, day, month, burn_status, zone, snow_phase) %>%
  pivot_wider(names_from = aspect, values_from = mean_temp) %>%
  rename(north_temp = north, south_temp = south) %>%
  summarize(north_temp, south_temp)
```

```{r}
#creating a full n/s green burned, db, ub dataframe to do comparisons on later
## USE THIS DATAFRAME FOR FURTHER DATA ANALYSIS
full_trans_ns_data <- final_transb_ns_temps %>%
  full_join(final_trans_ub_ns_temps)
```

## DB/GB/UB Dataframes \*\*\*

### Pers_b

Creating a clean dataframe with columns for gb, db, and ub temps, instead of rows (wide format)

```{r}
#persistent burned 
full_pers_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "PSZ") %>%
  #remove gb_w tree
  filter(!(tree_name == "gb_w")) %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_gbdb_temps <- full_pers_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))
```

### Trans_b

```{r}
#transitional burned 
full_trans_gbdb_data <- full_burned_combined_data %>%
  filter(!(aspect %in% c("charred", "uncharred"))) %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_gbdb_temps <- full_trans_gbdb_data %>%
  group_by(datetime, day, month, aspect, zone, snow_phase) %>%
  pivot_wider(names_from = burn_status, values_from = mean_temp) %>%
  rename(gb_temp = "green burn", db_temp = "dead burn") %>%
  summarize(gb_temp, db_temp)
```

### Pers_ub

```{r}
#persistent unburned final dataset
full_pers_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "PSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_pers_ub_temps <- full_pers_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(datetime, day, month, aspect, zone, snow_phase, ub_temp) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

### Trans_ub

```{r}
#transitional unburned final dataset
full_trans_ub_data <- full_unburned_combined_data %>%
  mutate(burn_status = "live") %>%
  filter(zone == "TSZ") %>%
  #getting mean temperature of all trees, instead of having individual trees identified
  group_by(datetime, day, month, aspect, zone, snow_phase, burn_status) %>%
  # Calculate the mean temperature for each aspect
  summarize(mean_temp = mean(temp, na.rm = TRUE), .groups = 'drop') %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S", tz = "MST"))

final_trans_ub_temps <- full_trans_ub_data %>%
  rename(ub_temp = mean_temp) %>%
  select(-burn_status) %>%
  mutate(
    datetime = floor_date(datetime, unit = "15 minutes")  # Rounds down to nearest 15 minutes
  )
```

```{r}
final_pers_gbdb_temps <- final_pers_gbdb_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))

final_pers_ub_temps <- final_pers_ub_temps %>%
  mutate(datetime = force_tz(datetime, tz = "MST"))
```

```{r}
full_pers_gbdbub_data <- final_pers_gbdb_temps %>%
  left_join(final_pers_ub_temps)
```

```{r}
full_trans_gbdbub_data <- final_trans_gbdb_temps %>%
  left_join(final_trans_ub_temps)
```

```{r}
full_15min_gbdbub <- rbind(full_pers_gbdbub_data, full_trans_gbdbub_data)
```

```{r}
write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")

write_csv(full_pers_gbdbub_data, "nsf/treetemp_data/full_pers_gbdbub_data.csv")

write_csv(full_trans_gbdbub_data, "nsf/treetemp_data/full_trans_gbdbub_data.csv")
```

# Final 15-minute Dataframes

Combining burned and unburned dataframes by zone and 15-minute timestep - this is what is used for stats calculations

-   important step
-   Also adding avg_temp for 15-minute timestep

## North/South Aspects

```{r}
#combining persistent and transitional unburned to create a 15min_unburned datatable

final_15min_ns_ub <- final_pers_ub_ns_temps %>% 
  full_join(final_trans_ub_ns_temps) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((south_temp + north_temp) / 2)) %>%
  mutate(burn_status = "live") %>%
  select(datetime, day, month, zone, snow_phase, burn_status, everything())
```

```{r}
#combining persistent and transitional burnedto create a 15min_burned datatable
final_15min_ns_b <- final_persb_ns_temps %>%
  full_join(final_transb_ns_temps) %>%
  group_by(datetime, day, month, zone, snow_phase, burn_status) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((south_temp + north_temp) / 2)) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, everything())
```

```{r}
#adding in an hourly version of this final_ns dataframe
##unburned
final_hourly_ns_ub <- final_15min_ns_ub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, burn_status) %>%
  summarize(north_temp = mean(north_temp, na.rm = TRUE), south_temp = mean(south_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
final_hourly_ns_b <- final_15min_ns_b %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, burn_status) %>%
  summarize(north_temp = mean(north_temp, na.rm = TRUE), south_temp = mean(south_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#writing final n/s combined dataframes to n-drive
write_csv(final_15min_ns_b, "nsf/treetemp_data/final_15min_ns_b.csv")

write_csv(final_15min_ns_ub, "nsf/treetemp_data/final_15min_ns_ub.csv")

write_csv(final_hourly_ns_b, "nsf/treetemp_data/final_hourly_ns_b.csv")

write_csv(final_hourly_ns_ub, "nsf/treetemp_data/final_hourly_ns_ub.csv")
```

## GB/DB/UB Aspects

\*\* keep working on this - 10/18/24!

```{r}
# GB/DB comparison
final_15min_gbdb <- final_pers_gbdb_temps %>%
  full_join(final_trans_gbdb_temps) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((gb_temp + db_temp) / 2)) %>%
  select(datetime, day, month, zone, snow_phase, everything())

#adding in an hourly version of this final_15min_gbdb dataframe
final_hourly_gbdb <- final_15min_gbdb %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, aspect) %>%
  summarize(gb_temp = mean(gb_temp, na.rm = TRUE), db_temp = mean(db_temp, na.rm = TRUE), avg_temp = mean(avg_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#creating dataframe for combined ub temps
final_15min_gbdb_ub <- final_pers_ub_temps %>%
  full_join(final_trans_ub_temps) %>%
  #adding in day & month
  mutate(day = as.POSIXct(format(datetime, "%Y-%m-%d")),
         month = as.numeric(format(datetime, "%m"))) %>%
  select(datetime, day, month, zone, snow_phase, everything()) 

#adding in an hourly version of this final_15min_ub dataframe
final_hourly_gbdb_ub <- final_15min_gbdb_ub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, aspect) %>%
  summarize(ub_temp = mean(ub_temp, na.rm = TRUE), .groups = 'drop')
```

```{r}
#creating a combined gbdbub full dataframe
## this is an important step in data wrangling - need this for final csvs
full_15min_gbdbub <- final_15min_gbdb %>%
  select(-avg_temp) %>%
  full_join(final_15min_gbdb_ub) %>%
  #remove weird NAs
  drop_na()
```

```{r}
write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

```{r}
#GB/UB comparison

##have to create a new dataframe for this step that gets rid of db temps
final_gb_temps <- final_15min_gbdb %>%
  select(c(-db_temp, -avg_temp))

final_15min_gbub <- final_gb_temps %>%
  full_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "aspect", "zone")) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((gb_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything())

#adding in an hourly version of this final_15min_gbdb dataframe
final_hourly_gbub <- final_15min_gbub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, snow_phase, aspect) %>%
  summarize(gb_temp = mean(gb_temp, na.rm = TRUE), 
            # Replace ub_temp with the correct column name if needed
            ub_temp = mean(ub_temp, na.rm = TRUE), 
            avg_temp = mean(avg_temp, na.rm = TRUE), 
            .groups = 'drop')
```

```{r}
#DB/UB comparison
final_db_temps <- final_15min_gbdb %>%
  select(c(-gb_temp, -avg_temp))

final_15min_dbub <- final_db_temps %>%
  full_join(final_15min_gbdb_ub, by = c("datetime", "day", "month", "aspect", "zone")) %>%
  #adding in mean temp of north and south temps
  mutate(avg_temp = mean((db_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything())

final_hourly_dbub <- final_15min_dbub %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, aspect) %>%
  summarize(db_temp = mean(db_temp, na.rm = TRUE), 
            # Replace ub_temp with the correct column name if needed
            ub_temp = mean(ub_temp, na.rm = TRUE), 
            avg_temp = mean(avg_temp, na.rm = TRUE), 
            .groups = 'drop')
```

### Writing CSV

\*\* don't think I need any of these exports!

```{r}
#writing final gb/db/ub 15-minute comparison dataframes to n-drive
write_csv(final_15min_gbdb, "nsf/treetemp_data/final_15min_gbdb.csv")

write_csv(final_hourly_gbdb, "nsf/treetemp_data/final_hourly_gbdb.csv")

#this one is only the unburned temps! Not a comparison
write_csv(final_15min_gbdb_ub, "nsf/treetemp_data/final_15min_gbdb_ub.csv")

write_csv(final_hourly_gbdb_ub, "nsf/treetemp_data/final_hourly_gbdb_ub.csv")

write_csv(final_15min_gbub, "nsf/treetemp_data/final_15min_gbub.csv")

write_csv(final_hourly_gbub, "nsf/treetemp_data/final_hourly_gbub.csv")

write_csv(final_15min_dbub, "nsf/treetemp_data/final_15min_dbub.csv")

write_csv(final_hourly_dbub, "nsf/treetemp_data/final_hourly_dbub.csv")
```

# Day/Night timeframes

## Creating day-only dataframes

```{r}
#persistent daytime dataframes
dt_pers_b_aspects <- filtered_pers_b_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_pers_ub_aspects <- filtered_pers_ub_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

#transitional daytime dataframes
dt_trans_b_aspects <- filtered_trans_b_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_trans_ub_aspects <- filtered_trans_ub_aspects %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

## Plotting - DB/GB differencing

### Persistent Zone

```{r}
#manipulating the dataset to only be hours between 6 am and 9 pm MST
dt_persb_north_diff_long <- persb_north_diff_long %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)

dt_persb_south_diff_long <- persb_south_diff_long %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

Plotting

```{r}
#plotting
dt_persb_north_diff_plot <- ggplot(dt_persb_north_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime North Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(dt_persb_north_diff_plot)

dt_persb_south_diff_plot <- ggplot(dt_persb_south_diff_long, aes(x = comparison, y = difference, fill = comparison)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime South Aspect",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Comparison")

print(dt_persb_south_diff_plot)
```

## Plotting - aspect differencing

### Persistent Zone

```{r}
#creating new dataframe with only daytime hours
dt_persb_gbdb_combined_data <- persb_gbdb_combined_data %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

```{r}
#plotting
dt_persb_gbdb_diff_plot <- ggplot(dt_persb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime Persistent Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(dt_persb_gbdb_diff_plot)
```

### Transitional Zone

```{r}
#creating new dataframe with only daytime hours
dt_transb_gbdb_combined_data <- transb_gbdb_combined_data %>%
  filter(hour(datetime) >= 6 & hour(datetime) <= 21)
```

Plotting

```{r}
#plotting
dt_transb_gbdb_diff_plot <- ggplot(dt_transb_gbdb_combined_data, aes(x = comparison, y = difference, fill = treename)) +
  geom_boxplot() +
  theme_minimal() +
  labs(title = "Daytime Transitional Zone",
       x = "Comparison",
       y = "Temperature Difference (°C)",
       fill = "Tree Name")

print(dt_transb_gbdb_diff_plot)
```

# Day vs Night Stats Dataframes

Data Wrangling V2 - with proper solar sunrise/sunset applied

\*\* Using final full_15min_ns data that is from the end of this rmd! Important to note! Because it has the correct phases on it based on sunrise/sunset

```{r}
##this dataframe that I am starting from is created below this in the code! At the end of the phase section towards the bottom!
dtnt_full_15min_ns <- full_15min_ns_dtnt %>%
  pivot_wider(
    names_from = phase,
    values_from = c(north_temp, south_temp))

dt_alltemps <- dtnt_full_15min_ns %>%
  select(-c(north_temp_night, south_temp_night)) %>%
  filter(!is.na(north_temp_day) & !is.na(south_temp_day)) %>%
  select(datetime, day, month, zone, snow_phase, burn_status, north_temp_day, south_temp_day) %>%
  rename(dt_north_temp = "north_temp_day", dt_south_temp = "south_temp_day") %>%
  group_by(day, month, zone, snow_phase, burn_status) %>%
  summarize(dt_north_temp = mean(dt_north_temp), dt_south_temp = mean(dt_south_temp))

nt_alltemps <- dtnt_full_15min_ns %>%
  select(-c(north_temp_day, south_temp_day)) %>%
  filter(!is.na(north_temp_night) & !is.na(south_temp_night)) %>%
  select(datetime, day, month, zone, burn_status, north_temp_night, south_temp_night) %>%
  rename(nt_north_temp = "north_temp_night", nt_south_temp = "south_temp_night") %>%
  group_by(day, month, zone , burn_status) %>%
  summarize(nt_north_temp = mean(nt_north_temp), nt_south_temp = mean(nt_south_temp))
```

```{r}
final_dtnt_ns_temps <- dt_alltemps %>%
  full_join(nt_alltemps) %>%
  mutate(dt_avg_temp = ((dt_north_temp + dt_south_temp) / 2), nt_avg_temp = ((nt_north_temp + nt_south_temp) / 2), n_avg_temp = ((dt_north_temp + nt_north_temp) / 2), s_avg_temp = ((dt_south_temp + nt_south_temp) / 2)) %>%
  #changing zone names to be PSZ and TSZ
  mutate(zone = case_when(
    zone == "persistent" ~ "PSZ",
    zone == "transitional" ~ "TSZ",
    TRUE ~ zone))
```

```{r}
#calculating stats for these! Parsed at a daily mean timestep to compare day/night temps

full_dtnt_ns_temps <- final_dtnt_ns_temps %>%
  mutate(n_dtnt_temp_diff = (dt_north_temp - nt_north_temp), s_dtnt_temp_diff = (dt_south_temp - nt_south_temp))
```

```{r}
#exporting this csv - it has the correct sunset/sunrise based on date already!

write.csv(full_dtnt_ns_temps, "nsf/treetemp_data/full_dtnt_ns_temps.csv")
```

# Creating a single comparison dataframe for N/S and GB/DB/UB comparisons and stats

\*\* mean of all gb/db trees, averaged by burn_status

\*\* These are the final dataframes that I can pull stats from!

### 15-minute timestep

```{r}
full_15min_ns <- full_pers_ns_data %>%
  full_join(full_trans_ns_data) %>%
#adding in mean temp of n/s temps
  mutate(ns_avg_temp = ((north_temp + south_temp)/2), ns_temp_diff = (south_temp - north_temp))

#don't think I need the below exports!
write_csv(final_hourly_ns_b, "nsf/treetemp_data/final_hourly_ns_b.csv")

write_csv(final_hourly_ns_ub, "nsf/treetemp_data/final_hourly_ns_ub.csv")
```

```{r}
# GB/DB/UB comparison
full_15min_gbdbub <- full_15min_gbdbub %>%
  #adding in mean temp of gbdb/gbub/dbub temps
  mutate(gbdb_avg_temp = ((gb_temp + db_temp) / 2)) %>%
  mutate(gbub_avg_temp = ((gb_temp + ub_temp) / 2)) %>%
  mutate(dbub_avg_temp = ((db_temp + ub_temp) / 2)) %>%
  select(datetime, day, month, zone, everything()) %>%
  #calculating temp_diff for 15-minute interval
  mutate(gbdb_temp_diff = gb_temp - db_temp) %>%
  mutate(gbub_temp_diff = gb_temp - ub_temp) %>%
  mutate(dbub_temp_diff = db_temp - ub_temp)
```

Exporting full 15-minute gbdbub and n/s comparison dataframes

\*\* don't use the following export

```{r}
write_csv(full_15min_ns, "nsf/treetemp_data/full_15min_ns.csv")

write_csv(full_15min_gbdbub, "nsf/treetemp_data/full_15min_gbdbub.csv")
```

### hourly timestep

-   don't think I need this step either!

```{r}
full_hourly_ns<- final_15min_ns_b %>%
  full_join(final_15min_ns_ub) %>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>% 
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  group_by(datetime_hour, day, month, zone, burn_status) %>%
  summarize(north_temp = mean(north_temp), south_temp = mean(south_temp), ns_avg_temp = ((north_temp + south_temp)/2), ns_temp_diff = (south_temp - north_temp))
```

```{r}
#gb/db/ub comparisons
# Combine datasets and calculate hourly averages
full_hourly_gbdbub <- full_15min_gbdbub %>%
  mutate(datetime_hour = floor_date(datetime, unit = "hour")) %>%
  # Calculate temperature differences
  mutate(gbdb_temp_diff = gb_temp - db_temp, 
         gbub_temp_diff = gb_temp - ub_temp, 
         dbub_temp_diff = db_temp - ub_temp) %>%
  # Group by the new hourly datetime column
  group_by(datetime_hour, day, month, zone, aspect) %>%
  summarize(
    db_temp = mean(db_temp, na.rm = TRUE), 
    ub_temp = mean(ub_temp, na.rm = TRUE), 
    gb_temp = mean(gb_temp, na.rm = TRUE), 
    gbdb_avg_temp = mean(c(gb_temp, db_temp), na.rm = TRUE), 
    gbub_avg_temp = mean(c(gb_temp, ub_temp), na.rm = TRUE), 
    dbub_avg_temp = mean(c(db_temp, ub_temp), na.rm = TRUE), 
    gbdb_temp_diff = mean(gbdb_temp_diff, na.rm = TRUE), 
    gbub_temp_diff = mean(gbub_temp_diff, na.rm = TRUE), 
    dbub_temp_diff = mean(dbub_temp_diff, na.rm = TRUE), 
    .groups = 'drop'
  )
```

\*\*

```{r}
write_csv(full_hourly_ns, "nsf/treetemp_data/full_hourly_ns.csv")

write_csv(full_hourly_gbdbub, "nsf/treetemp_data/full_hourly_gbdbub.csv")
```

As of 2024-10-18 -\> I've cleaned up everything above this!

## Plotting charred/uncharred comparisons

```{r}
#loading in finalized dataframe 
full_hourly_gb <- read_csv("nsf/treetemp_data/full_hourly_gb.csv") %>%
  mutate(datetime_hour = as.POSIXct(datetime_hour, format=("%Y-%m-%d %H:%M:%S")),
                               datetime_hour = force_tz(datetime_hour, tz = 'MST'))  %>%
  select(-1)
```

```{r}
#making a daily dt/nt dataset
daily_final_gb_temps <- final_hourly_gb_temps %>%
  mutate(
    date = date(datetime_hour),  # Create the date column
    phase = case_when(
      hour(datetime_hour) >= 8 & hour(datetime_hour) < 17 ~ "day",
      TRUE ~ "night"
    )) %>%
  group_by(date, month, phase, zone, burn_status) %>%
  summarize(mean_north_temp = mean(mean_north_temp), mean_south_temp = mean(mean_south_temp), mean_charred_temp = mean(mean_charred_temp), mean_uncharred_temp = mean(mean_uncharred_temp))

#adding day/night columns to hourly gb_temps
final_hourly_gb_temps <- final_hourly_gb_temps %>%
  mutate(
    date = date(datetime_hour),  # Create the date column
    phase = case_when(
      hour(datetime_hour) >= 8 & hour(datetime_hour) < 17 ~ "day",
      TRUE ~ "night"
    )) %>%
  select(datetime_hour, date, month, phase, everything())
```

```{r}
#day/night mean temps plots
dtnt_gb_aspect_comparison <- ggplot(data = daily_final_gb_temps) +
  geom_line(aes(x = date, y = mean_north_temp, color = "North Temp")) +
  geom_line(aes(x = date, y = mean_south_temp, color = "South Temp")) +
  geom_line(aes(x = date, y = mean_charred_temp, color = "Charred Temp")) +
  geom_line(aes(x = date, y = mean_uncharred_temp, color = "Uncharred Temp")) +
  facet_wrap(~ zone + phase, dir = "v") +
  scale_color_manual(
    values = c("North Temp" = "green", "South Temp" = "red", "Charred Temp" = "black", "Uncharred Temp" = "grey"),
    name = "Temperature Type"  # Legend title
  ) + 
  facet_wrap(~phase + zone)


dtnt_gb_aspect_comparison 
```

The above plot makes it look like the charred temps are the highest, not sure why?

```{r}
#day/night hourly temps plots
hourly_gb_aspect_comparison <- ggplot(data = final_hourly_gb_temps) +
  geom_line(aes(x = date, y = mean_north_temp, color = "North Temp")) +
  geom_line(aes(x = date, y = mean_south_temp, color = "South Temp")) +
  geom_line(aes(x = date, y = mean_charred_temp, color = "Charred Temp")) +
  geom_line(aes(x = date, y = mean_uncharred_temp, color = "Uncharred Temp")) +
  facet_wrap(~ zone + phase, dir = "v") +
  scale_color_manual(
    values = c("North Temp" = "green", "South Temp" = "red", "Charred Temp" = "black", "Uncharred Temp" = "grey"),
    name = "Temperature Type"  # Legend title
  ) + 
  facet_wrap(~phase + zone)


hourly_gb_aspect_comparison 
```

### Box Plot of hourly data

\*\* USE THIS FINAL COMBINED PLOT (dt_hourly_gb_comb_plots) IN THESIS!

```{r}
## have to make a new, long, dataframe first
final_hourly_gb_temps_long <- full_hourly_gb %>%
  select(datetime_hour, day, month, phase, zone, burn_status, north_temp, south_temp, charred_temp, uncharred_temp) %>%
  #renameing zones with appreviations
  mutate(zone = case_when(
    zone == "persistent" ~ "PSZ",
    zone == "transitional" ~ "TSZ"
  )) %>%
  #renaming temp columns for long pivot
  rename(north = "north_temp", south = "south_temp", charred = "charred_temp", uncharred = "uncharred_temp" ) %>%
  pivot_longer(cols = c(north, south, charred, uncharred),
               names_to = "aspect", values_to = "temperature")
```

```{r}
##use this!
#creating a daytime dataframe and a night time dataframe to make the facet_wrapped plots more easy to see
##persistent
dt_p_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "day", zone == "PSZ")

nt_p_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "night", zone == "PSZ")

##transitional
dt_t_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "day", zone == "TSZ")

nt_t_final_hourly_gb_temps_long <- final_hourly_gb_temps_long %>%
  filter(phase == "night", zone == "TSZ")
```

```{r}
##use this plot!
#dt hourly boxplot - persistent
dt_p_hourly_gb_aspect_comparison_boxplot <- ggplot(data = dt_p_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = aspect, y = temperature, fill = aspect)) + 
  facet_wrap(~ month, nrow = 1) + 
  scale_fill_manual(
    values = c("north" = "#2166AC", "south" = "#D95F02", "charred" = "grey30", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(title = "a) PSZ", x = NULL, y = "Temperature (°C)") +
  scale_y_continuous(limits = c(-20, 30)) +  # Set the y-axis limits here
  theme_bw()

# Display the plot
print(dt_p_hourly_gb_aspect_comparison_boxplot)
```

```{r}
##use this plot!
dt_t_hourly_gb_aspect_comparison_boxplot <- ggplot(data = dt_t_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = aspect, y = temperature, fill = aspect)) + 
  facet_wrap(~ month, nrow = 1) + 
  scale_fill_manual(
    values = c("north" = "#2166AC", "south" = "#D95F02", "charred" = "grey30", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(title = "b) TSZ", x = "Aspect", y = "Temperature (°C)") +
  scale_y_continuous(limits = c(-20, 30)) +  # Set the y-axis limits here
  theme_bw()

# Display the plot
print(dt_t_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt hourly mean plots next to each other
dt_hourly_gb_comb_plots <- ggarrange(
  dt_p_hourly_gb_aspect_comparison_boxplot,
  dt_t_hourly_gb_aspect_comparison_boxplot,
  nrow = 2,
  common.legend = TRUE,
  legend = "right")

# adding a title to combined plot
#dt_hourly_gb_comb_plots <- annotate_figure(
  #dt_hourly_gb_comb_plots,
  #top = text_grob("Daytime Hourly Mean Temps", size = 14)
#)

# Display the combined plot
dt_hourly_gb_comb_plots
```

```{r}
#hourly boxplot
hourly_gb_aspect_comparison_boxplot <- ggplot(data = final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = aspect, y = temperature, fill = aspect)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(hourly_gb_aspect_comparison_boxplot)
```

Night Time Plots by zone

```{r}
#nt hourly boxplot - persistent
nt_p_hourly_gb_aspect_comparison_boxplot <- ggplot(data = nt_p_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(nt_p_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#nighttime hourly boxplot - transitional
nt_t_hourly_gb_aspect_comparison_boxplot <- ggplot(data = nt_t_final_hourly_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(nt_t_hourly_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt plots next to each other
nt_hourly_gb_comb_plots <- ggarrange(
  nt_p_hourly_gb_aspect_comparison_boxplot, 
  nt_t_hourly_gb_aspect_comparison_boxplot,
  nrow = 1,
  common.legend = TRUE,
  legend = "right")

nt_hourly_gb_comb_plots
```

Boxplots by Mean Daily Data

```{r}
## have to make a new, long, dataframe first
final_daily_gb_temps_long <- daily_final_gb_temps %>%
  #renaming aspect variables so it only says "south" not "mean_south_temp"
  rename_with(~ gsub("_temp", "", gsub("mean_", "", .)), starts_with("mean_")) %>%
  pivot_longer(cols = c(north, south, charred, uncharred),
               names_to = "temp_type", values_to = "temperature")
```

Day/Night Long Dataframes - for boxplot plotting

```{r}
dt_p_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "day", zone == "persistent")

nt_p_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "night", zone == "persistent")

##transitional
dt_t_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "day", zone == "transitional")

nt_t_final_daily_gb_temps_long <- final_daily_gb_temps_long %>%
  filter(phase == "night", zone == "transitional")
```

Plotting Daily Means by Phase + Zone

```{r}
#dt hourly boxplot - persistent
dt_p_daily_gb_aspect_comparison_boxplot <- ggplot(data = dt_p_final_daily_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_p_daily_gb_aspect_comparison_boxplot)
```

```{r}
#dt hourly boxplot - persistent
dt_t_daily_gb_aspect_comparison_boxplot <- ggplot(data = dt_t_final_daily_gb_temps_long) +
  geom_boxplot(aes(x = temp_type, y = temperature, fill = temp_type)) + 
  facet_wrap(~ zone + phase + month, scales = "free") + 
  scale_fill_manual(
    values = c("north" = "green", "south" = "red", "charred" = "black", "uncharred" = "grey"),
    name = "Aspect of Temp Sensor"  # Legend title
  ) + 
  labs(x = "Aspect of Temp Sensor", y = "Temperature (°C)") +
  theme_bw()

# Display the plot
print(dt_t_daily_gb_aspect_comparison_boxplot)
```

```{r}
#arranging dt plots next to each other
dt_daily_gb_comb_plots <- ggarrange(
  dt_p_daily_gb_aspect_comparison_boxplot,
  dt_t_daily_gb_aspect_comparison_boxplot,
  nrow = 1,
  common.legend = TRUE,
  legend = "right")

# adding a title to combined plot
dt_daily_gb_comb_plots <- annotate_figure(
  dt_daily_gb_comb_plots,
  top = text_grob("Daytime Daily Mean Temps", size = 14)
)

# Display the combined plot
dt_daily_gb_comb_plots
```

# Display the combined plot

dt_hourly_gb_comb_plots

```{r}
#new dataframe with charred/uncharred/north/south aspects for feb

feb_t_full_aspects <- filtered_trans_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(month == "2")

feb_p_full_aspects <- filtered_pers_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(tree_name !="gb_w") %>%
  filter(month == "2")

may_t_full_aspects <- filtered_trans_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(month == "5")

may_p_full_aspects <- filtered_pers_b_aspects %>%
  select(datetime, zone, burn_status, aspect, tree_name, temp) %>%
  mutate(month = month(datetime)) %>%
  filter(tree_name !="gb_w") %>%
  filter(month == "5")

```

```{r}
#plotting feb dates to see if there is a substantial difference between char patterns
t_feb_aspect_plot <- ggplot(data = feb_t_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

t_feb_aspect_plot

p_feb_aspect_plot <- ggplot(data = feb_p_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

p_feb_aspect_plot
```

```{r}
t_may_aspect_plot <- ggplot(data = may_t_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

t_may_aspect_plot

p_may_aspect_plot <- ggplot(data = feb_p_full_aspects) +
  geom_line(aes(x = datetime, y = temp, color = aspect, linetype = burn_status)) +
  facet_wrap(~tree_name)

p_may_aspect_plot
```

\*\* at the pers may plot -\> looks like charred trees are actually getting much warmer than in feb.

# Solar Exposure Calculations

```{r}
# Specify PSZ and TSZ locations
##pers_b - from 105E location!
p_b_lat <- 40.56379284
p_b_lon <- -105.867215
##pers_ub - from 105E locations
p_ub_lat <- 40.5638075
p_ub_lon <- -105.8705715

##trans_b - from 105E locations!
t_b_lat <- 40.60600629
t_b_lon <- -105.8333335

##trans_ub - from 105E locations!
t_ub_lat <- 40.568396
t_ub_lon <- -105.587383

## use the p_b_lat/long since all of these dates/times are almost identical for sunrise/sunset

# Create a date sequence from February 2nd to May 24th, 2024 for solar exposure
solex_dates <- seq(as.Date("2024-02-02"), as.Date("2024-07-01"), by = "day")

# Get solar times for each day
p_b_solar_times <- getSunlightTimes(date = solex_dates, lat = p_b_lat, lon = p_b_lon, tz = "MST") %>%
  mutate(
    day = as.Date(date),  # Ensure 'day' is of Date type for joining
    sunrise = with_tz(as.POSIXct(sunrise), tzone = "MST"),  # Set MST timezone
    sunset = with_tz(as.POSIXct(sunset), tzone = "MST")     # Set MST timezone
  )
```

```{r}
write_csv(p_b_solar_times, "nsf/treetemp_data/p_b_solar_times.csv")
```

## Adding Day/Night Phase to final CSVs

Full aspect CSV's with all sensors, not an average

```{r}
filtered_pers_b_aspects_r <- filtered_pers_b_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_pers_ub_aspects_r <- filtered_pers_ub_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_trans_b_aspects_r <- filtered_trans_b_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
filtered_trans_ub_aspects_r <- filtered_trans_ub_aspects %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, burn_status, aspect, sensor_id, tree_name, temp, solarNoon, sunrise, sunset, dawn, dusk)
```

```{r}
#exporting above dataframes
write.csv(filtered_pers_b_aspects_r, "nsf/treetemp_data/filtered_pers_b_aspects_r.csv")

write.csv(filtered_pers_ub_aspects_r, "nsf/treetemp_data/filtered_pers_ub_aspects_r.csv")

write.csv(filtered_trans_b_aspects_r, "nsf/treetemp_data/filtered_trans_b_aspects_r.csv")

write.csv(filtered_trans_ub_aspects_r, "nsf/treetemp_data/filtered_trans_ub_aspects_r.csv")
```

Sensor-averaged dataframes

```{r}
#15-minute NS data
full_15min_ns_dtnt <- full_15min_ns %>%
  mutate(
    day = as.Date.POSIXct(datetime, tz = "MST"),
    datetime = with_tz(datetime, tzone = "MST")
  ) %>%
  filter(datetime >= as.POSIXct("2024-02-02 00:00:00", tz = "MST") & 
         datetime < as.POSIXct("2024-07-01 00:00:00", tz = "MST")) %>%
  inner_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, burn_status, north_temp, south_temp, ns_avg_temp, ns_temp_diff, sunrise, sunset, solarNoon)
```

```{r}
#hourly NS data
full_hourly_ns_dtnt <- full_hourly_ns %>%
  #adjusting TZ for datetime and day columns
  mutate(
    day = as.Date.POSIXct(datetime_hour, tz = "MST"),
    datetime = with_tz(datetime_hour, tzone = "MST")
  ) %>%
  left_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime_hour >= sunrise & datetime_hour < sunset ~ "day",
      datetime_hour < sunrise ~ "night",
      datetime_hour >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime_hour, day, month, phase, zone, burn_status, north_temp, south_temp, ns_avg_temp, ns_temp_diff, sunrise, sunset, solarNoon)
```

```{r}
#15-minute dbgbub dataframe
#hourly NS data
full_15min_gbdbub_dtnt <- full_15min_gbdbub %>%
  #adjusting TZ for datetime and day columns
  mutate(
    day = as.Date.POSIXct(datetime, tz = "MST"),
    datetime = with_tz(datetime, tzone = "MST")
  ) %>%
  left_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime >= sunrise & datetime < sunset ~ "day",
      datetime < sunrise ~ "night",
      datetime >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime, day, month, phase, zone, snow_phase, aspect, gb_temp, db_temp, ub_temp, gbdb_avg_temp, gbub_avg_temp, dbub_avg_temp, gbdb_temp_diff, gbub_temp_diff, dbub_temp_diff, sunrise, sunset, solarNoon)
```

```{r}
#hourly GBDBUB data
full_hourly_gbdbub_dtnt <- full_hourly_gbdbub %>%
  #adjusting TZ for datetime and day columns
  mutate(
    day = as.Date.POSIXct(datetime_hour, tz = "MST"),
    datetime = with_tz(datetime_hour, tzone = "MST")
  ) %>%
  #adding in solar exposure dataframe with sunrise/sunset times
  left_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime_hour >= sunrise & datetime_hour < sunset ~ "day",
      datetime_hour < sunrise ~ "night",
      datetime_hour >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime_hour, day, month, phase, zone, snow_phase, aspect, gb_temp, db_temp, ub_temp, gbdb_avg_temp, gbub_avg_temp, dbub_avg_temp, gbdb_temp_diff, gbub_temp_diff, dbub_temp_diff, sunrise, sunset, solarNoon)
```

```{r}
#hourly Charred/Uncharred/North/south temps
##didn't do a 15-minute version of this!

full_hourly_gb_dtnt <- final_hourly_gb_temps %>%
  #adjusting TZ for datetime and day columns
  mutate(
    day = as.Date.POSIXct(datetime_hour, tz = "MST"),
    datetime = with_tz(datetime_hour, tzone = "MST")
  ) %>%
  #adding in solar exposure dataframe with sunrise/sunset times
  left_join(p_b_solar_times, by = "day") %>%
  #creating a phase column that parses by daily sunrise/sunset times
  mutate(
    phase = case_when(
      datetime_hour >= sunrise & datetime_hour < sunset ~ "day",
      datetime_hour < sunrise ~ "night",
      datetime_hour >= sunset ~ "night",
      TRUE ~ NA_character_)) %>%
  select(datetime_hour, day, month, phase, zone, snow_phase, burn_status, north_temp, south_temp, charred_temp, uncharred_temp, ns_avg_temp, cuc_avg_temp, sc_avg_temp, suc_avg_temp, nc_avg_temp, nuc_avg_temp, ns_temp_diff, cuc_temp_diff, sc_temp_diff, suc_temp_diff, nc_temp_diff, nuc_temp_diff, sunrise, sunset, solarNoon)
```

```{r}
#day/nigth comparison dataframe 
final_daynight_singleavg_temps
```

Exporting final dataframes

```{r}
write.csv(full_15min_ns_dtnt, "nsf/treetemp_data/full_15min_ns.csv")

#write.csv(full_hourly_ns_dtnt, "nsf/treetemp_data/full_hourly_ns.csv")

write.csv(full_15min_gbdbub_dtnt, "nsf/treetemp_data/full_15min_gbdbub.csv")

#write.csv(full_hourly_gbdbub_dtnt, "nsf/treetemp_data/full_hourly_gbdbub.csv")

write.csv(full_hourly_gb_dtnt, "nsf/treetemp_data/full_hourly_gb.csv")
```
